---
title: "Analysis of Cant estimates"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r parent, child = here::here("analysis/child/setup.Rmd")}
```

# Libraries

Loading libraries specific to the the analysis performed in this section.

```{r load_libraries_specific, message=FALSE, warning=FALSE}
library(marelac)
library(kableExtra)
```

# Data sources

## Sabine 2004


```{r read_S04_cant_file}

cant_inv_S04 <- 
  read_csv(here::here("data/interim",
                       "S04_cant_inv.csv"))

cant_inv_S04 <- cant_inv_S04 %>% 
  select(-eras)

cant_zonal_S04 <- 
  read_csv(here::here("data/interim",
                       "S04_cant_zonal.csv"))

cant_zonal_S04 <- cant_zonal_S04 %>%
  select(-eras)


```


## Gruber 2019

```{r read_G19_cant_file}

cant_inv_G19 <- 
  read_csv(here::here("data/interim",
                       "G19_cant_inv.csv"))

cant_inv_G19 <- cant_inv_G19 %>% 
  select(-eras)

cant_zonal_G19 <- 
  read_csv(here::here("data/interim",
                       "G19_cant_zonal.csv"))

cant_zonal_G19 <- cant_zonal_G19 %>%
  select(-eras)


```

# Comparison of previous estimates

Cant inventory estimates of S04 (Sabine et al, 2004) and G19 (Gruber et al, 2019) were compared.

## Merge data sets

```{r merge_cant_data_sets}

cant_inv <- full_join(cant_inv_G19 %>% mutate(estimate = "G19"),
                      cant_inv_S04 %>% mutate(estimate = "S04")) %>% 
  select(-cant_inv)

rm(cant_inv_G19, cant_inv_S04)

```

## Inventory maps

Spanning different time periods, the Cant inventories differ in magnitude.

```{r cant_inv_comparison, fig.asp=1}
 
map +
  geom_raster(data = cant_inv,
              aes(lon, lat, fill = cant_pos_inv)) +
  scale_fill_viridis_c() +
  facet_wrap( ~ estimate, ncol = 1) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )

```

## C~ant~ budgets

Global Cant inventories were estimated in Pg-C. Please note that here we only added positive Cant values in the upper `r parameters$inventory_depth`m and do not apply additional corrections for areas not covered.

```{r calculate_global_inventory}

cant_inv <- cant_inv %>% 
  mutate(surface_area = earth_surf(lat, lon),
         cant_pos_inv_grid = cant_pos_inv*surface_area)

cant_inv_budget <- cant_inv %>% 
  group_by(estimate, basin_AIP) %>% 
  summarise(cant_pos_total = sum(cant_pos_inv_grid)*12*1e-15,
            cant_pos_total = round(cant_pos_total,1)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = cant_pos_total, names_from = basin_AIP) %>% 
  mutate(total = Atlantic + Indian + Pacific)

cant_inv_budget %>% 
  kable() %>% 
  add_header_above() %>%
  kable_styling(full_width = FALSE)

```


## Relative inventories

```{r calculate_cant_inv_ratio}

cant_inv_wide <- cant_inv %>%
  pivot_wider(values_from = c(cant_pos_inv, cant_pos_inv_grid),
              names_from = estimate)

cant_inv_wide <- cant_inv_wide %>% 
  drop_na() %>% 
  mutate(G19_rel = cant_pos_inv_grid_G19 / sum(cant_pos_inv_grid_G19),
         S04_rel = cant_pos_inv_grid_S04 / sum(cant_pos_inv_grid_S04),
         cant_ratio_rel = G19_rel / S04_rel)

cant_inv_rel <- cant_inv_wide %>% 
  pivot_longer(cols = c(G19_rel, S04_rel),
                        names_to = "estimate",
                        values_to = "cant_inv_rel")

```



```{r cant_inv_rel_maps, fig.asp=1}

map +
  geom_raster(data = cant_inv_rel,
              aes(lon, lat, fill = cant_inv_rel*100)) +
  scale_fill_viridis_c() +
  facet_wrap( ~ estimate, ncol = 1) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )

```

## Relative inventory ratios

```{r cant_inv_ratio_map, fig.asp=0.6}

map +
  geom_contour_filled(data = cant_inv_wide %>%
  filter(cant_ratio_rel < 10,
         cant_ratio_rel > 0.1),
                      aes(lon, lat, z = log10(cant_ratio_rel))) +
  coord_quickmap(expand = 0) +
  scale_fill_brewer(palette = "RdBu", direction = -1) +
  labs(title = "Cant inventory distribution | 1994-2007 vs preind-1994",
       subtitle = "Log ratio of relative contributions to total inventory") +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank()
  )

```

