---
title: "World Ocean Atlas 2018"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(tidync)
library(stars)
```


```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Data source
- Data source: [WOCE](http://icdc.cen.uni-hamburg.de/daten/ocean/nsbc/index.php?id=woce_climatology&L=0){target="_blank"}

# data

## Read ncdfs

```{r read_WOA18}

# temperature

WOCE <- tidync(here::here("data/WOCE",
                          "wghc_params.nc"))

WOCE

WOCE_tibble <- WOCE %>% hyper_tibble()

WOCE_tibble <- WOCE_tibble  %>% 
  select(LON, LAT, ZAX, TEMP, SALINITY, GAMMAN) %>% 
  drop_na()

```


```{r set_criteria_horizons_sections}

depth_surface_selection <- c(0)
Atl_lon <- 335.5 # subtract 360 from value used for GLODAP climatology
Pac_lon <- 190.5 # subtract 360 from value used for GLODAP climatology

```

Below, following subsets of the climatologies are plotted for all relevant parameters:  

- Horizontal planes at `r depth_surface_selection`m
- Meridional sections at longitudes:
  - Atlantic: `r Atl_lon`
  - Pacific: `r Pac_lon`

Section locations are indicated as white lines in maps.

Please note that longitudes in the climatologies range from -179.5 - 179.5, which is different from GLODAP mapped climatologies.



## Temperature plots

### Surface map

```{r temperature_surface_WOCE_map}

WOCE_tibble %>% 
  filter(ZAX == 0) %>% 
  ggplot(aes(LON, LAT, fill = TEMP)) +
  geom_raster() +
  geom_vline(xintercept = c(Atl_lon, Pac_lon), col = "white") +
  coord_quickmap(expand = 0) +
  scale_fill_viridis_c() +
  theme(legend.position = "top")


```

### Section

```{r temperature_section_WOCE_map}

WOCE_tibble %>% 
  filter(LON == Atl_lon) %>% 
  ggplot(aes(LAT, ZAX, z = TEMP)) +
  geom_contour_filled() +
  scale_y_reverse() +
  coord_cartesian(expand = 0) +
  theme(legend.position = "top")

```



## Neutral density plots

### Surface map

```{r gamma_surface_WOCE_map}

WOCE_tibble %>% 
  filter(ZAX == 0) %>% 
  ggplot(aes(LON, LAT, fill = GAMMAN)) +
  geom_raster() +
  geom_vline(xintercept = c(Atl_lon, Pac_lon), col = "white") +
  coord_quickmap(expand = 0) +
  scale_fill_viridis_c() +
  theme(legend.position = "top")


```

### Section

```{r gamma_section_WOCE_map}

WOCE_tibble %>% 
  filter(LON == Atl_lon) %>% 
  ggplot(aes(LAT, ZAX, z = GAMMAN)) +
  geom_contour_filled() +
  scale_y_reverse() +
  coord_cartesian(expand = 0) +
  theme(legend.position = "top")

```



## Write file

```{r write_WOA_predictor_file, eval=FALSE}

# WOA18_predictors <- full_join(WOA_sal_tibble, WOA_tem_tibble)
# 
# WOA18_predictors %>% 
#   write_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
#                        "WOA18_predictors.csv"))
# 
# rm(WOA18_predictors, WOA_sal, WOA_sal_tibble, WOA_tem, WOA_tem_tibble)

```


# Open tasks

- basin mask not yet applied to WOA18 data

# Questions

- Which version of the WOA to be used
  - Fields (currently: objectively analyzed mean)
  - Decades (currently: oall decades)
  - Grid (currently: o1 deg resolution)
- How to merge with GLODAP climatology (currently interpolated to GLODAP depth)
