---
title: "World Ocean Atlas 2013 - D. Clement"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = here::here("analysis/child/setup.Rmd")}
```

# Libraries

Loading libraries specific to the the analysis performed in this section.

```{r load_libraries_specific, message=FALSE, warning=FALSE}
library(tidync)
```


# Data source

Dominic Clement provided a netcdf file with the basin mask and neutral densities used in Clement and Gruber (2018), both derived from the World Ocean Atlas 2013.

# Read ncdfs

```{r read_WOA13_dclement}


nd_mask <- tidync(here::here("data/input/World_Ocean_Atlas_2013_Clement",
                             "nd_mask.nc"))

nd_mask_tibble <- nd_mask %>% hyper_tibble()

nd_mask_tibble <- nd_mask_tibble %>%
  mutate(gamma = if_else(gamma == -999, NaN, gamma)) %>%
  drop_na()

nd_mask_tibble <- nd_mask_tibble %>%
  rename(lat = latitude,
         lon = longitude) %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))



```

# Apply basin mask


# Plots

Below, following subsets of the climatologies are plotted for all relevant parameters:  

- Horizontal planes at `r parameters$depth_levels`m
- Meridional sections at longitudes: `r parameters$longitude_sections_basin`

Section locations are indicated as white lines in maps.

## Basin map

Follwoing basin mask is not further used in this project.

```{r basin_map_WOA13_dclement, fig.asp=0.6}

nd_mask_tibble %>%
  filter(depth == 0) %>%
  ggplot(aes(lon, lat, fill = as.factor(mask))) +
  geom_raster() +
  geom_vline(
    xintercept = c(
      parameters$lon_Atl_section,
      parameters$lon_Pac_section,
      parameters$lon_Ind_section
    ),
    col = "white"
  ) +
  coord_quickmap(expand = 0) +
  scale_fill_brewer(palette = "Set1",
                    name = "basin mask") +
  theme(legend.position = "top",
        axis.title = element_blank())

```

## Neutral density

### Maps

```{r neutral_density_surface_WOA18_map}

map_climatology(nd_mask_tibble, "gamma")

```

### Sections

```{r neutral_density_section_WOA18}

section_global(nd_mask_tibble, "gamma", "WOA13")

```



## Write file

```{r write_WOA13_dclement_file, eval=FALSE}

nd_mask_tibble %>%
  write_csv(here::here("data/World_Ocean_Atlas_2013_Clement/_summarized_files",
                       "WOA13_mask_gamma.csv"))

```
