



The reference year adjustment relies on an apriori estimate of Cant at a given location and depth, which is used as a scaling factor for the concurrent change in atmospheric CO~2~. The underlying assumption is a transient steady state for the oceanic Cant uptake. Here, Cant from the GLODAP mapped Climatology was used.

Note that eq. 6 in Clement and Gruber (2018) misses pCO~2~ pre-industrial in the denominator. Here we use the equation published in Gruber et al. (2019).

### Combine GLODAP + Cant

```{r merge_cstar_cant, eval=FALSE}

S04_cant_3d_obs <- left_join(
  GLODAP_obs_grid,
  S04_cant_3d %>% select(-c(cant, eras))
)

# calculate number of cant data points per grid cell
S04_cant_3d_obs <- S04_cant_3d_obs %>%
  group_by(lon, lat) %>%
  mutate(n = n()) %>%
  ungroup()

# S04_cant_3d_obs %>%
#   filter(n <= 1) %>%
#   ggplot(aes(lon,lat)) +
#   geom_point(data = GLODAP_obs_grid, aes(lon, lat)) +
#   geom_point(col = "red")

rm(S04_cant_3d, GLODAP_obs_grid)

GLODAP_cant_obs <- full_join(GLODAP, S04_cant_3d_obs)

rm(GLODAP, S04_cant_3d_obs)

# fill number of cant data points per grid cell to all observations
GLODAP_cant_obs <- GLODAP_cant_obs %>%
  group_by(lon, lat) %>%
  fill(n, .direction = "updown") %>%
  ungroup()


```

The mapped Cant product was merged with GLODAP observation by:

  - using an identical 1x1° horizontal grid
- linear interpolation of C~ant~ from standard to sampling depth

```{r cant_interpolation_to_observations, eval=FALSE}

# interpolate cant to observation depth
GLODAP_cant_obs_int <- GLODAP_cant_obs %>%
  filter(n > 1) %>%
  group_by(lat, lon) %>%
  arrange(depth) %>%
  mutate(cant_pos_int = approxfun(depth, cant_pos, rule = 2)(depth)) %>%
  ungroup()

# set cant for observation depth if only one cant available
GLODAP_cant_obs_set <- GLODAP_cant_obs %>%
  filter(n == 1) %>%
  group_by(lat, lon) %>%
  mutate(cant_pos_int = mean(cant_pos, na.rm = TRUE)) %>%
  ungroup()

# bin data sets with interpolated and set cant
GLODAP_cant_obs <- bind_rows(GLODAP_cant_obs_int, GLODAP_cant_obs_set)
rm(GLODAP_cant_obs_int, GLODAP_cant_obs_set)


ggplot() +
  geom_path(
    data = GLODAP_cant_obs %>%
      filter(lat == 48.5, lon == 165.5,!is.na(cant_pos)) %>%
      arrange(depth),
    aes(cant_pos, depth, col = "mapped")
  ) +
  geom_point(
    data = GLODAP_cant_obs %>%
      filter(lat == 48.5, lon == 165.5,!is.na(cant_pos)) %>%
      arrange(depth),
    aes(cant_pos, depth, col = "mapped")
  ) +
  geom_point(
    data = GLODAP_cant_obs %>%
      filter(lat == 48.5, lon == 165.5, date == ymd("2018-06-27")),
    aes(cant_pos_int, depth, col = "interpolated")
  ) +
  scale_y_reverse() +
  scale_color_brewer(palette = "Dark2", name = "") +
  labs(title = "Cant interpolation to sampling depth - example profile")

# remove cant data at grid cells without observations
GLODAP <- GLODAP_cant_obs %>%
  filter(!is.na(cstar)) %>%
  mutate(cant_pos = cant_pos_int) %>%
  select(-cant_pos_int, n)

rm(GLODAP_cant_obs)

```

### Merge GLODAP + atm. pCO~2~

GLODAP observations were merged with mean annual atmospheric pCO~2~ levels by year.

```{r merge_cstar_pCO2, eval=FALSE}

GLODAP <- left_join(GLODAP, co2_atm)

```

### Calculation

C~ant~ for median year of each era was calculated by applying alpha = 0.28/13 * (median year - 1994) to the estimate of Sabine et al. (2004).

```{r adjust_reference_year, eval=FALSE}

# assign reference year
GLODAP <- GLODAP %>%
  group_by(era) %>%
  mutate(tref = median(year)) %>%
  ungroup()

# calculate reference year
tref <- GLODAP %>%
  group_by(era) %>%
  summarise(year = median(year)) %>%
  ungroup()

# extract atm pCO2 at reference year
co2_atm_tref <- right_join(co2_atm, tref) %>%
  select(-year) %>%
  rename(pCO2_tref = pCO2)

# merge atm pCO2 at tref with GLODAP
GLODAP <- full_join(GLODAP, co2_atm_tref)
rm(co2_atm, tref)

# scale cant to reference year
GLODAP <- GLODAP %>%
  mutate(alpha = (tref - 1994) * (0.28 / 13),
         cant_pos = cant_pos * (1 + alpha))

# calculate cstar for reference year
GLODAP <- GLODAP %>%
  mutate(
    cstar_tref_delta =
      ((pCO2 - pCO2_tref) / (pCO2_tref - params_local$preind_atm_pCO2)) * cant_pos,
    cstar_tref = cstar - cstar_tref_delta)

```
