---
title: "Analysis of cant estimates"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
params:
  Version_ID: "v_XXX"
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}

if (params$Version_ID == "v_XXX") {
  params_local <-
    read_rds(here::here("data/auxillary", "params_local.rds"))
  
} else {
  params_local <-
    read_rds(
      file = paste(path_root, "observations",
                   params$Version_ID,
                   "data/params_local.rds",
                   sep = "/")
    )
}

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_preprocessing_model    <-
  paste(path_root, "/model/preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}

library(purrr)

```


```{r select_basin_mask, include=FALSE}

basinmask <- basinmask %>% 
  filter(MLR_basins == params_local$MLR_basins) %>% 
  select(-c(MLR_basins, basin))

```


# Version ID

The results displayed on this site correspond to the Version_ID: `r params$Version_ID`

# Data sources

Required are: 

- Cant from Sabine 2004 (S04)
- Cant from Gruber 2019 (G19)
- annual mean atmospheric pCO~2~ 

- Mean and SD per grid cell (lat, lon, depth)


```{r read_cant_files}

cant_3d_tref <- read_csv(paste(path_version_data,
                               "cant_3d_tref.csv",
                               sep = ""))

cant_3d <-
  read_csv(paste(path_version_data,
                 "cant_3d.csv",
                 sep = ""))

co2_atm <-
  read_csv(paste(path_preprocessing,
                 "co2_atm.csv",
                 sep = ""))

tref <-
  read_csv(paste(path_version_data,
                 "tref.csv",
                 sep = ""))

```


```{r compare_cant}


cant_3d_tref_wide <- cant_3d_tref %>%
  pivot_wider(names_from = era,
              values_from = "cant_pos") %>% 
  mutate(cant_pos = `2010-2019` - `2000-2009`) %>% 
  # select(-c(`2010-2019`, `2000-2009`)) %>% 
  drop_na()


cant_3d_tref_wide <- inner_join(basinmask, cant_3d_tref_wide)

zonal_mean_section <- function(df) {

  df_zonal_mean_section <- df %>%
    fselect(-lon) %>% 
    fgroup_by(lat, depth, basin_AIP) %>% {
      add_vars(fgroup_vars(.,"unique"),
               fmean(., keep.group_vars = FALSE) %>% add_stub(pre = FALSE, "_mean"),
               fsd(., keep.group_vars = FALSE) %>% add_stub(pre = FALSE, "_sd"))
    }
  
  return(df_zonal_mean_section)

}

cant_tref_section <- cant_3d_tref_wide %>% 
  group_by(data_source) %>% 
  nest() %>% 
  mutate(section = map(.x = data, ~zonal_mean_section(.x))) %>% 
  select(-data) %>% 
  unnest(section)
  
cant_tref_section %>%
  group_by(basin_AIP, data_source) %>%
  group_split() %>%
  map( ~ p_section_zonal(
    df = .x,
    var = "cant_pos_mean",
    plot_slabs = "n"
  ))

```


```{r estimate_ss_cant}

# extract atm pCO2 at reference year
co2_atm_tref <- right_join(co2_atm, tref %>% rename(year = median_year)) %>% 
  select(-year) %>% 
  rename(pCO2_tref = pCO2)

pCO2_to <- params_local$preind_atm_pCO2
pCO2_t1 <- co2_atm_tref$pCO2_tref[1]
pCO2_t2 <- co2_atm_tref$pCO2_tref[2]

delta_pCO2_ratio <- 
  (pCO2_t2 - pCO2_t1) /
  (pCO2_t1 - pCO2_to)

rm(pCO2_to)
rm(pCO2_t1)
rm(pCO2_t2)

alpha = delta_pCO2_ratio * 0.94 * 0.94

# calculate cstar for reference year
cant_3d_tref_wide <- cant_3d_tref_wide %>%
  rename(cant_t1 = `2010-2019`,
         cant_t2 = `2000-2009`,
         delta_cant_G19 = cant_pos) %>%
  mutate(delta_cant_ss = alpha * cant_t1)

cant_tref_section <- cant_3d_tref_wide %>%
  group_by(data_source) %>%
  nest() %>%
  mutate(section = map(.x = data, ~zonal_mean_section(.x))) %>%
  select(-data) %>%
  unnest(section)

cant_tref_section_long <- cant_tref_section %>%
  pivot_longer(cols = c(delta_cant_G19_mean, delta_cant_ss_mean),
               values_to = "delta_cant",
               names_to = "estimate",
               names_prefix = "delta_cant_")
  
  
cant_tref_section_long %>% 
  group_by(basin_AIP, data_source, estimate) %>%
  group_split() %>% 
  map( ~ p_section_zonal(
    df = .x,
    var = "delta_cant",
    plot_slabs = "n",
    subtitle_text = paste(.x$basin_AIP, .x$data_source, .x$estimate)
  ))


```



## Calculation

To adjust observation-based C* values to the reference year of each observation period, we assume a transient steady state change of cant between the time of sampling the reference year. The adjustment requires an approximation of the cant concentration at the reference year. We approximate this concentration by adding the delta cant signal estimated by Gruber et al (2019) to the "base line" total cant concentration determined for 1994 by Sabine et al (2004):  

Cant(tref) = S04 + (tref-1994)/13 * G19

This way, we use exactly S04+G19 for tref=2007. For all other tref we scale Cant with the observed anomalous change over the 1994-2007 period, rather than assuming a transient steady state. However, one assumes a linear behaviour of the anomalous change over time, which might be wrong in particular for the years past 2007.

For the model data, we perform the adjustment based on the total Cant estimate for 1994, and the delta Cant estimate for 1994 - 2007.


```{r adjust_reference_year, eval=FALSE}


```

