---
title: "Analysis of cant estimates"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}
params_local <-
  read_rds(here::here("data/auxillary",
                       "params_local.rds"))

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(scales)
library(marelac)
library(kableExtra)
library(gt)
```

# Data sources

Cant estimates from this sensitivity case:

- Mean and SD per grid cell (lat, lon, depth)
- Zonal mean and SD (basin, lat, depth)
- Inventories (lat, lon)


```{r read_cant_files}

cant_3d <-
  read_csv(paste(path_version_data,
                 "cant_3d.csv",
                 sep = ""))

cant_zonal <-
  read_csv(paste(path_version_data,
                 "cant_zonal.csv",
                 sep = ""))

cant_predictor_zonal <-
  read_csv(paste(path_version_data,
                 "cant_predictor_zonal.csv",
                 sep = ""))

cant_inv <-
  read_csv(paste(path_version_data,
                 "cant_inv.csv",
                 sep = ""))

```

Target variable (`r params_local$MLR_target`) estimates from this sensitivity case:

- Mean and SD per grid cell (lat, lon, depth)
- Zonal mean and SD (basin, lat, depth)

```{r read_target_files}

target_3d <-
  read_csv(paste(path_version_data,
                 "target_3d.csv",
                 sep = ""))

target_zonal <-
  read_csv(paste(path_version_data,
                 "target_zonal.csv",
                 sep = ""))


```

Cleaned GLODAPv2_2020 file as used in this sensitivity case

```{r read_GLODAP_used_for_fitting}

GLODAP <-
  read_csv(paste(
    path_version_data,
    "GLODAPv2.2020_MLR_fitting_ready.csv",
    sep = ""
  ))

```

# Cant budgets

Global Cant inventories were estimated in units of Pg C. Please note that here we added Cant (all vs postitive only) values and do not apply additional corrections for areas not covered.

```{r calculate_global_inventory}

cant_inv_budget <- cant_inv %>% 
  mutate(surface_area = earth_surf(lat, lon),
         cant_inv_grid = cant_inv*surface_area,
         cant_pos_inv_grid = cant_pos_inv*surface_area) %>% 
  group_by(basin_AIP, eras, inv_depth) %>% 
  summarise(cant_total = sum(cant_inv_grid)*12*1e-15,
            cant_total = round(cant_total,1),
            cant_pos_total = sum(cant_pos_inv_grid)*12*1e-15,
            cant_pos_total = round(cant_pos_total,1)) %>% 
  ungroup()


```

## Standard depth

Results integrated over the upper `r params_global$inventory_depth_standard` m

```{r cant_budget_standard_depth}

cant_inv_budget %>%
  filter(inv_depth == params_global$inventory_depth_standard) %>% 
  gt(rowname_col = "basin_AIP",
     groupname_col = c("eras", "inv_depth"),
     row_group.sep = " | Depth: ") %>% 
  summary_rows(
    groups = TRUE,
    fns = list(total = "sum")
  )


```

## Other depths

Results integrated over the upper `r params_global$inventory_depths` m

```{r cant_budget_other_depth}

cant_inv_budget %>%
  filter(inv_depth != params_global$inventory_depth_standard) %>% 
  gt(rowname_col = "basin_AIP",
     groupname_col = c("eras", "inv_depth"),
     row_group.sep = " | Depth: ") %>% 
  summary_rows(
    groups = TRUE,
    fns = list(total = "sum")
  )

rm(cant_inv_budget)

```

The following analysis is restricted to the standard inventory depth of `r params_global$inventory_depth_standard` m.

```{r filter_standard_inventory_depth}

cant_inv <- cant_inv %>%
  filter(inv_depth == params_global$inventory_depth_standard)

```


# Cant - positive

In a first series of plots we explore the distribution of Cant, taking only positive estimates into account (positive here refers to the mean cant estimate across MLR model predictions available for each grid cell). Negative values were set to zero before calculating mean sections and inventories.

## Zonal mean sections


```{r cant_pos_zonal_mean_sections, fig.asp=0.6}

# i_basin_AIP <- unique(cant_zonal$basin_AIP)[2]
# i_eras <- unique(cant_zonal$eras)[1]

for (i_basin_AIP in unique(cant_zonal$basin_AIP)) {
  for (i_eras in unique(cant_zonal$eras)) {
   
     print(
      p_section_zonal(
        df = cant_zonal %>%
          filter(basin_AIP == i_basin_AIP,
                 eras == i_eras),
        var = "cant_pos_mean",
        subtitle_text =
          paste("Basin:", i_basin_AIP, "| eras:", i_eras))
    )
    
  }
}
  
```

## Isoneutral slab distribution

Mean of positive cant within each horizontal grid cell (lon x lat) per isoneutral slab.

Please note that:

- density slabs covering values >28.1 occur by definition only either in the Atlantic or Indo-Pacific basin
- gaps in the maps represent areas where (thin) density layers fit between discrete depth levels used for mapping

```{r calculate_gamma_slab_maps}

cant_gamma_maps <- m_cant_slab(cant_3d)

cant_gamma_maps <- cant_gamma_maps %>% 
  arrange(gamma_slab, eras)

```

```{r cant_pos_gamma_slab_maps, fig.asp=0.5}

# i_eras <- unique(cant_gamma_maps$eras)[1]
# i_gamma_slab <- unique(cant_gamma_maps$gamma_slab)[1]

for (i_eras in unique(cant_gamma_maps$eras)) {
  for (i_gamma_slab in unique(cant_gamma_maps$gamma_slab)) {
    print(
      p_map_cant_slab(
        df = cant_gamma_maps %>%
          filter(eras == i_eras,
                 gamma_slab == i_gamma_slab),
        subtitle_text = paste(
          "Eras:", i_eras,
          "| Neutral density:", i_gamma_slab)
        )
    )
    
  }
}


```

## Inventory map

Column inventory of positive cant between the surface and `r params_global$inventory_depth`m water depth per horizontal grid cell (lat x lon).

```{r cant_pos_inventory_map, fig.asp=0.5}

# i_eras <- unique(cant_inv$eras)[1]

for (i_eras in unique(cant_inv$eras)) {
  
  print(
    p_map_cant_inv(
      df = cant_inv %>% filter(eras == i_eras),
      var = "cant_pos_inv",
      subtitle_text = paste("Eras:", i_eras))
  )
  
}

```


## Global sections

```{r cant_sections_positive_mean_one_lon_JGOFS_GO, fig.asp=0.6}


for (i_eras in unique(cant_3d$eras)) {
  print(
    p_section_global(
      df = cant_3d %>% filter(eras == i_eras),
      var = "cant_pos",
      subtitle_text = paste("Eras:", i_eras)
    )
  )
  
}

```


# Cant - all

In a second series of plots we explore the distribution of all Cant, taking positive and negative estimates into account.

## Zonal mean sections

```{r cant_all_zonal_mean_sections, fig.asp=0.6}

# i_eras <- unique(cant_zonal$eras)[1]
# i_basin_AIP <- unique(cant_zonal$basin_AIP)[1]

for (i_basin_AIP in unique(cant_zonal$basin_AIP)) {
  for (i_eras in unique(cant_zonal$eras)) {
    print(
      p_section_zonal(
        df = cant_zonal %>%
          filter(basin_AIP == i_basin_AIP,
                 eras == i_eras),
        var = "cant_mean",
        gamma = "gamma_mean",
        breaks = params_global$breaks_cant,
        col = "divergent",
        subtitle_text =
          paste("Basin:", i_basin_AIP, "| eras:", i_eras))
    )
    
  }
}
  
```

## Isoneutral slab distribution

Mean of all Cant within each horizontal grid cell (lon x lat) per isoneutral slab.

Please note that:

- density slabs covering values >28.1 occur by definition only either in the Atlantic or Indo-Pacific basin
- gaps in the maps represent areas where (thin) density layers fit between discrete depth levels used for mapping


```{r cant_all_gamma_slab_maps, fig.asp=0.5}

# i_eras <- unique(cant_gamma_maps$eras)[1]
# i_gamma_slab <- unique(cant_gamma_maps$gamma_slab)[5]

for (i_eras in unique(cant_gamma_maps$eras)) {
  for (i_gamma_slab in unique(cant_gamma_maps$gamma_slab)) {
    print(
      p_map_cant_slab(
        df = cant_gamma_maps %>%
          filter(eras == i_eras,
                 gamma_slab == i_gamma_slab),
        var = "cant",
        col = "divergent",
        subtitle_text = paste(
          "Eras:", i_eras,
          "| Neutral density:", i_gamma_slab))
    )
    
  }
}


```

## Inventory map

Column inventory of all Cant between the surface and `r params_global$inventory_depth_standard`m water depth per horizontal grid cell (lat x lon).

```{r cant_all_inventory_map, fig.asp=0.5}

# i_eras <- unique(cant_inv$eras)[1]

for (i_eras in unique(cant_inv$eras)) {
  
  print(
    p_map_cant_inv(
      df = cant_inv %>% filter(eras == i_eras),
      var = "cant_inv",
      col = "divergent",
      subtitle_text = paste("Eras:", i_eras))
  )
  
}

```

# Cant variability

## Across models

Standard deviation across Cant from all MLR models was calculate for each grid cell (XYZ). The zonal mean of this standard deviation should reflect the uncertainty associated to the predictor selection within each slab and era. 

```{r cant_sections_sd_models, fig.asp=0.6}

# i_eras <- unique(cant_zonal$eras)[1]
# i_basin_AIP <- unique(cant_zonal$basin_AIP)[2]

for (i_basin_AIP in unique(cant_zonal$basin_AIP)) {
  for (i_eras in unique(cant_zonal$eras)) {
    print(
      p_section_zonal(
        df = cant_zonal %>%
          filter(basin_AIP == i_basin_AIP,
                 eras == i_eras),
        var = "cant_sd_mean",
        gamma = "gamma_mean",
        legend_title = "sd",
        title_text = "Zonal mean section of SD across models",
        subtitle_text =
          paste("Basin:", i_basin_AIP, "| eras:", i_eras))
    )
    
  }
}
  

```

## Across basins

Standard deviation of mean cant values was calculate across all longitudes. This standard deviation should reflect the zonal variability of cant within the basin and era.

```{r cant_sections_sd_cant, fig.asp=0.6}

# i_eras <- unique(cant_zonal$eras)[1]
# i_basin_AIP <- unique(cant_zonal$basin_AIP)[2]

for (i_basin_AIP in unique(cant_zonal$basin_AIP)) {
  for (i_eras in unique(cant_zonal$eras)) {
    print(
      p_section_zonal(
        df = cant_zonal %>%
          filter(basin_AIP == i_basin_AIP,
                 eras == i_eras),
        var = "cant_sd",
        gamma = "gamma_mean",
        legend_title = "sd",
        title_text = "Zonal mean section of Cant SD",
        subtitle_text =
          paste("Basin:", i_basin_AIP, "| eras:", i_eras))
    )
    
  }
}

```

## Correlation

### Cant vs model SD

#### Era vs basin

```{r cant_vs_sd_by_basin_era}

cant_3d %>% 
  ggplot(aes(cant, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(basin_AIP ~ eras)

```

#### Basin vs gamma

```{r cant_vs_sd_by_basin_gamma, fig.asp=2}

cant_3d %>% 
  ggplot(aes(cant, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(gamma_slab ~ basin_AIP)

```

### Cant vs regional SD

#### Era vs basin

```{r cant_vs_sd_by_basin_era_zonal}

cant_zonal %>% 
  ggplot(aes(cant_mean, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(basin_AIP ~ eras)

```

#### Era vs basin

```{r cant_vs_sd_by_basin_gamma_zonal, fig.asp=2}

cant_zonal %>% 
  ggplot(aes(cant_mean, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(gamma_slab ~ basin_AIP)

```

# Cant - predictor contribution

```{r cant_section_predictor_contribution, fig.asp=1}

for (i_var in paste("cant",
                    c("intercept", params_local$MLR_predictors),
                    sep = "_")) {

  print(
    p_section_zonal_divergent_gamma_eras_basin(df = cant_predictor_zonal,
                                               var = i_var,
                                               gamma = "gamma")
  )
  
}

rm(i_var)
```


# Neutral density

## Slab depth

The plot below shows the depths of individual gamma slabs (color) together with the observations available in the respective slab.

Please note that:

- density slabs covering values >28.1 occur by definition only either in the Atlantic or Indo-Pacific basin

```{r gamma_maps, fig.asp=2.5}

GLODAP_obs_coverage <- GLODAP %>% 
  count(lat, lon, gamma_slab, era)

map +
  geom_raster(data = cant_gamma_maps,
              aes(lon, lat, fill = depth_max)) +
  geom_raster(data = GLODAP_obs_coverage,
              aes(lon, lat), fill = "red") +
  facet_grid(gamma_slab ~ era) +
  scale_fill_viridis_c(direction = -1) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "top")

rm(GLODAP_obs_coverage)

```

# Target variable

The predicted target variable (`r params_local$MLR_target` in this sensitivity case) is based on fitted MLRs and climatological fields of predictor variables, and calculated for each era.

## Zonal mean sections

```{r target_zonal_mean_era}

slab_breaks <- c(params_local$slabs_Atl[1:12], Inf)

for (i_basin_AIP in unique(target_zonal$basin_AIP)) {

  print(
  target_zonal %>%
    filter(basin_AIP == i_basin_AIP) %>%
    ggplot(aes(lat, depth,
               z = !!sym(
                 paste(params_local$MLR_target, "mean", sep = "_")
               ))) +
    geom_contour_filled(bins = 11) +
    scale_fill_viridis_d(name = params_local$MLR_target) +
    geom_contour(aes(lat, depth, z = gamma_mean),
                 breaks = slab_breaks,
                 col = "white") +
    geom_text_contour(
      aes(lat, depth, z = gamma_mean),
      breaks = slab_breaks,
      col = "white",
      skip = 1
    ) +
    scale_y_reverse() +
    coord_cartesian(expand = 0,
                    ylim = c(params_global$plotting_depth, 0)) +
    scale_x_continuous(breaks = seq(-100, 100, 20)) +
    guides(fill = guide_colorsteps(barheight = unit(10, "cm"))) +
    facet_grid(era ~ eras,
               labeller = labeller(.default = label_both)) +
    labs(title = i_basin_AIP)
  )

}

rm(slab_breaks)

```
