---
title: "Analysis of cant estimates"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
params:
  Version_ID: "v_XXX"
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}

if (params$Version_ID == "v_XXX") {
  params_local <-
    read_rds(here::here("data/auxillary", "params_local.rds"))
  
} else {
  params_local <-
    read_rds(
      file = paste(path_root, "observations",
                   params$Version_ID,
                   "data/params_local.rds",
                   sep = "/")
    )
}

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(scales)
library(marelac)
library(kableExtra)
library(gt)
```

# Version ID

The results displayed on this site correspond to the Version_ID: `r params$Version_ID`

# Data sources

Cant estimates from this sensitivity case:

- Mean and SD per grid cell (lat, lon, depth)
- Zonal mean and SD (basin, lat, depth)
- Inventories (lat, lon)


```{r read_cant_files}

cant_3d <-
  read_csv(paste(path_version_data,
                 "cant_3d.csv",
                 sep = ""))

cant_3d_mod_truth <-
  read_csv(paste(path_version_data,
                 "cant_3d_mod_truth.csv",
                 sep = ""))

cant_3d <- bind_rows(cant_3d,
                     cant_3d_mod_truth)


```

Cleaned GLODAPv2_2020 file as used in this sensitivity case

```{r read_GLODAP_used_for_fitting}

GLODAP <-
  read_csv(paste(
    path_version_data,
    "GLODAPv2.2020_MLR_fitting_ready.csv",
    sep = ""
  ))

GLODAP_grid <- GLODAP %>% 
  distinct(lon, lat, gamma_slab)

```


```{r calculate_gamma_slab_maps}

cant_gamma_maps <- m_cant_slab_data_source(cant_3d)

cant_gamma_maps <- cant_gamma_maps %>% 
  arrange(gamma_slab, data_source)

```



# Cant - positive

In a first series of plots we explore the distribution of Cant, taking only positive estimates into account (positive here refers to the mean cant estimate across MLR model predictions available for each grid cell). Negative values were set to zero before calculating mean sections and inventories.

Mean of positive cant within each horizontal grid cell (lon x lat) per isoneutral slab.

Please note that:

- density slabs covering values >28.1 occur by definition only either in the Atlantic or Indo-Pacific basin
- gaps in the maps represent areas where (thin) density layers fit between discrete depth levels used for mapping


```{r cant_pos_gamma_slab_maps, fig.asp=0.5}

cant_gamma_maps %>%
  group_split(gamma_slab, data_source) %>%
  # head(1) %>%
  map(
    ~ p_map_cant_slab(
      df = .x,
      subtitle_text = paste(
        "data_source:",
        unique(.x$data_source),
        "| Neutral density:",
        unique(.x$gamma_slab)
      )
    ) +
      geom_raster(data = GLODAP_grid %>% filter(gamma_slab == unique(.x$gamma_slab)),
                  aes(lon, lat),
                  fill="grey70")
  )


```

## Bias

```{r cant_pos_gamma_slab_maps_bias, fig.asp=0.5}

cant_pos_slab_inv_bias <- cant_gamma_maps %>% 
  select(lat, lon, gamma_slab, cant_pos, data_source) %>% 
  pivot_wider(names_from = data_source,
              values_from = cant_pos) %>% 
  mutate(cant_pos_bias = mod - mod_truth)

cant_pos_slab_inv_bias %>%
  group_split(gamma_slab) %>%
  # head(1) %>%
  map(
    ~ p_map_cant_slab(
      df = .x,
      var = "cant_pos_bias",
      col = "divergent",
      subtitle_text = paste("| Neutral density:", unique(.x$gamma_slab))
    ) +
      geom_raster(
        data = GLODAP_grid %>% filter(gamma_slab == unique(.x$gamma_slab)),
        aes(lon, lat),
        fill = "grey30"
      )
  )


```

```{r test_inventory_map, eval=FALSE}

i_gamma_slab <- unique(cant_3d$gamma_slab)[2]
i_data_source <- unique(cant_3d$data_source)[2]

map +
  geom_contour_filled(
    data = cant_gamma_maps %>%
      filter(data_source == i_data_source,
             gamma_slab == i_gamma_slab),
    aes(lon, lat, z = cant)
  ) +
  scale_fill_viridis_d()

```


# Cant - all

In a second series of plots we explore the distribution of all Cant, taking positive and negative estimates into account.

Mean of all Cant within each horizontal grid cell (lon x lat) per isoneutral slab.

Please note that:

- density slabs covering values >28.1 occur by definition only either in the Atlantic or Indo-Pacific basin
- gaps in the maps represent areas where (thin) density layers fit between discrete depth levels used for mapping


```{r cant_all_gamma_slab_maps, fig.asp=0.5}

cant_gamma_maps %>%
  group_split(gamma_slab, data_source) %>%
  map(
    ~ p_map_cant_slab(
      df = .x,
      var = "cant",
      col = "divergent",
      subtitle_text = paste(
        "data_source:",
        unique(.x$data_source),
        "| Neutral density:",
        unique(.x$gamma_slab)
      )
    ) +
      geom_raster(
        data = GLODAP_grid %>% filter(gamma_slab == unique(.x$gamma_slab)),
        aes(lon, lat),
        fill = "grey70"
      )
  )



```

## Bias

```{r cant_gamma_slab_maps_bias, fig.asp=0.5}

cant_slab_inv_bias <- cant_gamma_maps %>% 
  select(lat, lon, gamma_slab, cant, data_source) %>% 
  pivot_wider(names_from = data_source,
              values_from = cant) %>% 
  mutate(cant_bias = mod - mod_truth)


cant_slab_inv_bias %>%
  group_split(gamma_slab) %>%
  # head(1) %>%
  map(
    ~ p_map_cant_slab(
      df = .x,
      var = "cant_bias",
      col = "divergent",
      subtitle_text = paste("| Neutral density:", unique(.x$gamma_slab))
    ) +
      geom_raster(
        data = GLODAP_grid %>% filter(gamma_slab == unique(.x$gamma_slab)),
        aes(lon, lat),
        fill = "grey30"
      )
  )


```


# Slab inventories

```{r calc_slab_inventory}

cant_slab_budget <- m_cant_slab_inv_data_source(cant_3d)

cant_slab_budget %>% 
  ggplot(aes(data_source, cant_pos_total, fill = gamma_slab)) +
  geom_col() +
  scale_fill_viridis_d(direction = -1) +
  facet_wrap(~ basin_AIP)

cant_slab_budget %>% 
  ggplot(aes(data_source, cant_total, fill = gamma_slab)) +
  geom_hline(yintercept = 0, col = "red") +
  geom_col() +
  scale_fill_viridis_d(direction = -1) +
  facet_wrap(~ basin_AIP)


```

```{r cant_pos_slab_inv_bias, fig.asp=1}

cant_slab_budget_bias <- cant_slab_budget %>%
  filter(data_source %in% c("mod", "mod_truth")) %>%
  select(basin_AIP, data_source, gamma_slab, cant_pos_total) %>%
  pivot_wider(names_from = data_source,
              values_from = cant_pos_total) %>%
  mutate(cant_pos_bias = mod - mod_truth)

cant_slab_budget_bias %>% 
  ggplot(aes(gamma_slab, cant_pos_bias)) +
  geom_col() +
  facet_grid(basin_AIP ~ ., scales = "free_y") +
  coord_flip()

```

```{r cant_slab_budget_bias, fig.asp=1}

cant_slab_budget_bias <- cant_slab_budget %>%
  filter(data_source %in% c("mod", "mod_truth")) %>%
  select(basin_AIP, data_source, gamma_slab, cant_total) %>%
  pivot_wider(names_from = data_source,
              values_from = cant_total) %>%
  mutate(cant_bias = mod - mod_truth)

cant_slab_budget_bias %>% 
  ggplot(aes(gamma_slab, cant_bias)) +
  geom_col() +
  facet_grid(basin_AIP ~ ., scales = "free_y") +
  coord_flip()

```

# Write files

```{r write_csv}

cant_slab_budget_bias %>%
  write_csv(paste0(path_version_data,
                   "cant_slab_budget_bias.csv"))

cant_slab_budget %>%
  write_csv(paste0(path_version_data,
                   "cant_slab_budget.csv"))

cant_slab_inv_bias %>%
  write_csv(paste0(path_version_data,
                   "cant_slab_inv_bias.csv"))

cant_pos_slab_inv_bias %>%
  write_csv(paste0(path_version_data,
                   "cant_slab_inv_bias_pos.csv"))

```


# Slab boundary depth

```{r gamma_slab_boundary_depth, eval=FALSE, fig.asp=2.5}

# gamma field in Atl
predictors_gamma_Atl <- cant_3d %>%
  select(data_source, basin_AIP, lat, lon, depth, gamma) %>% 
  filter(basin_AIP == "Atlantic") %>% 
  distinct()

# filter grid cells with at least two depth layers
predictors_gamma_Atl <- predictors_gamma_Atl %>%
  group_by(data_source, lat, lon) %>%
  mutate(n = n()) %>%
  ungroup() %>% 
  filter(n > 1) %>% 
  select(-n)

# 3D grid with Atl slab boundaries
slabs_Atl <- expand_grid(predictors_gamma_Atl %>% distinct(data_source, lat, lon),
                         gamma = params_local$slabs_Atl) %>% 
  mutate(basin_AIP = "flag")

# join slab boundary grid with gamma field
slabs_Atl_int <- full_join(slabs_Atl,
                           predictors_gamma_Atl) %>%
  arrange(lat, lon, gamma)

# interpolate depth to slab boundary
slabs_Atl_int <- slabs_Atl_int %>% 
  group_by(data_source, lat, lon) %>% 
  mutate(depth = approxfun(gamma, depth, rule = 1)(gamma)) %>% 
  ungroup()

# filter interpolated depths
slabs_Atl_int <- slabs_Atl_int %>% 
  filter(basin_AIP == "flag",
         !is.na(depth))

#### Indo-Pacific

# gamma field in Pac
predictors_gamma_Ind_Pac <- cant_3d %>%
  select(data_source, basin_AIP, lat, lon, depth, gamma) %>% 
  filter(basin_AIP %in% c("Indian", "Pacific")) %>% 
  distinct()

# filter grid cells with at least two depth layers
predictors_gamma_Ind_Pac <- predictors_gamma_Ind_Pac %>%
  group_by(data_source, lat, lon) %>%
  mutate(n = n()) %>%
  ungroup() %>% 
  filter(n > 1) %>% 
  select(-n)

# 3D grid with Pac slab boundaries
slabs_Ind_Pac <- expand_grid(predictors_gamma_Ind_Pac %>% distinct(data_source, lat, lon),
                         gamma = params_local$slabs_Ind_Pac) %>% 
  mutate(basin_AIP = "flag")

# join slab boundary grid with gamma field
slabs_Ind_Pac_int <- full_join(slabs_Ind_Pac,
                           predictors_gamma_Ind_Pac) %>%
  arrange(lat, lon, gamma)

# interpolate depth to slab boundary
slabs_Ind_Pac_int <- slabs_Ind_Pac_int %>% 
  group_by(data_source, lat, lon) %>% 
  mutate(depth = approxfun(gamma, depth, rule = 1)(gamma)) %>% 
  ungroup()

# filter interpolated depths
slabs_Ind_Pac_int <- slabs_Ind_Pac_int %>% 
  filter(basin_AIP == "flag",
         !is.na(depth))


# join interpolated gamma depth
slabs <- bind_rows(slabs_Atl_int,
                   slabs_Ind_Pac_int) %>%
  select(-basin_AIP)

slabs <- inner_join(
  basinmask %>% select(lat, lon, basin_AIP) %>% distinct(),
  slabs)


map + 
  geom_raster(data = slabs,
              aes(lon, lat, fill = depth)) +
  facet_grid(gamma ~ data_source) +
  scale_fill_viridis_c(direction = -1)

rm(slabs_Atl, slabs_Atl_int, slabs_Ind_Pac, slabs_Ind_Pac_int)

```

```{r calculate_slab_thickness, eval=FALSE, fig.asp=2.5}
 
# slabs_ex <- slabs %>% 
#   filter(lat == 0.5,
#          lon == 250.5)

slabs <- slabs %>% 
  group_by(data_source, lat, lon) %>% 
  arrange(depth) %>% 
  mutate(depth_lower_bound = depth,
         depth_upper_bound = lag(depth, default = 0),
         slab_thickness = depth_lower_bound - depth_upper_bound,
         gamma_in_slab = gamma - 0.01) %>% 
  ungroup()

slabs <- m_cut_gamma(slabs, "gamma_in_slab")

map + 
  geom_raster(data = slabs,
              aes(lon, lat, fill = slab_thickness)) +
  facet_grid(gamma_slab ~ data_source) +
  scale_fill_viridis_c(direction = -1)

rm(slabs)

```
