---
title: "Analysis of dcant estimates"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
params:
  Version_ID: "v_XXX"
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup_obs.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}

if (params$Version_ID == "v_XXX") {
  params_local <-
    read_rds(here::here("data/auxillary", "params_local.rds"))
  
} else {
  params_local <-
    read_rds(
      file = paste(path_root, "observations",
                   params$Version_ID,
                   "data/params_local.rds",
                   sep = "/")
    )
}

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(scales)
library(marelac)
library(kableExtra)
library(gt)
```

# Version ID

The results displayed on this site correspond to the Version_ID: `r params$Version_ID`

# Data sources

dcant estimates from this sensitivity case:

- Mean and SD per grid cell (lat, lon, depth)
- Zonal mean and SD (basin, lat, depth)
- Inventories (lat, lon)


```{r read_dcant_files}

dcant_3d <-
  read_csv(paste(path_version_data,
                 "dcant_3d.csv",
                 sep = ""))

dcant_3d_mod_truth <-
  read_csv(paste(path_version_data,
                 "dcant_3d_mod_truth.csv",
                 sep = ""))

dcant_zonal <-
  read_csv(paste(path_version_data,
                 "dcant_zonal.csv",
                 sep = ""))

dcant_zonal_mod_truth <-
  read_csv(paste(path_version_data,
                 "dcant_zonal_mod_truth.csv",
                 sep = ""))

dcant_zonal <- bind_rows(dcant_zonal,
                        dcant_zonal_mod_truth)

dcant_inv <-
  read_csv(paste(path_version_data,
                 "dcant_inv.csv",
                 sep = ""))

```

# Zonal sections

## SD across models

Standard deviation across dcant from all MLR models was calculate for each grid cell (XYZ). The zonal mean of this standard deviation should reflect the uncertainty associated to the predictor selection within each slab and era. 

```{r dcant_sections_sd_models, fig.asp=0.6}

dcant_zonal %>%
  group_by(basin_AIP, data_source) %>%
  group_split() %>%
  head(1) %>%
  map(
    ~ p_section_zonal(
      df = .x,
      var = "dcant_sd_mean",
      gamma = "gamma_mean",
      legend_title = "sd",
      breaks = c(seq(0, 3, 0.3), Inf),
      title_text = "Zonal mean section of SD across models",
      subtitle_text =
        paste(
          "Basin:",
          unique(.x$basin_AIP),
          "| data_source:",
          unique(.x$data_source)
        )
    )
  )

```

## SD across basins

Standard deviation of mean dcant values was calculate across all longitudes. This standard deviation should reflect the zonal variability of dcant within the basin and era.

```{r dcant_sections_sd_dcant, fig.asp=0.6}

dcant_zonal %>%
  group_by(basin_AIP, data_source) %>%
  group_split() %>%
  head(1) %>% 
  map(
    ~ p_section_zonal(
      df = .x,
      var = "dcant_sd",
      gamma = "gamma_mean",
      legend_title = "sd",
      breaks = c(seq(0, 3, 0.3),Inf),
      title_text = "Zonal mean section of dcant SD",
      subtitle_text =
        paste(
          "Basin:",
          unique(.x$basin_AIP),
          "| data_source:",
          unique(.x$data_source)
        )
    )
  )


```

# Slab

## SD across models

```{r SD_slabs}



```



# Correlation

### dcant vs model SD

#### Era vs basin

```{r dcant_vs_sd_by_basin_era}

dcant_3d %>% 
  ggplot(aes(dcant, dcant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(basin_AIP ~ data_source)

```

#### Basin vs gamma

```{r dcant_vs_sd_by_basin_gamma, fig.asp=2}

dcant_3d %>%
  group_split(basin_AIP) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(dcant, dcant_sd)) +
      geom_vline(xintercept = 0) +
      geom_hline(yintercept = 10) +
      geom_bin2d() +
      scale_fill_viridis_c(
        option = "magma",
        
        trans = "log10",
        name = "log10(n)"
      ) +
      facet_grid(gamma_slab ~ data_source) +
      labs(title = paste("Basin:", unique(.x$basin_AIP)))
  )

```

### dcant vs regional SD

#### Era vs basin

```{r dcant_vs_sd_by_basin_era_zonal}

dcant_zonal %>% 
  ggplot(aes(dcant_mean, dcant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(basin_AIP ~ data_source)

```

#### Era vs basin

```{r dcant_vs_sd_by_basin_gamma_zonal, fig.asp=2}

dcant_zonal %>%
  group_split(basin_AIP) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(dcant_mean, dcant_sd)) +
      geom_vline(xintercept = 0) +
      geom_hline(yintercept = 10) +
      scale_fill_viridis_c(
        option = "magma",
        
        trans = "log10"
      ) +
      geom_bin2d() +
      facet_grid(gamma_slab ~ data_source)  +
      labs(title = paste("Basin:", unique(.x$basin_AIP)))
  )

```

### dcant bias vs model SD

```{r dcant_bias_vs_dcant_sd}

dcant_3d <- bind_rows(dcant_3d,
                     dcant_3d_mod_truth)


dcant_3d_bias <- dcant_3d %>%
  filter(data_source %in% c("mod", "mod_truth")) %>%
  select(lat, lon, depth, data_source, basin_AIP, gamma_slab, dcant) %>%
  pivot_wider(names_from = data_source,
              values_from = dcant) %>%
  mutate(dcant_bias = mod - mod_truth)

dcant_3d_bias_sd <- dcant_3d %>% 
  filter(data_source %in% c("mod")) %>%
  select(lat, lon, depth, dcant, dcant_sd)

dcant_3d_bias <- full_join(dcant_3d_bias, dcant_3d_bias_sd)

dcant_3d_bias %>%
  ggplot(aes(dcant_sd, dcant_bias)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  scale_fill_viridis_c(option = "magma",
                       
                       trans = "log10") +
  geom_bin2d()


```


```{r dcant_bias_vs_dcant_sd_basin_gamma_slab, fig.asp=2}

dcant_3d_bias %>%
  ggplot(aes(dcant_sd, dcant_bias)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  scale_fill_viridis_c(option = "magma",
                       
                       trans = "log10") +
  geom_bin2d() +
  facet_grid(gamma_slab ~ basin_AIP)

```


### dcant bias vs dcant


```{r dcant_bias_vs_dcant}

dcant_3d_bias %>%
  ggplot(aes(dcant, dcant_bias)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  scale_fill_viridis_c(option = "magma",
                       
                       trans = "log10") +
  geom_bin2d()
```



```{r dcant_bias_vs_dcant_basin_gamma_slab, fig.asp=2}


dcant_3d_bias %>%
  ggplot(aes(dcant, dcant_bias)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  scale_fill_viridis_c(option = "magma",
                       
                       trans = "log10") +
  geom_bin2d() +
  facet_grid(gamma_slab ~ basin_AIP)

```



