---
title: "Analysis of cant estimates"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}
params_local <-
  read_rds(here::here("data/auxillary",
                       "params_local.rds"))

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(scales)
library(marelac)
library(kableExtra)
library(gt)
```

# Data sources

Cant estimates from this sensitivity case:

- Mean and SD per grid cell (lat, lon, depth)
- Zonal mean and SD (basin, lat, depth)
- Inventories (lat, lon)


```{r read_cant_files}

cant_3d <-
  read_csv(paste(path_version_data,
                 "cant_3d.csv",
                 sep = ""))

cant_zonal <-
  read_csv(paste(path_version_data,
                 "cant_zonal.csv",
                 sep = ""))

cant_zonal_mod_truth <-
  read_csv(paste(path_version_data,
                 "cant_zonal_mod_truth.csv",
                 sep = ""))

cant_zonal <- bind_rows(cant_zonal,
                        cant_zonal_mod_truth)

cant_inv <-
  read_csv(paste(path_version_data,
                 "cant_inv.csv",
                 sep = ""))

```

# Cant variability

## Across models

Standard deviation across Cant from all MLR models was calculate for each grid cell (XYZ). The zonal mean of this standard deviation should reflect the uncertainty associated to the predictor selection within each slab and era. 

```{r cant_sections_sd_models, fig.asp=0.6}

# i_data_source <- unique(cant_zonal$data_source)[1]
# i_basin_AIP <- unique(cant_zonal$basin_AIP)[2]

for (i_basin_AIP in unique(cant_zonal$basin_AIP)) {
  for (i_data_source in unique(cant_zonal$data_source)) {
    print(
      p_section_zonal(
        df = cant_zonal %>%
          filter(basin_AIP == i_basin_AIP,
                 data_source == i_data_source),
        var = "cant_sd_mean",
        gamma = "gamma_mean",
        legend_title = "sd",
        title_text = "Zonal mean section of SD across models",
        subtitle_text =
          paste("Basin:", i_basin_AIP, "| data_source:", i_data_source))
    )
    
  }
}
  

```

## Across basins

Standard deviation of mean cant values was calculate across all longitudes. This standard deviation should reflect the zonal variability of cant within the basin and era.

```{r cant_sections_sd_cant, fig.asp=0.6}

# i_data_source <- unique(cant_zonal$data_source)[1]
# i_basin_AIP <- unique(cant_zonal$basin_AIP)[2]

for (i_basin_AIP in unique(cant_zonal$basin_AIP)) {
  for (i_data_source in unique(cant_zonal$data_source)) {
    print(
      p_section_zonal(
        df = cant_zonal %>%
          filter(basin_AIP == i_basin_AIP,
                 data_source == i_data_source),
        var = "cant_sd",
        gamma = "gamma_mean",
        legend_title = "sd",
        title_text = "Zonal mean section of Cant SD",
        subtitle_text =
          paste("Basin:", i_basin_AIP, "| data_source:", i_data_source))
    )
    
  }
}

```

## Correlation

### Cant vs model SD

#### Era vs basin

```{r cant_vs_sd_by_basin_era}

cant_3d %>% 
  ggplot(aes(cant, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(basin_AIP ~ data_source)

```

#### Basin vs gamma

```{r cant_vs_sd_by_basin_gamma, fig.asp=2}

cant_3d %>%
  group_split(basin_AIP) %>%
  map(
    ~ ggplot(data = .x, aes(cant, cant_sd)) +
      geom_vline(xintercept = 0) +
      geom_hline(yintercept = 10) +
      geom_bin2d() +
      scale_fill_viridis_c(
        option = "magma",
        direction = -1,
        trans = "log10",
        name = "log10(n)"
      ) +
      facet_grid(gamma_slab ~ data_source) +
      labs(title = paste("Basin:", unique(.x$basin_AIP)))
  )

```

### Cant vs regional SD

#### Era vs basin

```{r cant_vs_sd_by_basin_era_zonal}

cant_zonal %>% 
  ggplot(aes(cant_mean, cant_sd)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 10) +
  geom_bin2d() +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10",
                       name = "log10(n)") +
  facet_grid(basin_AIP ~ data_source)

```

#### Era vs basin

```{r cant_vs_sd_by_basin_gamma_zonal, fig.asp=2}

cant_zonal %>%
  group_split(basin_AIP) %>%
  map(
    ~ ggplot(data = .x, aes(cant_mean, cant_sd)) +
      geom_vline(xintercept = 0) +
      geom_hline(yintercept = 10) +
      geom_bin2d() +
      scale_fill_viridis_c(
        option = "magma",
        direction = -1,
        trans = "log10",
        name = "log10(n)"
      ) +
      facet_grid(gamma_slab ~ data_source)  +
      labs(title = paste("Basin:", unique(.x$basin_AIP)))
  )

```
