---
title: "Dependencies"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

```{r library, message=FALSE, warning=FALSE}
library(reticulate)
```


# Required packages

Packages available on CRAN are not listed here.

Following packages need to be installed but are not available on CRAN:

- [cutr](https://github.com/moodymudskipper/cutr/){target="_blank"}

Run the code below for installation from Github repositories.

```{r install_r_packages_not_available_on_CRAN, eval=FALSE}

remotes::install_github("moodymudskipper/cutr")

```

# Python

## Code

Python code used in this project is stored in /code/python_scripts and sourced from there, when required in .Rmd files.

## Required packages

To install required python packages from within R-Studio, run the code below.

```{r install_required_python_packages_from_r, eval=FALSE}

# the commands below must only be run once, when you first run the python code and need to install required packages

py_install("pandas")
py_install("numpy")
py_install("shapely")
py_install("math")

```
