---
title: "eMLR - data preparation"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/updata/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}
params_local <-
  read_rds(here::here("data/auxillary",
                       "params_local.rds"))

```

```{r define_updata_paths, include = FALSE}
path_preprocessing    <-
  "/nfs/kryo/work/updata/emlr_cant/observations/preprocessing/"

path_version_data     <-
  paste(
    "/nfs/kryo/work/updata/emlr_cant/observations/",
    params_local$Version_ID,
    "/data/",
    sep = ""
  )

path_version_figures  <-
  paste(
    "/nfs/kryo/work/updata/emlr_cant/observations/",
    params_local$Version_ID,
    "/figures/",
    sep = ""
  )
```

```{r load_libraries_specific, include = FALSE}
library(marelac)
library(lubridate)
```

# Required data

Required are: 

- GLODAPv2.2020
  - cleaned data file
  - horizontal grid of sampling coordinates
- Cant from Sabine 2004
- annual mean atmospheric pCO~2~ 

```{r read_required_data_sets}

GLODAP <-
  read_csv(paste(path_version_data,
                 "GLODAPv2.2020_clean.csv",
                 sep = ""))

GLODAP_obs_grid <-
  read_csv(paste(path_version_data,
                 "GLODAPv2.2020_clean_obs_grid.csv",
                 sep = ""))

S04_cant_3d <-
  read_csv(paste(path_preprocessing,
                 "S04_cant_3d.csv",
                 sep = ""))

co2_atm <-
  read_csv(paste(path_preprocessing,
                 "co2_atm.csv",
                 sep = ""))


```


# PO~4~*

## Calculation 

The predictor PO~4~* was be calculated according to Clement and Gruber (2018), ie based on oxygen. Please note that an erroneous equations for PO4* calculation is given in the supplement of Gruber et al (2019), based on nitrate.

```{r calculate_phosphate_star}

GLODAP <- GLODAP %>% 
  mutate(phosphate_star = b_phosphate_star(phosphate, oxygen))

```



# C*

C* serves as a conservative tracer of anthropogenic CO~2~ uptake. It is derived from measured DIC by removing the impact of  

- organic matter formation and respiration
- calcification and calcium carbonate dissolution

Contributions of those processes are estimated from phosphate and alkalinity concentrations.

## Stoichiometric ratios

The stoichiometric nutrient ratios for the production and mineralization of organic matter were set to:

- C/P: `r params_local$rCP`
- N/P: `r params_local$rNP`

## Calculation

C* was calculated as:  

```{r calculate_cstar}

print(b_cstar)

GLODAP <- GLODAP %>% 
  mutate(rCP_phosphate = -params_local$rCP * phosphate,
         talk_05 = -0.5 * talk,
         rNP_phosphate_05 = -0.5 * params_local$rNP * phosphate,
         cstar = b_cstar(tco2, phosphate, talk))

```

## Reference year adjustment

The reference year adjustment relies on an apriori estimate of Cant at a given location and depth, which is used as a scaling factor for the concurrent change in atmospheric CO~2~. The underlying assumption is a transient steady state for the oceanic Cant uptake. Here, Cant from the GLODAP mapped Climatology was used.  

Note that eq. 6 in Clement and Gruber (2018) misses pCO~2~ pre-industrial in the denominator. Here we use the equation published in Gruber et al. (2019).

### Combine GLODAP + Cant


```{r merge_cstar_cant}

S04_cant_3d_obs <- left_join(
  GLODAP_obs_grid,
  S04_cant_3d %>% select(-c(cant, eras))
  )

# calculate number of cant data points per grid cell
S04_cant_3d_obs <- S04_cant_3d_obs %>%
  group_by(lon, lat) %>% 
  mutate(n = n()) %>% 
  ungroup()

# S04_cant_3d_obs %>%
#   filter(n <= 1) %>%
#   ggplot(aes(lon,lat)) +
#   geom_point(data = GLODAP_obs_grid, aes(lon, lat)) +
#   geom_point(col = "red")

rm(S04_cant_3d, GLODAP_obs_grid)

GLODAP_cant_obs <- full_join(GLODAP, S04_cant_3d_obs)

rm(GLODAP, S04_cant_3d_obs)

# fill number of cant data points per grid cell to all observations
GLODAP_cant_obs <- GLODAP_cant_obs %>%
  group_by(lon, lat) %>% 
  fill(n, .direction = "updown") %>% 
  ungroup()


```

The mapped Cant product was merged with GLODAP observation by:

- using an identical 1x1° horizontal grid
- linear interpolation of C~ant~ from standard to sampling depth

```{r cant_interpolation_to_observations}

# interpolate cant to observation depth
GLODAP_cant_obs_int <- GLODAP_cant_obs %>%
  filter(n > 1) %>% 
  group_by(lat, lon) %>%
  arrange(depth) %>%
  mutate(cant_pos_int = approxfun(depth, cant_pos, rule = 2)(depth)) %>%
  ungroup()

# set cant for observation depth if only one cant available
GLODAP_cant_obs_set <- GLODAP_cant_obs %>%
  filter(n == 1) %>%
  group_by(lat, lon) %>%
  mutate(cant_pos_int = mean(cant_pos, na.rm = TRUE)) %>%
  ungroup()

# bin data sets with interpolated and set cant
GLODAP_cant_obs <- bind_rows(GLODAP_cant_obs_int, GLODAP_cant_obs_set)
rm(GLODAP_cant_obs_int, GLODAP_cant_obs_set)


ggplot() +
  geom_path(
    data = GLODAP_cant_obs %>%
      filter(lat == 48.5, lon == 165.5,!is.na(cant_pos)) %>%
      arrange(depth),
    aes(cant_pos, depth, col = "mapped")
  ) +
  geom_point(
    data = GLODAP_cant_obs %>%
      filter(lat == 48.5, lon == 165.5,!is.na(cant_pos)) %>%
      arrange(depth),
    aes(cant_pos, depth, col = "mapped")
  ) +
  geom_point(
    data = GLODAP_cant_obs %>%
      filter(lat == 48.5, lon == 165.5, date == ymd("2018-06-27")),
    aes(cant_pos_int, depth, col = "interpolated")
  ) +
  scale_y_reverse() +
  scale_color_brewer(palette = "Dark2", name = "") +
  labs(title = "Cant interpolation to sampling depth - example profile")

# remove cant data at grid cells without observations
GLODAP <- GLODAP_cant_obs %>%
  filter(!is.na(cstar)) %>%
  mutate(cant_pos = cant_pos_int) %>%
  select(-cant_pos_int, n)

rm(GLODAP_cant_obs)

```

### Merge GLODAP + atm. pCO~2~

GLODAP observations were merged with mean annual atmospheric pCO~2~ levels by year.

```{r merge_cstar_pCO2}

GLODAP <- left_join(GLODAP, co2_atm)

```

### Calculation

C~ant~ for median year of each era was calculated by applying alpha = 0.28/13 * (median year - 1994) to the estimate of Sabine et al. (2004).

```{r adjust_reference_year}

# assign reference year
GLODAP <- GLODAP %>% 
  group_by(era) %>% 
  mutate(tref = median(year)) %>% 
  ungroup()

# calculate reference year
tref <- GLODAP %>% 
  group_by(era) %>% 
  summarise(year = median(year)) %>% 
  ungroup()

# extract atm pCO2 at reference year
co2_atm_tref <- right_join(co2_atm, tref) %>% 
  select(-year) %>% 
  rename(pCO2_tref = pCO2)

# merge atm pCO2 at tref with GLODAP
GLODAP <- full_join(GLODAP, co2_atm_tref)
rm(co2_atm, tref)

# scale cant to reference year
GLODAP <- GLODAP %>%
  mutate(alpha = (tref - 1994) * (0.28 / 13),
         cant_pos = cant_pos * (1 + alpha))

# calculate cstar for reference year
GLODAP <- GLODAP %>%
  mutate(
    cstar_tref_delta =
      ((pCO2 - pCO2_tref) / (pCO2_tref - params_local$preind_atm_pCO2)) * cant_pos,
    cstar_tref = cstar - cstar_tref_delta)

```

## Control plots

```{r cstar_reference_year_adjustment_histogram}

GLODAP %>% 
  ggplot(aes(cstar_tref_delta)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Histogramm with binwidth = 1")

```

```{r cstar_reference_year_adjustment_vs_time_cant}

GLODAP %>% 
  sample_n(1e4) %>% 
  ggplot(aes(year, cstar_tref_delta, col = cant_pos)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(title = "Time series of random subsample 1e4")


```

```{r cstar_reference_year_adjustment_vs_time_counts}

GLODAP %>% 
  ggplot(aes(year, cstar_tref_delta)) +
  geom_bin2d(binwidth = 1) +
  scale_fill_viridis_c(trans = "log10") +
  labs(title = "Heatmap with binwidth = 1")

```

# Selected section plots

A selected section is plotted to demonstrate the magnitude of various parameters and corrections relevant to C*.

```{r select_cruises}

GLODAP_cruise <- GLODAP %>% 
  filter(cruise %in% params_global$cruises_meridional)

```


```{r meridional_sections, fig.asp=0.5}

map +
  geom_path(data = GLODAP_cruise %>%
              arrange(date),
            aes(lon, lat)) +
  geom_point(data = GLODAP_cruise %>%
              arrange(date),
             aes(lon, lat, col = date)) +
  scale_color_viridis_c(trans = "date") +
  labs(title = paste("Cruise year:", mean(GLODAP_cruise$year)))

lat_section <- 
GLODAP_cruise %>%
  ggplot(aes(lat, depth)) +
  scale_y_reverse() +
  scale_fill_viridis_c() +
  theme(axis.title.x = element_blank())

lat_section +
  stat_summary_2d(aes(z = tco2))

lat_section +
  stat_summary_2d(aes(z = rCP_phosphate))

lat_section +
  stat_summary_2d(aes(z = talk_05))

lat_section +
  stat_summary_2d(aes(z = rNP_phosphate_05))

lat_section +
  stat_summary_2d(aes(z = cstar))

lat_section +
  stat_summary_2d(aes(z = cstar_tref))

rm(lat_section, GLODAP_cruise)

# remove columns not required for further analysis
GLODAP <- GLODAP %>% 
  select(-c(rCP_phosphate:cstar, cant_pos:cstar_tref_delta))

```


# Isoneutral slabs

The following boundaries for isoneutral slabs were defined:

- Atlantic: `r params_local$slabs_Atl`
- Indo-Pacific: `r params_local$slabs_Ind_Pac`

Continuous neutral densities (gamma) values from GLODAP are grouped into isoneutral slabs.

```{r cut_isoneutral_slabs}

GLODAP <- m_cut_gamma(GLODAP, "gamma")

```

```{r meridional_sections_isoneutral_slabs, fig.asp=0.6}

GLODAP_cruise <- GLODAP %>% 
  filter(cruise %in% params_global$cruises_meridional)

lat_section <- 
GLODAP_cruise %>%
  ggplot(aes(lat, depth)) +
  scale_y_reverse() +
  theme(legend.position = "bottom")

lat_section +
  geom_point(aes(col = gamma_slab)) +
  scale_color_viridis_d()

rm(lat_section, GLODAP_cruise)

```


```{r test_isoneutral_density_calculation, eval=FALSE, hide=TRUE}

# this section was only used to calculate gamma locally, and compare it to the value provided in GLODAP data set

GLODAP_cruise <- GLODAP %>% 
  filter(cruise %in% params_global$cruises_meridional)

library(oce)
library(gsw)
# calculate pressure from depth

GLODAP_cruise <- GLODAP_cruise %>% 
  mutate(CTDPRS = gsw_p_from_z(-depth,
                               lat))

GLODAP_cruise <- GLODAP_cruise %>% 
  mutate(THETA = swTheta(salinity = sal,
                         temperature = tem,
                         pressure = CTDPRS,
                         referencePressure = 0,
                         longitude = lon-180,
                         latitude = lat))

GLODAP_cruise <- GLODAP_cruise %>% 
  rename(LATITUDE = lat,
         LONGITUDE = lon,
         SALNTY = sal,
         gamma_provided = gamma)

library(reticulate)
source_python(here::here("code/python_scripts",
                         "Gamma_GLODAP_python.py"))

GLODAP_cruise <- calculate_gamma(GLODAP_cruise)

GLODAP_cruise <- GLODAP_cruise %>% 
  mutate(gamma_delta = gamma_provided - GAMMA)

lat_section <- 
GLODAP_cruise %>%
  ggplot(aes(LATITUDE, CTDPRS)) +
  scale_y_reverse() +
  theme(legend.position = "bottom")

lat_section +
  stat_summary_2d(aes(z = gamma_delta)) +
  scale_color_viridis_c()

GLODAP_cruise %>% 
  ggplot(aes(gamma_delta))+
  geom_histogram()

rm(lat_section, GLODAP_cruise, cruises_meridional)

```


# Observations coverage

```{r obs_gamma_slab_Atl, fig.asp=1}

GLODAP <- GLODAP %>% 
  mutate(gamma_slab = factor(gamma_slab), 
         gamma_slab = factor(gamma_slab, levels = rev(levels(gamma_slab))))

GLODAP %>% 
  filter(basin == "Atlantic") %>% 
  ggplot(aes(lat, gamma_slab)) +
  geom_bin2d(binwidth = 5) +
  scale_fill_viridis_c(option = "magma", direction = -1, trans = "log10") +
  scale_x_continuous(breaks = seq(-100,100,20)) +
  facet_grid(era~basin)

```


```{r obs_gamma_slab_Pac, fig.asp=1}

GLODAP %>% 
  filter(basin == "Indo-Pacific") %>% 
  ggplot(aes(lat, gamma_slab)) +
  geom_bin2d(binwidth = 5) +
  scale_fill_viridis_c(option = "magma", direction = -1, trans = "log10") +
  scale_x_continuous(breaks = seq(-100,100,20)) +
  facet_grid(era~basin)

```

## Histograms

```{r histogram_predictor, fig.asp=1}

GLODAP_vars <- GLODAP %>% 
  select(cstar_tref,
         sal,
         tem,
         oxygen,
         aou,
         silicate,
         phosphate,
         phosphate_star)

GLODAP_vars_long <- GLODAP_vars %>% 
  pivot_longer(cstar_tref:phosphate_star, names_to = "variable", values_to = "value")

GLODAP_vars_long %>% 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~ variable,
             ncol = 2,
             scales = "free")

rm(GLODAP_vars, GLODAP_vars_long)

```




# Individual cruise sections

Zonal and meridional section plots are produce for each cruise individually and can be downloaded [here](https://github.com/jens-daniel-mueller/Cant_eMLR/tree/master/output/figure/eMLR/data_preparation/all_cruises_clean){target="_blank"}.

```{r plot_all_individual_cruises_clean}

if (params_local$plot_all_figures == "y") {

cruises <- GLODAP %>% 
  group_by(cruise) %>% 
  summarise(date_mean = mean(date, na.rm = TRUE),
            n = n()) %>% 
  ungroup() %>% 
  arrange(date_mean)

GLODAP <- full_join(GLODAP, cruises)

n <- 0
for (i_cruise in unique(cruises$cruise)) {

# i_cruise <- unique(cruises$cruise)[1]
n <- n + 1
print(n)  
  
GLODAP_cruise <- GLODAP %>%
  filter(cruise == i_cruise) %>% 
  arrange(date)

cruises_cruise <- cruises %>%
  filter(cruise == i_cruise)
  
map_plot <- 
  map +
  geom_point(data = GLODAP_cruise,
             aes(lon, lat, col = date)) +
  scale_color_viridis_c(trans = "date") +
  labs(title = paste("Mean date:", cruises_cruise$date_mean,
                     "| cruise:", cruises_cruise$cruise,
                     "| n(samples):", cruises_cruise$n))


lon_section <- GLODAP_cruise %>%
  ggplot(aes(lon, depth)) +
  scale_y_reverse() +
  scale_fill_viridis_c()

lon_tco2 <- lon_section+
  stat_summary_2d(aes(z=tco2))

lon_talk <- lon_section+
  stat_summary_2d(aes(z=talk))

lon_phosphate <- lon_section+
  stat_summary_2d(aes(z=phosphate))

lon_oxygen <- lon_section+
  stat_summary_2d(aes(z=oxygen))

lon_aou <- lon_section+
  stat_summary_2d(aes(z=aou))

lon_phosphate_star <- lon_section+
  stat_summary_2d(aes(z=phosphate_star))

lon_nitrate <- lon_section+
  stat_summary_2d(aes(z=nitrate))

lon_cstar <- lon_section+
  stat_summary_2d(aes(z=cstar_tref))


lat_section <- GLODAP_cruise %>%
  ggplot(aes(lat, depth)) +
  scale_y_reverse() +
  scale_fill_viridis_c()

lat_tco2 <- lat_section+
  stat_summary_2d(aes(z=tco2))

lat_talk <- lat_section+
  stat_summary_2d(aes(z=talk))

lat_phosphate <- lat_section+
  stat_summary_2d(aes(z=phosphate))

lat_oxygen <- lat_section+
  stat_summary_2d(aes(z=oxygen))

lat_aou <- lat_section+
  stat_summary_2d(aes(z=aou))

lat_phosphate_star <- lat_section+
  stat_summary_2d(aes(z=phosphate_star))

lat_nitrate <- lat_section+
  stat_summary_2d(aes(z=nitrate))

lat_cstar <- lat_section+
  stat_summary_2d(aes(z=cstar_tref))

hist_tco2 <- GLODAP_cruise %>%
  ggplot(aes(tco2)) +
  geom_histogram()

hist_talk <- GLODAP_cruise %>%
  ggplot(aes(talk)) +
  geom_histogram()

hist_phosphate <- GLODAP_cruise %>%
  ggplot(aes(phosphate)) +
  geom_histogram()

hist_oxygen <- GLODAP_cruise %>%
  ggplot(aes(oxygen)) +
  geom_histogram()

hist_aou <- GLODAP_cruise %>%
  ggplot(aes(aou)) +
  geom_histogram()

hist_phosphate_star <- GLODAP_cruise %>%
  ggplot(aes(phosphate_star)) +
  geom_histogram()

hist_nitrate <- GLODAP_cruise %>%
  ggplot(aes(nitrate)) +
  geom_histogram()

hist_cstar <- GLODAP_cruise %>%
  ggplot(aes(cstar_tref)) +
  geom_histogram()

(map_plot /
    ((hist_tco2 / hist_talk / hist_phosphate / hist_cstar) |
       (hist_oxygen / hist_phosphate_star / hist_nitrate / hist_aou)
    )) |
  ((lat_tco2 / lat_talk / lat_phosphate / lat_oxygen / lat_aou / lat_phosphate_star / lat_nitrate / lat_cstar) |
     (lon_tco2 / lon_talk / lon_phosphate / lon_oxygen /  lon_aou /lon_phosphate_star / lon_nitrate / lon_cstar))    

ggsave(
  path = paste(path_version_figures, "Cruise_sections_histograms/", sep = ""),
  filename = paste(
    "Cruise_date",
    cruises_cruise$date_mean,
    "count",
    cruises_cruise$n,
    "cruiseID",
    cruises_cruise$cruise,
    ".png",
    sep = "_"
  ),
width = 20, height = 12)

rm(map_plot,
   lon_section, lat_section,
   lat_tco2, lat_talk, lat_phosphate, lon_tco2, lon_talk, lon_phosphate,
   GLODAP_cruise, cruises_cruise)

}

}

```


# Write files

```{r write_csv_GLODAP_used_for_fitting}

# rename the cstar variable,
# in scripts downstrea from here, we refer to cstar as the value adjusted to the reference year
GLODAP <- GLODAP %>% 
  rename(cstar = cstar_tref)

GLODAP %>% write_csv(paste(path_version_data,
                             "GLODAPv2.2020_MLR_fitting_ready.csv",
                             sep = ""))

co2_atm_tref %>%  write_csv(paste(path_version_data,
                             "co2_atm_tref.csv",
                             sep = ""))

```
