---
title: "GLODAPv2_2020"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/updata/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}
params_local <-
  read_rds(here::here("data/auxillary",
                       "params_local.rds"))

```

```{r define_updata_paths, include = FALSE}
path_preprocessing    <-
  "/nfs/kryo/work/updata/emlr_cant/observations/preprocessing/"

path_version_data     <-
  paste(
    "/nfs/kryo/work/updata/emlr_cant/observations/",
    params_local$Version_ID,
    "/data/",
    sep = ""
  )

path_version_figures  <-
  paste(
    "/nfs/kryo/work/updata/emlr_cant/observations/",
    params_local$Version_ID,
    "/figures/",
    sep = ""
  )
```

```{r load_libraries_specific, include = FALSE}
library(lubridate)
```

# Read files

Main data source for this project is the preprocessed version of the `GLODAPv2.2020_Merged_Master_File.csv` downloaded from [glodap.info](https://www.glodap.info/){target="_blank"} in June 2020.

```{r read_GLODAPv2_2020_merged_master_file}

GLODAP <-
  read_csv(paste(path_preprocessing,
                 "GLODAPv2.2020_preprocessed.csv",
                 sep = ""))


```


# Data preparation

## Reference eras

Samples were assigned to following eras:

```{r assign_eras}

# create clear labels for era
labels <- bind_cols(
  start = params_local$era_breaks+1,
  end = lead(params_local$era_breaks))

labels <- labels %>% 
  filter(!is.na(end)) %>% 
  mutate(end = if_else(end == Inf, max(GLODAP$year), end),
         label = paste(start, end, sep = "-")) %>% 
  select(label) %>% 
  pull()

# cut observation years into era applying the labels
GLODAP <- GLODAP %>%
  filter(year > params_local$era_breaks[1]) %>% 
  mutate(era = cut(year,
                   params_local$era_breaks,
                   labels = labels))

levels(GLODAP$era)
rm(labels)

```


## Spatial boundaries

### Depth

Observations collected shallower than:

- minimum sampling depth:  `r params_local$depth_min`m

were excluded from the analysis to avoid seasonal bias.

```{r apply_depth_min}

GLODAP <- GLODAP %>% 
  filter(depth >= params_local$depth_min)

```

### Bottomdepth

Observations collected in an area with a:

- minimum bottom depth:  `r params_local$bottomdepth_min`m

were excluded from the analysis to avoid coastal impacts. Please note that minimum bottom depth criterion of 0m means that no filtering was applied here.

```{r apply_bottomdepth_min}

GLODAP <- GLODAP %>% 
  filter(bottomdepth >= params_local$bottomdepth_min)

```


## Flags and missing data

Only rows (samples) for which all relevant parameters are available were selected, ie NA's were removed.

According to Olsen et al (2020), flags within the merged master file identify:

- f:  

  - 2: Acceptable
  - 0: Interpolated (nutrients/oxygen) or calculated (CO[2] params_local) value
  - 9: Data not used (so, only NA data should have this flag)
  
  
- qc:  

  - 1: Adjusted or unadjusted data
  - 0: Data appear of good quality but have not been subjected to full secondary QC
  - data with poor or uncertain quality are excluded.

Following flagging criteria were taken into account:

- flag_f:     `r params_local$flag_f`
- flag_qc:    `r params_local$flag_qc`

Summary statistics were calculated during cleaning process.

### tco2

#### NA

Rows with missing tco2 observations were already removed in the preprocessing.

```{r tco2_na_subset}

GLODAP_stats <- GLODAP %>% 
  summarise(tco2_values = n())

GLODAP_obs_grid <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "tco2_values")

```

```{r WOA18_basin_mask_GLODAP_observations_map, fig.asp=0.6}

GLODAP_obs <- GLODAP %>% 
  group_by(lat, lon) %>% 
  summarise(n = n()) %>% 
  ungroup()

map +
  geom_raster(data = basinmask, aes(lon, lat, fill = basin)) +
  geom_raster(data = GLODAP_obs, aes(lon, lat)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "top",
        legend.title = element_blank())

rm(GLODAP_obs)

```


#### f flag

```{r tco2_f_subset, fig.asp=0.8}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, tco2f)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ tco2f) +
  theme(legend.position = "top")

rm(GLODAP_obs_grid_temp)

GLODAP <- GLODAP %>%
  filter(tco2f %in% params_local$flag_f)


```

#### qc flag

```{r tco2_qc_subset, fig.asp=0.8}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, tco2qc)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ tco2qc) +
  theme(legend.position = "top")

##

GLODAP <- GLODAP %>%
  filter(tco2qc %in% params_local$flag_qc)

GLODAP_stats_temp <- GLODAP %>%
  summarise(tco2_flag = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "tco2_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### talk

#### NA

```{r talk_na_subset, fig.asp=0.8}

GLODAP <- GLODAP %>% 
  mutate(talkna = if_else(is.na(talk), "NA", "Value"))

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkna)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkna) +
  theme(legend.position = "top")

GLODAP <- GLODAP %>% 
  select(-talkna) %>% 
  filter(!is.na(talk))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(talk_values = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "talk_values")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

#### f flag

```{r talk_f_subset, fig.asp=0.8}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkf)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkf) +
  theme(legend.position = "top",
        legend.title = element_blank())


# ###

GLODAP <- GLODAP %>%
  filter(talkf %in% params_local$flag_f)


```

#### qc flag

```{r talk_qc_subset, fig.asp=0.8}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkqc)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkqc) +
  theme(legend.position = "top",
        legend.title = element_blank())

###

GLODAP <- GLODAP %>%
  filter(talkqc %in% params_local$flag_qc)

##

GLODAP_stats_temp <- GLODAP %>%
  summarise(talk_flag = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "talk_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### Phosphate

#### NA

```{r phosphate_na_subset, fig.asp=0.8}

GLODAP <- GLODAP %>% 
  mutate(phosphatena = if_else(is.na(phosphate), "NA", "Value"))

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphatena)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ phosphatena) +
  theme(legend.position = "top")

GLODAP <- GLODAP %>% 
  select(-phosphatena) %>% 
  filter(!is.na(phosphate))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(phosphate_values = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "phosphate_values")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

#### f flag

```{r phosphate_f_subset, fig.asp=0.8}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphatef)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~phosphatef) +
  theme(legend.position = "top",
        legend.title = element_blank())

###

GLODAP <- GLODAP %>%
  filter(phosphatef %in% params_local$flag_f)


```

#### qc flag

```{r phosphate_qc_subset, fig.asp=0.8}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphateqc)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~phosphateqc) +
  theme(legend.position = "top",
        legend.title = element_blank())


###

GLODAP <- GLODAP %>%
  filter(phosphateqc %in% params_local$flag_qc)

##

GLODAP_stats_temp <- GLODAP %>%
  summarise(phosphate_flag = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "phosphate_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### eMLR variables

```{r eMLR_subset}

GLODAP <- GLODAP %>% 
  filter(!is.na(tem))

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(sal))

GLODAP <- GLODAP %>%
  filter(salinityf %in% params_local$flag_f)

GLODAP <- GLODAP %>%
  filter(salinityqc %in% params_local$flag_qc)

##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(silicate))

GLODAP <- GLODAP %>%
  filter(silicatef %in% params_local$flag_f)

GLODAP <- GLODAP %>%
  filter(silicateqc %in% params_local$flag_qc)
##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(oxygen))

GLODAP <- GLODAP %>%
  filter(oxygenf %in% params_local$flag_f)

GLODAP <- GLODAP %>%
  filter(oxygenqc %in% params_local$flag_qc)
##

GLODAP <- GLODAP %>% 
  filter(!is.na(aou))

GLODAP <- GLODAP %>%
  filter(aouf %in% params_local$flag_f)
##
  
GLODAP <- GLODAP %>% 
  filter(!is.na(nitrate))

GLODAP <- GLODAP %>%
  filter(nitratef %in% params_local$flag_f)

GLODAP <- GLODAP %>%
  filter(nitrateqc %in% params_local$flag_qc)
##

GLODAP <- GLODAP %>% 
  filter(!is.na(depth))

GLODAP <- GLODAP %>% 
  filter(!is.na(gamma))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(eMLR_variables = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)

rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "eMLR_variables")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

```{r remove_flag_columns}

GLODAP <- GLODAP %>% 
  select(-ends_with(c("f", "qc")))

```


### Manual adjustment A16 cruise

For harmonization with Gruber et al. (2019), cruises 1041 (A16N) and 1042 (A16S) were grouped into the `r sort(unique(GLODAP$era))[2]` era despite taking place in 2013/14.

```{r GLODAP_Atl_cruises_2013_2014_map}

GLODAP_cruises <- GLODAP %>% 
  filter(basin == "Atlantic",
         year %in% c(2013, 2014)) %>% 
  count(lat, lon, cruise)

map +
  geom_raster(data = GLODAP_cruises, aes(lon, lat, fill = as.factor(cruise))) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "top",
        legend.title = element_blank())

rm(GLODAP_cruises)

```

```{r switch_eras_manually}

GLODAP <- GLODAP %>%
  mutate(era = if_else(cruise %in% c(1041, 1042),
                       sort(unique(GLODAP$era))[2], era))

```

## Create clean observations grid

Grid containing all grid cells where at least one observation remains available after cleaning.

```{r create_clean_obs_grid}

GLODAP_obs_grid_clean <- GLODAP %>% 
  count(lat, lon) %>% 
  select(-n)

```



## Write summary file
 
```{r write_clean_data_files}

GLODAP_obs_grid_clean  %>%  write_csv(paste(path_version_data,
                                            "GLODAPv2.2020_clean_obs_grid.csv",
                                            sep = ""))

# select relevant columns for further analysis
GLODAP <- GLODAP %>% 
  select(year, date, era, basin, lat, lon, cruise,
         bottomdepth, depth, tem, sal, gamma,
         tco2, talk, phosphate,
         oxygen, aou, nitrate, silicate)


GLODAP  %>%  write_csv(paste(path_version_data,
                             "GLODAPv2.2020_clean.csv",
                             sep = ""))


```


# Overview plots

## Cleaning stats

Number of observations at various steps of data cleaning.

```{r GLODAP_cleaning_stats}

GLODAP_stats_long <- GLODAP_stats %>%
  pivot_longer(1:length(GLODAP_stats),
               names_to = "parameter",
               values_to = "n")

GLODAP_stats_long <- GLODAP_stats_long %>%
  mutate(parameter = fct_reorder(parameter, n))

GLODAP_stats_long %>% 
  ggplot(aes(parameter, n/1000)) +
  geom_col() +
  coord_flip() +
  labs(y = "n (1000)") +
  theme(axis.title.y = element_blank())

rm(GLODAP_stats_long)

```


## Assign coarse spatial grid

For the following plots, the cleaned data set was re-opened and observations were gridded spatially to intervals of:  

- 5° x 5°

```{r grid_spatially_5x5}

GLODAP <- m_grid_horizontal_coarse(GLODAP)

```


## Histogram Zonal coverage

```{r coverage_histogram_zonal}

GLODAP_histogram_lat <- GLODAP %>%
  group_by(era, lat_grid, basin) %>%
  tally() %>%
  ungroup()

GLODAP_histogram_lat %>%
  ggplot(aes(lat_grid, n, fill = era)) +
  geom_col() +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap( ~ basin) +
  coord_flip(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank())

rm(GLODAP_histogram_lat)

```

## Histogram temporal coverage

```{r coverage_histogram_temporal}

GLODAP_histogram_year <- GLODAP %>%
  group_by(year, basin) %>%
  tally() %>%
  ungroup()

era_median_year <- GLODAP %>%
  group_by(era) %>%
  summarise(t_ref = median(year)) %>%
  ungroup()

era_median_year

GLODAP_histogram_year %>%
  ggplot() +
  geom_vline(xintercept = c(
    params_local$era_breaks + 0.5
  )) +
  geom_col(aes(year, n, fill = basin)) +
  geom_point(
    data = era_median_year,
    aes(t_ref, 0, shape = "Median year"),
    size = 2,
    fill = "white"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  scale_shape_manual(values = 24, name = "") +
  scale_y_continuous() +
  coord_cartesian(expand = 0) +
  theme(
    legend.position = "top",
    legend.direction = "vertical",
    legend.title = element_blank(),
    axis.title.x = element_blank()
  )

rm(GLODAP_histogram_year,
   era_median_year)

```

## Zonal temporal coverage (Hovmoeller)

```{r coverage_hovmoeller, fig.asp=1}

GLODAP_hovmoeller_year <- GLODAP %>%
  group_by(year, lat_grid, basin) %>%
  tally() %>%
  ungroup()

GLODAP_hovmoeller_year %>%
  ggplot(aes(year, lat_grid, fill = n)) +
  geom_tile() +
  geom_vline(xintercept = c(1999.5, 2012.5)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_wrap( ~ basin, ncol = 1) +
  theme(legend.position = "top",
        axis.title.x = element_blank())

rm(GLODAP_hovmoeller_year)

```

## Coverage maps by era

### Subsetting process

The following plots show the remaining data after several cleaning steps, separately for each era.

```{r coverage_maps_era_subsetting_era_cleaning_level, fig.asp=1.1}

GLODAP_obs_grid <- GLODAP_obs_grid %>%
  mutate(cleaning_level = factor(
           cleaning_level,
           unique(GLODAP_obs_grid$cleaning_level)
         ))

map +
  geom_raster(data = GLODAP_obs_grid %>%
                filter(cleaning_level == "tco2_values") %>%
                select(-cleaning_level),
              aes(lon, lat, fill = "tco2_values")) +
  geom_raster(data = GLODAP_obs_grid %>%
                filter(cleaning_level != "tco2_values"),
              aes(lon, lat, fill = "subset")) +
  scale_fill_brewer(palette = "Set1", name = "") +
  facet_grid(cleaning_level ~ era) +
  theme(legend.position = "top",
        axis.title = element_blank())

```

### Final input data

The following plots show the remaining data density in each grid cell after all cleaning steps, separately for each era.

```{r coverage_maps_era_clean, fig.asp=1.3}

GLODAP_tco2_grid <- GLODAP %>%
  count(lat, lon)

map +
  # geom_raster(data = GLODAP_tco2_grid, aes(lon, lat), fill = "grey80") +
  geom_bin2d(data = GLODAP,
             aes(lon, lat),
             binwidth = c(1,1)) +
  scale_fill_viridis_c(option = "magma", direction = -1, trans = "log10") +
  facet_wrap(~era, ncol = 1) +
  labs(title = "Cleaned GLODAP observations",
       subtitle = paste("Version:", params_local$Version_ID)) +
  theme(axis.title = element_blank())

ggsave(path = path_version_figures,
       filename = "data_distribution_era.png",
       height = 8,
       width = 5)

```

