---
title: "eMLR - model fitting"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
params:
  Version_ID: "v_XXX"
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup_obs.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}

if (params$Version_ID == "v_XXX") {
  params_local <-
    read_rds(here::here("data/auxillary", "params_local.rds"))
  
} else {
  params_local <-
    read_rds(
      file = paste(path_root, "observations",
                   params$Version_ID,
                   "data/params_local.rds",
                   sep = "/")
    )
  }

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(lubridate)
library(GGally)
library(olsrr)
library(knitr)
library(kableExtra)
library(broom)
library(corrr)
library(GGally)
library(gt)
library(tictoc)
```

# Version ID

The results displayed on this site correspond to the Version_ID: `r params$Version_ID`

# Required data

Required are:

- cleaned and prepared GLODAPv2_2020 file

```{r read_GLODAP_used_for_fitting}

GLODAP <-
  read_csv(paste(path_version_data,
                 "GLODAPv2.2020_MLR_fitting_ready.csv",
                 sep = ""))

```

# Predictor combinations

Find all possible combinations of following considered predictor variables:  

- `r params_local$MLR_predictors`

```{r eMLR_find_all_predictor_combinations, results='hide'}

# the following code is a workaround to find all predictor combinations
# using the olsrr package and fit all models for one era, slab, and basin

i_basin <- unique(GLODAP$basin)[1]
i_era   <- unique(GLODAP$era)[1]

# subset one basin and era for fitting
GLODAP_basin_era <- GLODAP %>%
  filter(basin == i_basin, era == i_era,
         data_source == "obs")

i_gamma_slab <- unique(GLODAP_basin_era$gamma_slab)[3]
print(i_gamma_slab)

# subset one gamma slab
GLODAP_basin_era_slab <- GLODAP_basin_era %>%
  filter(gamma_slab == i_gamma_slab)

# fit the full linear model, i.e. all predictor combinations
lm_full <- lm(paste(
  params_local$MLR_target,
  paste(params_local$MLR_predictors, collapse = " + "),
  sep = " ~ "
),
data = GLODAP_basin_era_slab)

# fit linear models for all possible predictor combinations
# unfortunately, this functions does not provide model coefficients (yet)
lm_all <- ols_step_all_possible(lm_full)

# convert to tibble
lm_all <- as_tibble(lm_all)

# format model formula
lm_all <- lm_all %>% 
  select(n, predictors) %>% 
  mutate(model = str_replace_all(predictors, " ", " + "),
         model = paste(params_local$MLR_target, "~", model))

# remove helper objects
rm(i_gamma_slab,
   i_era,
   i_basin,
   GLODAP_basin_era,
   GLODAP_basin_era_slab,
   lm_full)

```

# Apply predictor threshold

Select combinations with a total number of predictors in the range:  

- Minimum: `r params_local$MLR_predictors_min`  
- Maximum: `r params_local$MLR_predictors_max`

```{r apply_MLR_predictor_threshold}

lm_all <- lm_all %>% 
  filter(n >= params_local$MLR_predictors_min,
         n <= params_local$MLR_predictors_max)

```

This results in a total number of MLR models of:  

- `r nrow(lm_all)`

# Fit all models

Individual linear regression models were fitted for the chosen target variable:  

- `r params_local$MLR_target`

as a function of each predictor combination. Fitting was performed separately within each basin, era, and slab. Model diagnostics, such as the root mean squared error (RMSE), were calculated for each fitted model.


```{r fit_all_models_purrr, results='hide'}

# prepare nested data frame
GLODAP %>%
  filter_all(any_vars(is.na(.)))

GLODAP_nested <- GLODAP %>% 
  # filter(gamma_slab %in% unique(GLODAP$gamma_slab)[c(5,6)]) %>%
  nest(data = !c(gamma_slab, era, basin, data_source))

# expand with model definitions
GLODAP_nested_lm <- expand_grid(
  GLODAP_nested,
  lm_all#[1:3,]
)

# fit models and extract tidy model output
if (params_local$MLR_function == "lm") {
GLODAP_nested_lm_fit <- GLODAP_nested_lm %>% 
  mutate(
    fit = map2(.x = data, .y = model, 
               ~ lm(as.formula(.y), data = .x))
  )
}

if (params_local$MLR_function == "rlm") {
GLODAP_nested_lm_fit <- GLODAP_nested_lm %>% 
  mutate(
    fit = map2(.x = data, .y = model, 
               ~ MASS::rlm(as.formula(.y), data = .x))
  )
}


GLODAP_nested_lm_fit <- GLODAP_nested_lm_fit %>% 
  mutate(
    tidied = map(fit, tidy),
    glanced = map(fit, glance),
    augmented = map(fit, augment),
    vif = map(fit, ols_vif_tol)
  )

```

# Tidy models

```{r tidy_model_output}

# extract glanced model output (model diagnostics, such as AIC)
GLODAP_glanced <- GLODAP_nested_lm_fit %>%
  select(-c(data, fit, tidied, augmented, vif)) %>%
  unnest(glanced) %>% 
  rename(n_predictors = n)

# extract tidy model output (model coefficients)
GLODAP_tidy <- GLODAP_nested_lm_fit %>%
  select(-c(data, fit, glanced, augmented, vif)) %>%
  unnest(tidied)

# extract augmented model output (fitted values and residuals)
GLODAP_augmented <- GLODAP_nested_lm_fit %>% 
  select(-c(data, fit, tidied, glanced, vif)) %>% 
  unnest(augmented) 

# print(object.size(GLODAP_augmented), units = "MB")

# extract VIF from output
GLODAP_glanced_vif <- GLODAP_nested_lm_fit %>%
  select(-c(data, fit, tidied, augmented, glanced)) %>%
  unnest(vif)

# calculte max VIF per model
GLODAP_glanced_vif_max <- GLODAP_glanced_vif %>% 
  group_by(data_source, era, basin, gamma_slab, model, n) %>% 
  summarise(vif_max = max(VIF)) %>% 
  ungroup()

# calculate RMSE from augmented output
GLODAP_glanced_rmse <- GLODAP_augmented %>%
  group_by(data_source, era, basin, gamma_slab, model) %>%
  summarise(rmse = sqrt(c(crossprod(.resid)) / length(.resid))) %>%
  ungroup()

# add RMSE and vif_max to glanced output
GLODAP_glanced <- full_join(GLODAP_glanced, GLODAP_glanced_rmse)
GLODAP_glanced <- full_join(GLODAP_glanced, GLODAP_glanced_vif_max)
rm(GLODAP_glanced_rmse)
rm(GLODAP_glanced_vif_max)


# extract input data
GLODAP_data <- GLODAP_nested_lm_fit %>%
  select(-c(fit, tidied, glanced, augmented, vif)) %>%
  unnest(data)

# append input data with augmented data
GLODAP_augmented <- bind_cols(
  GLODAP_data,
  GLODAP_augmented %>% select(.fitted, .resid)
)

rm(GLODAP, GLODAP_nested, GLODAP_nested_lm, GLODAP_nested_lm_fit, lm_all,
   GLODAP_data)

```




# Prepare coeffcients

Coefficients are prepared for the mapping of Cant and the chosen target variable.


## VIF threshold

To avoid multicollinearity among predictors, models were excluded with a VIF above:  

- `r params_local$vif_max`

After removing models affected by multicollinearity, the targeted number of MLRs (`r params_local$MLR_number`) was undercut in following fitting units:

```{r VIF_threshold}

GLODAP_glanced_clean <- GLODAP_glanced %>% 
  filter(vif_max <= params_local$vif_max)

GLODAP_glanced_clean %>% 
  group_by(data_source, era, basin, gamma_slab) %>% 
  count() %>% 
  filter(n < params_local$MLR_number)

```


## Predictor selection

Within each basin and slab, the following number of best linear regression models was selected:  

- `r params_local$MLR_number`

The criterion used to select the best models was:

- `r params_local$MLR_criterion`

The criterion was summed up for two adjacent eras, and the models with lowest summed values were selected.


```{r predictor_selection}

# calculate RMSE sum for adjacent eras
lm_all <- GLODAP_glanced_clean  %>%
  select(data_source, basin, gamma_slab, model, era, AIC, BIC, rmse) %>% 
  arrange(era) %>% 
  group_by(data_source, basin, gamma_slab, model) %>% 
  mutate(rmse_sum = rmse + lag(rmse),
         aic_sum = AIC + lag(AIC),
         bic_sum = BIC + lag(BIC)
         ) %>% 
  ungroup() %>% 
  select(-c(era)) %>% 
  drop_na()

# subset models with lowest summed criterion
# chose which criterion is applied

if (params_local$MLR_criterion == "aic") {
  lm_best <- lm_all %>%
    group_by(data_source, basin, gamma_slab) %>%
    slice_min(order_by = aic_sum,
              with_ties = FALSE,
              n = params_local$MLR_number) %>%
    ungroup() %>%
    arrange(basin, gamma_slab, model)
} 

if (params_local$MLR_criterion == "bic") {
  lm_best <- lm_all %>%
    group_by(data_source, basin, gamma_slab) %>%
    slice_min(order_by = bic_sum,
              with_ties = FALSE,
              n = params_local$MLR_number) %>%
    ungroup() %>%
    arrange(basin, gamma_slab, model)
}

if (params_local$MLR_criterion == "rmse") {
  lm_best <- lm_all %>%
    group_by(data_source, basin, gamma_slab) %>%
    slice_min(order_by = rmse_sum,
              with_ties = FALSE,
              n = params_local$MLR_number) %>%
    ungroup() %>%
    arrange(basin, gamma_slab, model)
}


# print table
lm_best %>% 
  kable() %>%
  add_header_above() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")

```

## Alternative model selection

```{r vif_rmse_model_selection, fig.asp=2, eval=FALSE}

GLODAP_glanced %>% 
  filter(data_source == "obs",
         era == "2010-2019") %>% 
  ggplot(aes(rmse, log10(vif_max))) +
  geom_bin2d() +
  scale_fill_viridis_c() +
  facet_grid(gamma_slab ~ basin)

GLODAP_glanced %>% 
  filter(data_source == "obs",
         era == "2010-2019") %>% 
  ggplot(aes(rmse, r.squared)) +
  geom_bin2d() +
  scale_fill_viridis_c() +
  facet_grid(gamma_slab ~ basin)

# calculate RMSE sum for adjacent eras
lm_all_new <- GLODAP_glanced  %>%
  select(data_source, basin, gamma_slab, model, era, AIC, BIC, rmse) %>% 
  arrange(era) %>% 
  group_by(data_source, basin, gamma_slab, model) %>% 
  mutate(rmse_sum = rmse + lag(rmse),
         aic_sum = AIC + lag(AIC),
         bic_sum = BIC + lag(BIC)
         ) %>% 
  ungroup() %>% 
  select(-c(era)) %>% 
  drop_na()

lm_best_rmse_sum_max <- lm_all_new %>%
  group_by(data_source, basin, gamma_slab) %>%
  slice_min(order_by = rmse_sum,
            with_ties = FALSE,
            n = 1) %>%
  ungroup() %>%
  arrange(basin, gamma_slab, model) %>% 
  mutate(rmse_sum_max = rmse_sum*1.1) %>% 
  select(data_source, basin, gamma_slab, rmse_sum_max)

lm_best_new <- 
  full_join(lm_all_new, lm_best_rmse_sum_max) %>% 
  filter(rmse_sum < rmse_sum_max)

GLODAP_glanced_new <- 
  left_join(lm_best_new %>% select(data_source, basin, gamma_slab, model),
            GLODAP_glanced) 

GLODAP_glanced_new %>% 
  filter(data_source == "obs",
         era == "2010-2019") %>% 
  ggplot(aes(rmse, log10(vif_max))) +
  geom_bin2d() +
  scale_fill_viridis_c() +
  facet_grid(gamma_slab ~ basin)

GLODAP_glanced_new %>% 
  filter(data_source == "obs",
         era == "2010-2019") %>% 
  ggplot(aes(rmse, r.squared)) +
  geom_bin2d() +
  scale_fill_viridis_c() +
  facet_grid(gamma_slab ~ basin)

# GLODAP_glanced_new %>%
#   group_by(data_source, basin, gamma_slab) %>%
#   count() %>% 
#   arrange(n)

lm_best_new <- GLODAP_glanced_new %>%
  select(data_source, basin, gamma_slab, model, era, rmse, vif_max) %>%
  arrange(era) %>%
  group_by(data_source, basin, gamma_slab, model) %>%
  mutate(
    vif_max_sum = vif_max + lag(vif_max)
  ) %>%
  ungroup() %>%
  select(-c(era)) %>%
  drop_na() %>%
  group_by(data_source, basin, gamma_slab) %>%
  slice_min(order_by = vif_max_sum,
            with_ties = FALSE,
            n = params_local$MLR_number) %>%
  ungroup() %>%
  arrange(basin, gamma_slab, model)

# ggplot() +
#   geom_point(data = GLODAP_glanced %>%
#                filter(data_source == "obs",
#                       era == "2010-2019"),
#              aes(rmse, log10(vif_max), col = "all")) +
#   geom_point(
#     data = GLODAP_glanced_new %>%
#       filter(data_source == "obs",
#              era == "2010-2019"),
#     aes(rmse, log10(vif_max), col = "rmse max")
#   ) +
#   geom_point(
#     data = GLODAP_glanced_clean_new %>%
#       filter(data_source == "obs",
#              era == "2010-2019"),
#     aes(rmse, log10(vif_max), col = "rmse + vif max")
#   ) +
#   facet_grid(gamma_slab ~ basin)

lm_all <- lm_all_new

lm_best <- 
  left_join(lm_best_new %>% select(-c(vif_max, vif_max_sum)),
            lm_all)

GLODAP_glanced_clean <- GLODAP_glanced_new

```




## Target variable coefficients

A data frame to map the target variable is prepared.

```{r prepare_model_coefficients_table_target_variable_mapping}

# extend best model selection with coefficients

lm_best <- left_join(
  lm_best,
  GLODAP_tidy %>% select(data_source, basin, gamma_slab, era, model, term, estimate)
)

```

## dcant coeffcients

A data frame of coefficient offsets is prepared to facilitate the direct mapping of dcant.

```{r prepare_delta_coefficients_for_cant_mapping}

# subtract coefficients of adjacent era  
lm_best_dcant <- lm_best %>%
  select(-c(AIC, BIC, rmse, aic_sum, bic_sum, rmse_sum)) %>% 
  arrange(era) %>%
  group_by(data_source, basin, gamma_slab, model, term) %>%
  mutate(delta_coeff = estimate - lag(estimate)) %>%
  ungroup() %>%
  arrange(basin, gamma_slab, model, term) %>%
  drop_na() %>%
  select(-c(era,estimate))

# pivot to wide format
lm_best_dcant <- lm_best_dcant %>%
  pivot_wider(values_from = delta_coeff,
              names_from = term,
              names_prefix = "delta_coeff_",
              values_fill = 0)

```

```{r target_coefficient_table_wide}

# create table of target variable coefficients in wide format
lm_best_target <- lm_best %>% 
  select(-c(AIC, BIC, rmse, aic_sum, bic_sum, rmse_sum)) %>% 
  pivot_wider(names_from = "term",
              names_prefix = "coeff_",
              values_from = "estimate",
              values_fill = 0
              )

```



```{r subset_best_models}

GLODAP_augmented_best <- left_join(
  lm_best_target %>% select(data_source, basin, gamma_slab, era, model),
  GLODAP_augmented
)

```


## Write files

```{r write_model_coefficients_files}


lm_best_target %>%
  write_csv(paste(path_version_data,
                  "lm_best_target.csv",
                  sep = ""))

lm_best_dcant %>%
  write_csv(paste(path_version_data,
                  "lm_best_dcant.csv",
                  sep = ""))

GLODAP_augmented_best %>%
  write_csv(paste(path_version_data,
                  "lm_residuals_best_models.csv",
                  sep = ""))

GLODAP_glanced %>%
  write_csv(paste(path_version_data,
                  "lm_model_metrics.csv",
                  sep = ""))


```

# Model diagnotics

## Selection criterion vs predictors

The selection criterion (`r params_local$MLR_criterion`) was plotted against the number of predictors (limited to `r params_local$MLR_predictors_min` - `r params_local$MLR_predictors_max`).

### All models

```{r eMLR_predictor_model_quality_all, fig.asp=2}

GLODAP_glanced %>%
  group_split(era) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(
               as.factor(n_predictors),
               !!sym(params_local$MLR_criterion),
               col = data_source
             )) +
      geom_hline(yintercept = c(0, 10)) +
      geom_boxplot() +
      facet_grid(gamma_slab ~ basin) +
      scale_color_brewer(palette = "Set1") +
      ylim(c(0, NA)) +
      labs(x = "Number of predictors",
           title = paste("Era:", unique(.x$era)))
  )



```


### Best models

```{r eMLR_predictor_model_quality_best, fig.asp=2}

left_join(lm_best_target %>% select(basin, gamma_slab, era, model),
          GLODAP_glanced) %>%
  group_split(era) %>%
  # head(1) %>%
  imap(
  ~ ggplot(data = .x,
         aes("",
             !!sym(params_local$MLR_criterion),
             col = data_source)) +
  geom_hline(yintercept = c(0, 10)) +
  geom_boxplot() +
  facet_grid(gamma_slab ~ basin) +
  scale_color_brewer(palette = "Set1") +
  ylim(c(0, NA)) +
  labs(x = "Number of predictors pooled",
           title = paste("Era:", unique(.x$era)))
  )

```


## RMSE correlation between eras

RMSE was plotted to compare the agreement for one model applied to two adjecent eras (ie check whether the same predictor combination performs equal in both eras).

### All models

```{r rmse_correlation_across_eras_all_models}

# find max rmse to scale axis
max_rmse <-
  max(c(lm_all$rmse,
        lm_all$rmse_sum - lm_all$rmse))

lm_all %>%
  ggplot(aes(rmse, rmse_sum - rmse, col = gamma_slab)) +
  geom_point() +
  scale_color_viridis_d() +
  coord_equal(xlim = c(0, max_rmse),
              ylim = c(0, max_rmse)) +
  geom_abline(slope = 1,
              col = 'red') +
  facet_grid(data_source ~ basin)

rm(max_rmse)

```


### Best models

```{r rmse_correlation_across_eras_best_models}

# find max rmse to scale axis
max_rmse <-
  max(c(lm_best$rmse,
        lm_best$rmse_sum - lm_best$rmse))

lm_best %>%
  ggplot(aes(rmse, rmse_sum - rmse, col = gamma_slab)) +
  geom_point() +
  scale_color_viridis_d() +
  coord_equal(xlim = c(0, max_rmse),
              ylim = c(0, max_rmse)) +
  geom_abline(slope = 1,
              col = 'red') +
  facet_grid(data_source ~ basin)

rm(max_rmse)
```

## Predictor counts

The number of models where a particular predictor was included were counted for each basin, density slab and compared eras

```{r predictor_counts}

# calculate cases of predictor used
lm_best_predictor_counts <- lm_best_dcant %>% 
  pivot_longer(starts_with("delta_coeff_"),
               names_to = "term",
               names_prefix = "delta_coeff_",
               values_to = "delta_coeff") %>% 
  filter(term != "(Intercept)",
         delta_coeff != 0) %>% 
  group_by(data_source, basin, gamma_slab) %>% 
  count(term) %>% 
  ungroup() %>% 
  pivot_wider(values_from = n, names_from = term)

# print table
lm_best_predictor_counts %>%
  gt(rowname_col = "gamma_slab",
     groupname_col = c("basin", "data_source")) %>% 
  summary_rows(
    groups = TRUE,
    fns = list(total = "sum")
  )

```


## RMSE alternatives

### AIC

AIC is an alternative criterion to RMSE to judge model quality, but not (yet) taken into account.

```{r aic_vs_rmse_all_models}
lm_all %>% 
  ggplot(aes(rmse, AIC, col = gamma_slab)) +
  geom_point() +
  scale_color_viridis_d() +
  facet_grid(data_source~basin)

```


```{r aic_vs_rmse_selected_models}
lm_best %>% 
  ggplot(aes(rmse, AIC, col = gamma_slab)) +
  geom_point() +
  scale_color_viridis_d() +
  facet_grid(data_source~basin)

```


### AIC vs BIC

BIC is an alternative criterion to RMSE to judge model quality, but not (yet) taken into account.

```{r BIC_vs_AIC_all_models}
lm_all %>% 
  ggplot(aes(AIC, BIC, col = gamma_slab)) +
  geom_point() +
  scale_color_viridis_d() +
  facet_grid(data_source~basin)

```

## RMSE vs VIF

```{r rmse_vs_VIF_all_models, fig.asp=2}

GLODAP_glanced %>%
  group_by(data_source) %>%
  group_split() %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x,
             aes(rmse, log10(vif_max), col = as.factor(n_predictors))) +
      geom_hline(yintercept = log10(params_local$vif_max)) +
      geom_point() +
      scale_color_viridis_d(name = "n_predictors") +
      labs(title = "All models",
           subtitle = paste("data_source:", unique(.x$data_source))) +
      facet_grid(gamma_slab ~ basin)
  )

```

```{r rmse_vs_VIF_clean_models, fig.asp=2}

GLODAP_glanced_clean %>%
  group_by(data_source) %>%
  group_split() %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x,
             aes(rmse, log10(vif_max), col = as.factor(n_predictors))) +
      geom_hline(yintercept = log10(params_local$vif_max)) +
      geom_point() +
      scale_color_viridis_d(name = "n_predictors") +
      labs(title = "Selected models",
           subtitle = paste("data_source:", unique(.x$data_source))) +
      facet_grid(gamma_slab ~ basin)
  )

```


## Residual patterns

### Fitted vs true

Plotted are fitted vs actual target variable values, here: `r params_local$MLR_target`

```{r fitted_vs_actual_all_property_property, fig.asp=1}

i_binwidth <- 2

p_prop_prop(df = GLODAP_augmented,
            var1 = params_local$MLR_target,
            var2 = ".fitted") +
  facet_grid(basin_AIP ~ data_source) +
  labs(title = "All models")

```

```{r fitted_vs_actual_all_histogram, fig.asp=0.5}

GLODAP_augmented %>%
  ggplot(aes(!!sym(params_local$MLR_target) - .fitted)) +
  geom_histogram(binwidth = i_binwidth) +
  labs(title = "All models") +
  scale_x_continuous(limits = c(-50,50)) +
  facet_wrap(~data_source)

```


```{r fitted_vs_actual_best_property_property, fig.asp=1}

p_prop_prop(df = GLODAP_augmented_best,
            var1 = params_local$MLR_target,
            var2 = ".fitted") +
  labs(title = "Selected models") +
  facet_grid(basin_AIP ~ data_source)


```


```{r fitted_vs_actual_best_histogram, fig.asp=0.5}

GLODAP_augmented_best %>%
  ggplot(aes(!!sym(params_local$MLR_target) - .fitted)) +
  geom_histogram(binwidth = i_binwidth) +
  labs(title = "Selected models") +
    scale_x_continuous(limits = c(-50,50)) +
  facet_wrap(~data_source)

```


### Pooled

In the following, we present residual patterns vs latitude across all domains.

```{r residual_patterns_full_range, fig.asp=0.5}

i_ylim <- c(-30,30)

GLODAP_augmented_best %>%
  ggplot(aes(lat, .resid)) +
  geom_bin2d(binwidth = i_binwidth) +
  geom_hline(yintercept = 0, col = "black") +
  scale_fill_viridis_c(trans = "log10") +
  labs(
    title = paste(
      "Target variable:",
      params_local$MLR_target,
      "| Selected models",
      "| All domains"
    )
  ) +
  facet_wrap(~data_source)


```

Due to the few large residuals, we limit the y axis range for the plots below.

```{r residual_patterns_y_range_limit, fig.asp=0.5}

GLODAP_augmented_best %>%
  ggplot(aes(lat, .resid)) +
  geom_bin2d(binwidth = i_binwidth) +
  geom_hline(yintercept = 0, col = "black") +
  scale_fill_viridis_c(trans = "log10") +
  coord_cartesian(ylim = i_ylim) +
  labs(
    title = paste(
      "Target variable:",
      params_local$MLR_target,
      "| Selected models",
      "| All domains"
    )
  ) +
  facet_wrap(~data_source)


```


### By model domain

In the following, we present residual patterns vs latitude for separate model domains, ie basins, density slabs and eras.

```{r residual_patterns_by_model_domain, fig.asp=2}

GLODAP_augmented_best %>% 
  group_split(basin, data_source) %>% 
  # head(1) %>% 
  map(~ ggplot(data = .x, aes(lat, .resid)) +
      geom_bin2d(binwidth = i_binwidth) +
      geom_hline(yintercept = 0, col = "black") +
      scale_fill_viridis_c(trans = "log10") +
      facet_grid(gamma_slab ~ era) +
      coord_cartesian(ylim = i_ylim) +
      labs(
        title = paste(
          "Target:",
          params_local$MLR_target,
          "| best models | basin:",
          unique(.x$basin),
          "| data_source:",
          unique(.x$data_source)
        )
      ))
  

```

### Latitudinal mean

```{r latitudinal_residual, fig.asp=2}

GLODAP_augmented_best <- GLODAP_augmented_best %>%
  mutate(lat_grid = as.numeric(as.character(cut(
    lat,
    seq(-90, 90, 10),
    seq(-85, 85, 10)
  ))))

lat_residual <- GLODAP_augmented_best %>%
  group_by(data_source, basin, gamma_slab, era, lat_grid) %>% 
  summarise(.resid_mean = mean(.resid)) %>% 
  ungroup()

lat_residual %>%
  group_split(basin) %>% 
  map(
  ~ggplot(data = .x,
          aes(lat_grid, .resid_mean, col=era)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0, col = "black") +
  facet_grid(gamma_slab ~ data_source) +
    labs(title = paste("Basin:", unique(.x$basin)))
  )


```



### Latitudinal offset

```{r latitudinal_residual_offset, fig.asp=2}

# calculate residual offset for adjacent eras
lat_residual_offset <- lat_residual  %>%
  select(data_source, basin, gamma_slab, era, lat_grid, .resid_mean) %>% 
  arrange(era) %>% 
  group_by(data_source, basin, gamma_slab, lat_grid) %>% 
  mutate(.resid_mean_offset = .resid_mean - lag(.resid_mean)) %>% 
  ungroup() %>% 
  drop_na()

lat_residual_offset %>%
  ggplot(aes(lat_grid, .resid_mean_offset, col=data_source)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0, col = "black") +
  facet_grid(gamma_slab ~ basin)

```

### Spatial mean

```{r spatial_residual, fig.asp=0.8}

GLODAP_augmented_best <- GLODAP_augmented_best %>%
  mutate(lon_grid = as.numeric(as.character(cut(
    lon,
    seq(-360, 500, 10),
    seq(-355, 500, 10)
  ))))

spatial_residual <- GLODAP_augmented_best %>%
  group_by(data_source, gamma_slab, era, lat_grid, lon_grid) %>% 
  summarise(.resid_mean = mean(.resid)) %>% 
  ungroup()

spatial_obs <- GLODAP_augmented_best %>%
  distinct(gamma_slab, era, lat, lon)

spatial_residual %>%
  rename(lat = lat_grid,
         lon = lon_grid) %>%
  group_split(data_source, gamma_slab) %>%
  # head(1) %>%
  map(~ p_map_dcant_slab(
    df = .x,
    var = ".resid_mean",
    col = "divergent",
    title_text = "Residual distribution",
    subtitle_text = paste(
      "Data source:",
      unique(.x$data_source),
      "| density slab:",
      unique(.x$gamma_slab)
      ),
    legend_title = paste(params_local$MLR_target, "\nbias")
  ) +
    geom_raster(data = spatial_obs,
                aes(lon, lat)) +
    facet_grid(era ~ .)
  )


```

### Spatial offset

```{r Spatial_residual_offset, fig.asp=0.5}

# calculate residual offset for adjacent eras
spatial_residual_offset <- spatial_residual  %>%
  select(data_source, gamma_slab, era, lat_grid, lon_grid, .resid_mean) %>% 
  arrange(era) %>% 
  group_by(data_source, gamma_slab, lat_grid, lon_grid) %>% 
  mutate(.resid_mean_offset = .resid_mean - lag(.resid_mean)) %>% 
  ungroup() %>% 
  drop_na()

spatial_obs <- spatial_obs %>% 
  count(lon, lat, gamma_slab) %>% 
  filter(n == 2) %>% 
  select(-n)

spatial_residual_offset %>%
  rename(lat = lat_grid,
         lon = lon_grid) %>%
  group_split(data_source, gamma_slab) %>%
  # head(1) %>%
  map(~ p_map_dcant_slab(
    df = .x,
    var = ".resid_mean_offset",
    col = "divergent",
    title_text = "Residual offset distribution (era 2 - era 1)",
    subtitle_text = paste(
      "Data source:",
      unique(.x$data_source),
      "| density slab:",
      unique(.x$gamma_slab)
      ),
    legend_title = paste(params_local$MLR_target, "\nbias offset")
  ) +
    geom_raster(data = spatial_obs,
                aes(lon, lat))
  )

```

### Predictor

```{r residual_vs_predictor}

GLODAP_augmented_long <- GLODAP_augmented %>%
  select(era, basin_AIP, gamma_slab, data_source,
         sal:params_local$MLR_target, .resid) %>% 
  pivot_longer(sal:params_local$MLR_target,
               names_to = "predictor",
               values_to = "value")


GLODAP_augmented_long %>%
  filter(gamma_slab %in% params_local$plot_slabs) %>%
  group_split(basin_AIP,
              gamma_slab,
              predictor) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x) +
      geom_hline(yintercept = 0) +
      geom_bin_2d(aes(value, .resid)) +
      facet_grid(data_source ~ era) +
      scale_fill_viridis_c(trans = "log10") +
      labs(
        title = paste("basin_AIP:", .x$basin_AIP,
                      "| gamma_slab:", .x$gamma_slab),
        x = .x$predictor
      ) +
      coord_cartesian(ylim = c(-30, 30))
  )

```



# VIF

```{r vif_distribution_n_predictors, fig.asp=2}

GLODAP_glanced %>% 
  ggplot(aes(as.factor(n),
             log10(vif_max),
             col=data_source)) + 
  geom_hline(yintercept = log10(params_local$vif_max)) +
  scale_color_brewer(palette = "Set1") +
  geom_violin() +
  facet_grid(gamma_slab~basin)
  

```

```{r vif_distribution_type_predictor, fig.asp=3.5}

GLODAP_glanced_vif %>% 
  ggplot(aes(Variables,
             log10(VIF),
             col = data_source)) + 
  geom_hline(yintercept = log10(params_local$vif_max)) +
  geom_violin() +
  scale_color_brewer(palette = "Set1") +
  facet_grid(gamma_slab~basin) +
  coord_flip()
  
```

## Write files

```{r write_diagnostic_files}


lm_best_predictor_counts %>%
  write_csv(paste(path_version_data,
                  "lm_best_predictor_counts.csv",
                  sep = ""))

lat_residual %>%
  write_csv(paste(path_version_data,
                  "lm_lat_residual.csv",
                  sep = ""))

lat_residual_offset %>%
  write_csv(paste(path_version_data,
                  "lm_lat_residual_offset.csv",
                  sep = ""))

spatial_residual %>%
  write_csv(paste(path_version_data,
                  "lm_spatial_residual.csv",
                  sep = ""))

spatial_residual_offset %>%
  write_csv(paste(path_version_data,
                  "lm_spatial_residual_offset.csv",
                  sep = ""))

```
