---
title: "GLODAPv2_2020 subsetting"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}
params_local <-
  read_rds(here::here("data/auxillary",
                       "params_local.rds"))

```

```{r define_paths, include = FALSE}

path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_preprocessing_model    <-
  paste(path_root, "/model/preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(lubridate)
library(gt)
```

```{r select_basin_mask, include=FALSE}

basinmask <- basinmask %>% 
  filter(MLR_basins == params_local$MLR_basins) %>% 
  select(-c(MLR_basins, basin_AIP))

```


# Read files

Main data source for this project is the preprocessed version of the `GLODAPv2.2020_Merged_Master_File.csv` downloaded from [glodap.info](https://www.glodap.info/){target="_blank"} in June 2020.

**CAVEAT:** This file still contains neutral densities `gamma` calculated with a preliminary method. However, this is consistent with the way gamma is currently calculated in this script and should therefore be maintained until changed on all levels.

```{r read_GLODAPv2_2020_merged_master_file}

GLODAP <-
  read_csv(
    paste(
      path_preprocessing_model,
      "GLODAPv2.2020_preprocessed_model_runA_both.csv",
      sep = ""
    )
  )

GLODAP <- GLODAP %>%
  arrange(date) %>%
  relocate(row_number)

GLODAP_CB <-
  read_csv(paste(path_preprocessing,
                 "GLODAPv2.2020_Canyon-B.csv",
                 sep = ""))

```


# Data preparation

## Reference eras

Samples were assigned to following eras:

```{r assign_eras}

# create labels for era
era_labels <- bind_cols(
  start = params_local$era_start,
  end = params_local$era_end)

era_labels <- era_labels %>% 
  mutate(start = if_else(start == -Inf, max(GLODAP$year), start),
         end = if_else(end == Inf, max(GLODAP$year), end),
         era = as.factor(paste(start, end, sep = "-")))

# filter GLODAP data within eras
GLODAP <- expand_grid(
  GLODAP,
  era_labels
)

# extract A16 cruise data from 2013/14
GLODAP_A16_backup <- GLODAP %>% 
  filter(cruise %in% c(1041, 1042),
         era == sort(unique(GLODAP$era))[2])

# select data within each era
GLODAP <- GLODAP %>% 
  filter(year >= start & year <= end)


if (params_local$A16_GO_SHIP == "y") {
  GLODAP <- bind_rows(GLODAP,
                      GLODAP_A16_backup)
  
  map +
    geom_raster(data = GLODAP_A16_backup %>% distinct(lat, lon, cruise),
                aes(lon, lat, fill = as.factor(cruise))) +
    scale_fill_brewer(palette = "Dark2") +
    labs(
      title = "Manual adjustment of A16 cruise",
      subtitle = paste(
        "For harmonization with Gruber et al. (2019),
        cruises 1041 (A16N) and 1042 (A16S) were grouped into the",
        sort(unique(GLODAP$era))[2], "era
        despite taking place in 2013/14"
      )
    ) +
    theme(legend.title = element_blank())
  
}

GLODAP <- GLODAP %>% 
  select(-c(start, end))

rm(era_labels, GLODAP_A16_backup)

```

## Spatial boundaries

### Basin mask

The basin mask from the World Ocean Atlas was used. For details consult the data base subsection for [WOA18](https://jens-daniel-mueller.github.io/Cant_eMLR/read_World_Ocean_Atlas_2018.html) data.

Please note that some GLODAP observations were made outside the WOA18 basin mask (i.e. in marginal seas) and will be removed for further analysis.

```{r join_GLODAP_basin_mask}

# use only data inside basinmask
GLODAP <- inner_join(GLODAP, basinmask)

```

### Depth

Observations collected shallower than:

- minimum sampling depth:  `r params_local$depth_min`m

were excluded from the analysis to avoid seasonal bias.

```{r apply_depth_min}

GLODAP <- GLODAP %>% 
  filter(depth >= params_local$depth_min)

```

### Bottomdepth

Observations collected in an area with a:

- minimum bottom depth:  `r params_local$bottomdepth_min`m

were excluded from the analysis to avoid coastal impacts. Please note that minimum bottom depth criterion of 0m means that no filtering was applied here.

```{r apply_bottomdepth_min}

GLODAP <- GLODAP %>% 
  filter(bottomdepth >= params_local$bottomdepth_min)

```


## Flags and missing data

Only rows (samples) for which all relevant parameters are available were selected, ie NA's were removed.

According to Olsen et al (2020), flags within the merged master file identify:

- f:  

  - 2: Acceptable
  - 0: Interpolated (nutrients/oxygen) or calculated (CO[2] variables)
  - 9: Data not used (so, only NA data should have this flag)
  
  
- qc:  

  - 1: Adjusted or unadjusted data
  - 0: Data appear of good quality but have not been subjected to full secondary QC
  - data with poor or uncertain quality are excluded.

Following flagging criteria were taken into account:

- flag_f:     `r params_local$flag_f`
- flag_qc:    `r params_local$flag_qc`

The cleaning process was performed successively and the maps below represent the data coverage at various cleaning levels. 

Summary statistics were calculated during cleaning process.

### tco2

#### NA

Rows with missing tco2 observations were already removed in the preprocessing. The map below shows the coverage of preprocessed GLODAP data.

```{r tco2_na_subset}

GLODAP_stats <- GLODAP %>% 
  summarise(tco2_values = n())

GLODAP_obs_grid <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "tco2_values")

```

```{r WOA18_basin_mask_GLODAP_observations_map, fig.asp=0.6}

GLODAP_obs <- GLODAP %>% 
  group_by(lat, lon) %>% 
  summarise(n = n()) %>% 
  ungroup()

map +
  geom_raster(data = basinmask, aes(lon, lat, fill = basin)) +
  geom_raster(data = GLODAP_obs, aes(lon, lat)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "top",
        legend.title = element_blank())

rm(GLODAP_obs)

```


#### f flag

```{r tco2_f_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, tco2f)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ tco2f) +
  theme(legend.position = "top")

rm(GLODAP_obs_grid_temp)

GLODAP <- GLODAP %>%
  filter(tco2f %in% params_local$flag_f)


```

#### qc flag

```{r tco2_qc_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, tco2qc)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ tco2qc) +
  theme(legend.position = "top")

##

GLODAP <- GLODAP %>%
  filter(tco2qc %in% params_local$flag_qc)

GLODAP_stats_temp <- GLODAP %>%
  summarise(tco2_flag = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "tco2_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### talk

#### NA

```{r talk_na_subset, fig.asp=0.6}

GLODAP <- GLODAP %>% 
  mutate(talkna = if_else(is.na(talk), "NA", "Value"))

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkna)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkna) +
  theme(legend.position = "top")

GLODAP <- GLODAP %>% 
  select(-talkna) %>% 
  filter(!is.na(talk))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(talk_values = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "talk_values")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

#### f flag

```{r talk_f_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkf)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkf) +
  theme(legend.position = "top",
        legend.title = element_blank())


# ###

GLODAP <- GLODAP %>%
  filter(talkf %in% params_local$flag_f)


```

#### qc flag

```{r talk_qc_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkqc)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkqc) +
  theme(legend.position = "top",
        legend.title = element_blank())

###

GLODAP <- GLODAP %>%
  filter(talkqc %in% params_local$flag_qc)

##

GLODAP_stats_temp <- GLODAP %>%
  summarise(talk_flag = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "talk_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### Phosphate

#### NA

```{r phosphate_na_subset, fig.asp=0.6}

GLODAP <- GLODAP %>% 
  mutate(phosphatena = if_else(is.na(phosphate), "NA", "Value"))

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphatena)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ phosphatena) +
  theme(legend.position = "top")

GLODAP <- GLODAP %>% 
  select(-phosphatena) %>% 
  filter(!is.na(phosphate))

##

GLODAP_stats_temp <- GLODAP %>% 
  summarise(phosphate_values = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "phosphate_values")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

#### f flag

```{r phosphate_f_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphatef)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~phosphatef) +
  theme(legend.position = "top",
        legend.title = element_blank())

###

GLODAP <- GLODAP %>%
  filter(phosphatef %in% params_local$flag_f)


```

#### qc flag

```{r phosphate_qc_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphateqc)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~phosphateqc) +
  theme(legend.position = "top",
        legend.title = element_blank())


###

GLODAP <- GLODAP %>%
  filter(phosphateqc %in% params_local$flag_qc)

##

GLODAP_stats_temp <- GLODAP %>%
  summarise(phosphate_flag = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)
rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "phosphate_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### eMLR variables

Variables required as predictors for the MLR fits, are subsetted for NAs and flags.

```{r eMLR_subset}

if ("temp" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(temp))
}

##

if ("sal" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(sal))
  
  GLODAP <- GLODAP %>%
    filter(salinityf %in% params_local$flag_f)
  
  GLODAP <- GLODAP %>%
    filter(salinityqc %in% params_local$flag_qc)
}

##

if ("silicate" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(silicate))
  
  GLODAP <- GLODAP %>%
    filter(silicatef %in% params_local$flag_f)
  
  GLODAP <- GLODAP %>%
    filter(silicateqc %in% params_local$flag_qc)
}

##

if ("oxygen" %in% params_local$MLR_predictors |
    "phosphate_star" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(oxygen))
  
  GLODAP <- GLODAP %>%
    filter(oxygenf %in% params_local$flag_f)
  
  GLODAP <- GLODAP %>%
    filter(oxygenqc %in% params_local$flag_qc)
}

##

if ("aou" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(aou))
  
  GLODAP <- GLODAP %>%
    filter(aouf %in% params_local$flag_f)
}

##

if ("nitrate" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(nitrate))
  
  GLODAP <- GLODAP %>%
    filter(nitratef %in% params_local$flag_f)
  
  GLODAP <- GLODAP %>%
    filter(nitrateqc %in% params_local$flag_qc)
}

##

GLODAP <- GLODAP %>%
  filter(!is.na(depth))

GLODAP <- GLODAP %>%
  filter(!is.na(gamma))

##

GLODAP_stats_temp <- GLODAP %>%
  summarise(eMLR_variables = n())

GLODAP_stats <- cbind(GLODAP_stats, GLODAP_stats_temp)

rm(GLODAP_stats_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "eMLR_variables")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

```{r remove_flag_columns}

GLODAP <- GLODAP %>% 
  select(-ends_with(c("f", "qc")))

```

## CANYON-B comparison

Cruises are removed, when the mean offset of the observation from the value predicted with CANYON-B is higher than `r params_local$CANYON_B_max` times the standard deviation of all cruise mean offsets. This critrion is evaluated individually for each variable involved in the eMLR approach.

```{r combine_GLODAP_CANYON_B}

# join data frames
GLODAP_combined <- left_join(GLODAP,
                             GLODAP_CB)

# calculate offset by parameter
GLODAP_combined <- GLODAP_combined %>%
  mutate(
    offset_talk = talk - talk_CANYONB,
    offset_tco2 = tco2 - tco2_CANYONB,
    offset_nitrate = nitrate - nitrate_CANYONB,
    offset_phosphate = phosphate - phosphate_CANYONB,
    offset_silicate = silicate - silicate_CANYONB
  ) %>%
  select(row_number,
         year,
         cruise,
         basin_AIP,
         lat,
         lon,
         starts_with("offset_"))

# pivot to long format
GLODAP_combined <- GLODAP_combined %>%
  pivot_longer(
    starts_with("offset"),
    names_to = "parameter",
    names_prefix = "offset_",
    values_to = "offset"
  )


```

### Residual histograms

```{r CANYON_B_residual_histograms, fig.asp=0.6}

GLODAP_combined %>%
  ggplot(aes(offset)) +
  geom_histogram() +
  scale_y_continuous(trans = "log10") +
  facet_grid(basin_AIP ~ parameter, scales = "free_x")

```

### Ranked residuals

```{r CANYON_B_ranked_residuals}

# calculate mean cruise offset by parameter
cruise_all <- GLODAP_combined %>%
  group_by(cruise, parameter) %>%
  summarise(
    mean_offset = mean(offset, na.rm = TRUE),
    sd_offset = sd(offset, na.rm = TRUE)
  ) %>%
  ungroup()

# rank offsets and calculate offset threshols
cruise_all <- cruise_all %>%
  group_by(parameter) %>%
  mutate(
    rank_offset = rank(mean_offset),
    threshold_offset = sd(mean_offset) * params_local$CANYON_B_max,
    cruise = as.factor(cruise)
  ) %>%
  ungroup() %>%
  arrange(parameter, rank_offset)

cruise_out <- cruise_all %>%
  filter(parameter %in% c("tco2", "talk", params_local$MLR_predictors)) %>% 
  filter(abs(mean_offset) > threshold_offset)

# length(unique(cruise_out$cruise))

for (i_parameter in unique(cruise_all$parameter)) {
  # i_parameter <- unique(cruise_all$parameter)[1]
  
  i_cruise_all <- cruise_all %>%
    filter(parameter == i_parameter)
  
  i_cruise_out <- cruise_out %>%
    filter(parameter == i_parameter)
  
  print(
  ggplot() +
    geom_hline(data = i_cruise_all,
               aes(yintercept = c(-1, 1) * threshold_offset),
               lty = 2) +
    geom_ribbon(
      data = i_cruise_all,
      aes(
        x = rank_offset,
        ymax = mean_offset + sd_offset,
        ymin = mean_offset-+sd_offset
      ),
      alpha = 0.3
    ) +
    geom_path(data = i_cruise_all,
              aes(rank_offset, mean_offset)) +
    geom_point(data = i_cruise_out,
               aes(rank_offset, mean_offset, col = cruise)) +
    labs(title = i_parameter)
  )
  
}

rm(i_parameter, i_cruise_all, i_cruise_out)

```

### Remove cruises

```{r CANYON_B_removed_cruises_map, fig.asp=1}

GLODAP_out <- GLODAP %>%
  filter(cruise %in% cruise_out$cruise)

if (nrow(GLODAP_out) > 0) {
  map +
  geom_raster(data = GLODAP_out %>% distinct(lat, lon, era),
              aes(lon, lat)) +
  facet_wrap( ~ era, ncol = 1) +
  labs(title = "Maps of removed cruises")

} else {
  print("no cruises removed")
  
}

```

The ratio (%) of removed cruises is:

```{r CANYON_B_remove_cruises}

nrow(GLODAP_out) / nrow(GLODAP) * 100

GLODAP <- GLODAP %>%
  filter(!(cruise %in% cruise_out$cruise))

```


## Create clean observations grid

Grid containing all grid cells where at least one observation remains available after cleaning.

```{r create_clean_obs_grid}

GLODAP_obs_grid_clean <- GLODAP %>% 
  count(lat, lon) %>% 
  select(-n)

```



## Write summary file
 
```{r write_clean_data_files}

GLODAP_obs_grid_clean  %>%  write_csv(paste(path_version_data,
                                            "GLODAPv2.2020_clean_obs_grid.csv",
                                            sep = ""))

# select relevant columns for further analysis
GLODAP <- GLODAP %>% 
  select(year, date, era, basin, basin_AIP, lat, lon, cruise,
         bottomdepth, depth,
         temp, sal, gamma,
         tco2, talk, phosphate,
         oxygen, aou, nitrate, silicate,
         temp_model, sal_model = so_model, gamma_model,
         tco2_model = dissic_model, talk_model, phosphate_model = po4_model,
         oxygen_model = o2_model, aou_model,
         nitrate_model = no3_model, silicate_model = si_model
         )


GLODAP  %>%  write_csv(paste(path_version_data,
                             "GLODAPv2.2020_clean.csv",
                             sep = ""))


```


# Overview plots

## Number of overservations

```{r counts}

GLODAP %>% 
  group_by(era, basin_AIP) %>% 
  count() %>% 
  ggplot(aes(basin_AIP, n, fill = era)) +
  geom_col() +
  scale_fill_brewer(palette = "Dark2")

```


## Cleaning stats

Number of observations at various steps of data cleaning.

```{r GLODAP_cleaning_stats}

GLODAP_stats_long <- GLODAP_stats %>%
  pivot_longer(1:length(GLODAP_stats),
               names_to = "parameter",
               values_to = "n")

GLODAP_stats_long <- GLODAP_stats_long %>%
  mutate(parameter = fct_reorder(parameter, n))

GLODAP_stats_long %>% 
  ggplot(aes(parameter, n/1000)) +
  geom_col() +
  coord_flip() +
  theme(axis.title.y = element_blank())

rm(GLODAP_stats_long)

```


## Assign coarse spatial grid

For the following plots, the cleaned data set was re-opened and observations were gridded spatially to intervals of:  

- 5° x 5°

```{r grid_spatially_5x5}

GLODAP <- m_grid_horizontal_coarse(GLODAP)

```


## Histogram Zonal coverage

```{r coverage_histogram_zonal}

GLODAP_histogram_lat <- GLODAP %>%
  group_by(era, lat_grid, basin_AIP) %>%
  tally() %>%
  ungroup()

GLODAP_histogram_lat %>%
  ggplot(aes(lat_grid, n, fill = era)) +
  geom_col() +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap( ~ basin_AIP) +
  coord_flip(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank())

rm(GLODAP_histogram_lat)

```

## Median years (tref)

Median years of each era (tref) were determined as:

```{r calculate_median_years}

era_median_year <- GLODAP %>%
  group_by(era) %>%
  summarise(t_ref = median(year)) %>%
  ungroup()

era_median_year

```



## Histogram temporal coverage

```{r coverage_histogram_temporal}

GLODAP_histogram_year <- GLODAP %>%
  group_by(year, basin_AIP) %>%
  tally() %>%
  ungroup()

GLODAP_histogram_year %>%
  ggplot() +
  geom_vline(xintercept = sort(params_local$era_end)[1] + 0.5) +
  geom_col(aes(year, n,
               fill = basin_AIP)) +
  geom_point(
    data = era_median_year,
    aes(t_ref, 100, shape = "Median year"),
    size = 2,
    fill = "white"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  scale_shape_manual(values = 24, name = "") +
  scale_y_continuous() +
  coord_cartesian(expand = 0) +
  theme(
    legend.position = "top",
    legend.direction = "vertical",
    legend.title = element_blank(),
    axis.title.x = element_blank()
  )

rm(GLODAP_histogram_year,
   era_median_year)

```

## Zonal temporal coverage (Hovmoeller)

```{r coverage_hovmoeller, fig.asp=1.3}

GLODAP_hovmoeller_year <- GLODAP %>%
  group_by(year, lat_grid, basin_AIP) %>%
  tally() %>%
  ungroup()

GLODAP_hovmoeller_year %>%
  ggplot(aes(year, lat_grid, fill = n)) +
  geom_tile() +
  geom_vline(xintercept = sort(params_local$era_end)[1] + 0.5) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  coord_cartesian(expand = 0) +
  facet_wrap( ~ basin_AIP, ncol = 1) +
  theme(legend.position = "top",
        axis.title.x = element_blank())

rm(GLODAP_hovmoeller_year)

```

## Coverage maps by era

### Subsetting process

The following plots show the remaining data after **individual cleaning steps**, separately for each era.

```{r coverage_maps_era_subsetting_era_cleaning_level, fig.asp=1.1}

GLODAP_obs_grid <- GLODAP_obs_grid %>%
  mutate(cleaning_level = factor(
           cleaning_level,
           unique(GLODAP_obs_grid$cleaning_level)
         ))

map +
  geom_raster(data = GLODAP_obs_grid %>%
                filter(cleaning_level == "tco2_values") %>%
                select(-cleaning_level),
              aes(lon, lat, fill = "tco2_values")) +
  geom_raster(data = GLODAP_obs_grid %>%
                filter(cleaning_level != "tco2_values"),
              aes(lon, lat, fill = "subset")) +
  scale_fill_brewer(palette = "Set1", name = "") +
  facet_grid(cleaning_level ~ era) +
  theme(legend.position = "top",
        axis.title = element_blank())

```

### Final input data

The following plots show the remaining data density in each grid cell **after all cleaning steps**, separately for each era.

```{r coverage_maps_era_clean, fig.asp=1}

GLODAP_tco2_grid <- GLODAP %>%
  count(lat, lon)

map +
  geom_bin2d(data = GLODAP,
             aes(lon, lat),
             binwidth = c(1,1)) +
  scale_fill_viridis_c(option = "magma", direction = -1, trans = "log10") +
  facet_wrap(~era, ncol = 1) +
  labs(title = "Cleaned GLODAP observations",
       subtitle = paste("Version:", params_local$Version_ID)) +
  theme(axis.title = element_blank())

ggsave(path = path_version_figures,
       filename = "data_distribution_era.png",
       height = 8,
       width = 5)

```


## Model vs observation

```{r model_obs_comparison}

variables <-
  c(
    "temp",
    "sal",
    "gamma",
    "tco2",
    "talk",
    "phosphate",
    "oxygen",
    "aou",
    "nitrate",
    "silicate"
  )

for (i_var in variables) {
  #  i_var <- variables[1]
  
  # calculate equal axis limits and binwidth
  axis_lims <- GLODAP %>%
    summarise(max_value = max(c(max(!!sym(
      i_var
    )),
    max(!!sym(
      paste0(i_var, "_model")
    )))),
    min_value = min(c(min(!!sym(
      i_var
    )),
    min(!!sym(
      paste0(i_var, "_model")
    )))))
  
  binwidth_value <- (axis_lims$max_value - axis_lims$min_value) / 40
  axis_lims <- c(axis_lims$min_value, axis_lims$max_value)
  
  # obs-model plot (season)
  print(
    ggplot(GLODAP, aes(
      x = !!sym(i_var),
      y = !!sym(paste0(i_var, "_model"))
    )) +
      geom_bin2d(binwidth = binwidth_value) +
      scale_fill_viridis_c() +
      geom_abline(slope = 1,
                  col = 'red') +
      coord_equal(xlim = axis_lims,
                  ylim = axis_lims) +
      labs(x = "observation", y = "model", title = i_var)
  )
}

```

