---
title: "eMLR - assumption testing"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
params:
  Version_ID: "v_XXX"
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}

if (params$Version_ID == "v_XXX") {
  params_local <-
    read_rds(here::here("data/auxillary", "params_local.rds"))
  
} else {
  params_local <-
    read_rds(
      file = paste(path_root, "observations",
                   params$Version_ID,
                   "data/params_local.rds",
                   sep = "/")
    )
  }

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(lubridate)
library(GGally)
library(olsrr)
library(knitr)
library(kableExtra)
library(broom)
library(corrr)
```

# Version ID

The results displayed on this site correspond to the Version_ID: `r params$Version_ID`

# Required data

Required are:

- cleaned and prepared GLODAPv2.2020 file

```{r read_GLODAP_used_for_fitting}

GLODAP <-
  read_csv(paste(path_version_data,
                 "GLODAPv2.2020_MLR_fitting_ready.csv",
                 sep = ""))

```


# Predictor correlation

The correlation between:  

- pairs of seven potential predictor variables and
- C* and seven potential predictor variables

were investigated based on:  

- property-property plots and
- calculated correlation coeffcients.

## Correlation plots

For an overview, a random subset of data from all eras was plotted separately 
for both basins, with color indicating neutral density slabs (high density = dark-purple color).

```{r predictor_correlation_per_basin, fig.asp=1}

plot_fun <- function(df){
  ggpairs(
    data = df,
    columns = c(params_local$MLR_target,
                params_local$MLR_predictors),
    upper = "blank",
    ggplot2::aes(col = gamma_slab, fill = gamma_slab, alpha = 0.01)
  ) +
    scale_fill_viridis_d(direction = -1) +
    scale_color_viridis_d(direction = -1) +
    labs(title = paste(
      "Basin:",
      unique(df$basin),
      "| era: all | random subsample:",
      nrow(df),
      "| data source:",
      unique(df$data_source)
    ))
}

GLODAP %>%
  slice_sample(n = 5e3) %>% 
  group_split(basin, data_source) %>%
  # head(1) %>%
  map(plot_fun)

GLODAP <- GLODAP %>%
  filter(data_source == "obs") %>%
  select(-data_source)

```

Individual correlation plots for each basin, era and neutral density (gamma) slab are available at:

`r paste(path_version_figures, "Observations_correlation/", sep = "")`

```{r predictor_correlation_per_basin_era_slab}

if (params_local$plot_all_figures == "y") {

for (i_basin in unique(GLODAP$basin)) {
  for (i_era in unique(GLODAP$era)) {
    # i_basin <- unique(GLODAP$basin)[1]
    # i_era   <- unique(GLODAP$era)[1]

    GLODAP_basin_era <- GLODAP %>%
      filter(basin == i_basin,
             era == i_era)
    
    for (i_gamma_slab in unique(GLODAP_basin_era$gamma_slab)) {
      # i_gamma_slab <- unique(GLODAP_basin_era$gamma_slab)[5]

      GLODAP_highlight <- GLODAP_basin_era %>%
        mutate(gamma_highlight = if_else(gamma_slab == i_gamma_slab,
                                         "in", "out")) %>%
        arrange(desc(gamma_highlight))
      
      p <- GLODAP_highlight %>%
        ggpairs(
          columns = c(params_local$MLR_target,
                      params_local$MLR_predictors),
          ggplot2::aes(
            col = gamma_highlight,
            fill = gamma_highlight,
            alpha = 0.01
          )
        ) +
        scale_fill_manual(values = c("red", "grey")) +
        scale_color_manual(values = c("red", "grey")) +
        labs(
          title = paste(
            i_era,
            "|",
            i_basin,
            "| Gamma slab",
            i_gamma_slab,
            "| # obs total",
            nrow(GLODAP_basin_era),
            "| # obs slab",
            nrow(GLODAP_highlight %>%
                   filter(gamma_highlight == "in"))
          )
        )
      
      png(
        filename = paste(
          path_version_figures,
          "Observations_correlation/",
          paste(
            "Predictor_correlation",
            i_era,
            i_basin,
            i_gamma_slab,
            ".png",
            sep = "_"
          ),
          sep = ""),
        width = 12,
        height = 12,
        units = "in",
        res = 300
      )
      
      print(p)
      
      dev.off()
      
    }
  }
}

}

```

## Correlation assesment

### Calculation of correlation coeffcients

Correlation coefficients were calculated individually within each slabs, era and basin.

```{r calculate_correlation_coefficients, results='hide'}

for (i_basin in unique(GLODAP$basin)) {
  for (i_era in unique(GLODAP$era)) {
    # i_basin <- unique(GLODAP$basin)[1]
    # i_era   <- unique(GLODAP$era)[1]
    
    GLODAP_basin_era <- GLODAP %>%
      filter(basin == i_basin,
             era == i_era) %>%
      select(basin,
             era,
             gamma_slab,
             params_local$MLR_target,
             params_local$MLR_predictors)
    
    for (i_gamma_slab in unique(GLODAP_basin_era$gamma_slab)) {
      # i_gamma_slab <- unique(GLODAP_basin_era$gamma_slab)[5]
      print(i_gamma_slab)
      
      GLODAP_basin_era_slab <- GLODAP_basin_era %>%
        filter(gamma_slab == i_gamma_slab)
      
      # calculate correlation table
      cor_target_predictor_temp <- GLODAP_basin_era_slab %>%
        select(-c(basin, era, gamma_slab)) %>%
        correlate() %>%
        focus(params_local$MLR_target) %>%
        mutate(basin = i_basin,
               era = i_era,
               gamma_slab = i_gamma_slab)
      
      if (exists("cor_target_predictor")) {
        cor_target_predictor <-
          bind_rows(cor_target_predictor, cor_target_predictor_temp)
      }
      
      if (!exists("cor_target_predictor")) {
        cor_target_predictor <- cor_target_predictor_temp
      }
      
      
      cor_predictors_temp <- GLODAP_basin_era_slab %>%
        select(-c(basin, era, gamma_slab)) %>%
        correlate() %>%
        shave %>%
        stretch() %>%
        filter(!is.na(r),
               x != params_local$MLR_target,
               y != params_local$MLR_target) %>%
        mutate(pair = paste(x, y, sep = " + ")) %>%
        select(-c(x, y)) %>%
        mutate(basin = i_basin,
               era = i_era,
               gamma_slab = i_gamma_slab)
      
      if (exists("cor_predictors")) {
        cor_predictors <- bind_rows(cor_predictors, cor_predictors_temp)
      }
      
      if (!exists("cor_predictors")) {
        cor_predictors <- cor_predictors_temp
      }
      
      
      
    }
  }
}

rm(cor_predictors_temp, cor_target_predictor_temp,
   i_gamma_slab, i_era, i_basin,
   GLODAP_basin_era, GLODAP_basin_era_slab)

```

### Predictor pairs

Below, the range of correlations coefficients for each predictor pair is plotted per basin (facet) and density slab (color). Note that the range indicates the min and max values of in total 3 calculated coefficients (one per era).

```{r correlation_predictor_pairs, fig.asp=1.8}

# calculate min, max, mean across all eras
cor_predictors_stats <- cor_predictors %>% 
  group_by(pair, basin, gamma_slab) %>% 
  summarise(mean_r = mean(r),
            min_r = min(r),
            max_r = max(r)) %>% 
  ungroup()

# plot figure
cor_predictors_stats %>% 
  mutate(pair = reorder(pair, mean_r)) %>%
  ggplot() +
  geom_vline(xintercept = c(-0.9, 0.9), col = "red") +
  geom_vline(xintercept = 0) +
  geom_linerange(
    aes(y = pair, xmin = min_r, xmax = max_r, col = gamma_slab),
    position = position_dodge(width = 0.6)) +
  facet_wrap(~basin) +
  scale_color_viridis_d(direction = -1) +
  labs(x = "correlation coefficient", y = "") +
  theme(legend.position = "top")

# print table
kable(cor_predictors_stats) %>%
  add_header_above() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")

rm(cor_predictors, cor_predictors_stats)

```

### C* vs individual predictors

Below, the range of correlations coefficients for C* with each predictor is plotted per basin (facet) and density slab (color). Note that the range indicates the min and max values of in total 3 calculated coefficients (one per era).

```{r correlation_predictors, fig.asp=1.5}

cor_target_predictor <- cor_target_predictor %>% 
  rename(predictor = term)

# calculate min, max, mean across all eras
cor_target_predictor_stats <- cor_target_predictor %>%
  select(-era) %>% 
  group_by(predictor, basin, gamma_slab) %>% 
  summarise_all(list(mean_r = mean, min_r = min, max_r = max)) %>% 
  ungroup()

# plot figure
cor_target_predictor_stats %>% 
  mutate(predictor = reorder(predictor, mean_r)) %>%
  ggplot() +
  geom_vline(xintercept = c(-0.9, 0.9), col = "red") +
  geom_vline(xintercept = 0) +
  geom_linerange(
    aes(y = predictor, xmin = min_r, xmax = max_r, col = gamma_slab),
    position = position_dodge(width = 0.6)) +
  facet_wrap(~basin) +
  scale_color_viridis_d(direction = -1) +
  labs(x = "correlation coefficient", y = "C* correlation with...") +
  theme(legend.position = "top")

# print table
kable(cor_target_predictor_stats) %>%
  add_header_above() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")

rm(cor_target_predictor, cor_target_predictor_stats)

```

