---
title: "eMLR data preparation"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}
params_local <-
  read_rds(here::here("data/auxillary",
                       "params_local.rds"))

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(marelac)
library(lubridate)
library(broom)
```

# Required data

Required are: 

- GLODAPv2.2020
  - cleaned data file
- Cant from Sabine 2004 (S04)
- Cant from Gruber 2019 (G19)
- annual mean atmospheric pCO~2~ 

```{r read_required_data_sets}

GLODAP <-
  read_csv(paste(path_version_data,
                 "GLODAPv2.2020_clean.csv",
                 sep = ""))

S04_cant_3d <-
  read_csv(paste(path_preprocessing,
                 "S04_cant_3d.csv",
                 sep = ""))

G19_cant_3d <-
  read_csv(paste(path_preprocessing,
                 "G19_cant_3d.csv",
                 sep = ""))

co2_atm <-
  read_csv(paste(path_preprocessing,
                 "co2_atm.csv",
                 sep = ""))


```

# PO~4~*

## Calculation 

The predictor PO~4~* was be calculated according to Clement and Gruber (2018), ie based on oxygen. Please note that an erroneous equations for PO4* calculation is given in the supplement of Gruber et al (2019), based on nitrate.  

Here we use following equation:

```{r calculate_phosphate_star}

print(b_phosphate_star)

if ("phosphate_star" %in% params_local$MLR_predictors) {
GLODAP <- GLODAP %>% 
  mutate(phosphate_star = b_phosphate_star(phosphate, oxygen))
}

```



# C*

C* serves as a conservative tracer of anthropogenic CO~2~ uptake. It is derived from measured DIC by removing the impact of  

- organic matter formation and respiration
- calcification and calcium carbonate dissolution

Contributions of those processes are estimated from phosphate and alkalinity concentrations.

## Stoichiometric ratios

The stoichiometric nutrient ratios for the production and mineralization of organic matter were set to:

- C/P: `r params_local$rCP`
- N/P: `r params_local$rNP`

## Calculation

C* was calculated as:  

```{r calculate_cstar}

print(b_cstar)

GLODAP <- GLODAP %>% 
  mutate(rCP_phosphate = -params_local$rCP * phosphate,
         talk_05 = -0.5 * talk,
         rNP_phosphate_05 = -0.5 * params_local$rNP * phosphate,
         cstar = b_cstar(tco2, phosphate, talk))

```

## Reference year adjustment

To adjust C* values to the reference year of each observation period, we assume a transient steady state change of cant between the time of sampling the reference year. The adjustment requires an approximation of the cant concentration at the reference year. We approximate this concentration by adding the delta cant signal estimated by Gruber et al (2019) to the "base line" total cant concentration determined for 1994 by Sabine et al (2004):  

Cant(tref) = S04 + (tref-1994)/13 * G19

This way, we use exactly S04+G19 for tref=2007. For all other tref we scale Cant with the observed anomalous change over the 1994-2007 period, rather than assuming a transient steady state. However, one assumes a linear behaviour of the anomalous change over time, which might be wrong in particular for the years past 2007.


### Join Cant fields

Join Cant fields of G19 and S04

```{r join_cant_estimates, fig.asp=0.5}

G19_cant_3d <- G19_cant_3d %>% 
  select(lon, lat, depth, cant_pos_G19 = cant_pos)

S04_cant_3d <- S04_cant_3d %>% 
  select(lon, lat, depth, cant_pos_S04 = cant_pos)

cant_3d_coverage <- full_join(
  S04_cant_3d %>% distinct(lat, lon),
  G19_cant_3d %>% distinct(lat, lon)
)

cant_3d_coverage <- full_join(
  cant_3d_coverage,
  G19_cant_3d %>% distinct(lat, lon) %>% mutate(G19 = "y")
)

cant_3d_coverage <- full_join(
  cant_3d_coverage,
  S04_cant_3d %>% distinct(lat, lon) %>% mutate(S04 = "y")
)

cant_3d_coverage <- cant_3d_coverage %>% 
  mutate(coverage = case_when(
    G19 == "y" & S04 == "y" ~ "both",
    is.na(G19) & S04 == "y" ~ "S04",
    G19 == "y" & is.na(S04) ~ "G19"))

map +
  geom_raster(data = cant_3d_coverage,
              aes(lon, lat, fill = coverage)) +
  geom_raster(data = GLODAP %>% distinct(lat, lon),
              aes(lon, lat)) +
  scale_fill_brewer(palette = "Dark2")

rm(cant_3d_coverage)

cant_3d <- full_join(S04_cant_3d, G19_cant_3d)

cant_3d <- cant_3d %>%
  mutate(cant_pos_S04 = replace_na(cant_pos_S04, 0),
         cant_pos_G19 = replace_na(cant_pos_G19, 0))

# cant_3d %>%
#   filter_all(any_vars(is.na(.)))

```

### Cant at tref

Calculate Cant at tref by adding G19, scaled for the time since 1994.

```{r calculate_cant_tref}

# calculate reference year
tref <- GLODAP %>%
  group_by(era) %>%
  summarise(year = median(year)) %>%
  ungroup()

# join cant with tref
cant_3d <- expand_grid(cant_3d, tref)

# calculate cant fields for all tref
cant_3d <- cant_3d %>%
  mutate(cant_pos =
           cant_pos_S04 +
           ((year - 1994) / 13 * cant_pos_G19))

# remove columns
cant_3d <- cant_3d %>% 
  select(lon, lat, depth, era, cant_pos)

# cant_3d %>%
#   filter_all(any_vars(is.na(.)))

```


### Combine GLODAP + Cant

```{r merge_cstar_cant, fig.asp=0.8}

# observations grid per era
GLODAP_obs_grid_era <- GLODAP %>% 
  distinct(lat, lon, era)

# cant data at observations grid
cant_3d_obs <- left_join(
  GLODAP_obs_grid_era,
  cant_3d)

cant_3d_obs <- cant_3d_obs %>%
  mutate(cant_pos = replace_na(cant_pos, 0),
         depth = replace_na(depth, 0))

# cant_3d_obs %>%
#   filter_all(any_vars(is.na(.)))

# calculate number of cant data points per grid cell
cant_3d_obs <- cant_3d_obs %>%
  group_by(lon, lat, era) %>% 
  mutate(n = n(),
         n_group = if_else(n > 1, "n > 1", "n <= 1")) %>% 
  ungroup()

# GLODAP observations with only one Cant value
map +
  geom_raster(data = cant_3d_obs,
             aes(lon, lat, fill = n_group)) +
  scale_fill_brewer(palette = "Set1", name="n") +
  facet_wrap( ~ era, ncol = 1) +
  labs(title = "Number of Cant depth levels",
       subtitle = "available per latxlon grid cell")


cant_3d_obs <- cant_3d_obs %>% 
  select(-n_group)

rm(cant_3d, GLODAP_obs_grid_era)

GLODAP_cant_obs <- full_join(GLODAP, cant_3d_obs)

rm(GLODAP, cant_3d_obs)

# fill number of cant data points per grid cell to all observations
GLODAP_cant_obs <- GLODAP_cant_obs %>%
  group_by(lon, lat, era) %>% 
  fill(n, .direction = "updown") %>% 
  ungroup()

```

The mapped Cant product was merged with GLODAP observation by:

- using an identical 1x1° horizontal grid
- linear interpolation of C~ant~ from standard to sampling depth

```{r cant_interpolation_to_observations}

# interpolate cant to observation depth
GLODAP_cant_obs_int <- GLODAP_cant_obs %>%
  filter(n > 1) %>% 
  group_by(lat, lon, era) %>%
  arrange(depth) %>%
  mutate(cant_pos_int = approxfun(depth, cant_pos, rule = 2)(depth)) %>%
  ungroup()

# GLODAP_cant_obs_int %>% 
#   filter(is.na(cant_pos_int))

# set cant for observation depth if only one cant available
GLODAP_cant_obs_set <- GLODAP_cant_obs %>%
  filter(n == 1) %>%
  group_by(lat, lon, era) %>%
  mutate(cant_pos_int = mean(cant_pos, na.rm = TRUE)) %>%
  ungroup()

# GLODAP_cant_obs_set %>% 
#   filter(is.na(cant_pos_int))

# bin data sets with interpolated and set cant
GLODAP_cant_obs <- bind_rows(GLODAP_cant_obs_int, GLODAP_cant_obs_set)
rm(GLODAP_cant_obs_int, GLODAP_cant_obs_set)


# ggplot() +
#   geom_path(
#     data = GLODAP_cant_obs %>%
#       filter(lat == 48.5, lon == 165.5,!is.na(cant_pos)) %>%
#       arrange(depth),
#     aes(cant_pos, depth, col = "mapped")
#   ) +
#   geom_point(
#     data = GLODAP_cant_obs %>%
#       filter(lat == 48.5, lon == 165.5,!is.na(cant_pos)) %>%
#       arrange(depth),
#     aes(cant_pos, depth, col = "mapped")
#   ) +
#   geom_point(
#     data = GLODAP_cant_obs %>%
#       filter(lat == 48.5, lon == 165.5, date == ymd("2018-06-27")),
#     aes(cant_pos_int, depth, col = "interpolated")
#   ) +
#   scale_y_reverse() +
#   facet_wrap(~era) +
#   scale_color_brewer(palette = "Dark2", name = "") +
#   labs(title = "Cant interpolation to sampling depth - example profile")

# remove cant data at grid cells without observations
GLODAP <- GLODAP_cant_obs %>%
  filter(!is.na(cstar)) %>%
  mutate(cant_pos = cant_pos_int) %>%
  select(-cant_pos_int, n)

rm(GLODAP_cant_obs)


# GLODAP %>%
#   filter_all(any_vars(is.na(.)))

```


### Merge GLODAP + atm. pCO~2~

GLODAP observations were merged with mean annual atmospheric pCO~2~ levels by year.

```{r merge_cstar_pCO2}

GLODAP <- left_join(GLODAP, co2_atm)

```

### Calculation

```{r adjust_reference_year}

# assign reference year
GLODAP <- GLODAP %>% 
  group_by(era) %>% 
  mutate(tref = median(year)) %>% 
  ungroup()

# extract atm pCO2 at reference year
co2_atm_tref <- right_join(co2_atm, tref) %>% 
  select(-year) %>% 
  rename(pCO2_tref = pCO2)

# merge atm pCO2 at tref with GLODAP
GLODAP <- full_join(GLODAP, co2_atm_tref)
rm(co2_atm, tref)

# calculate cstar for reference year
GLODAP <- GLODAP %>%
  mutate(
    cstar_tref_delta =
      ((pCO2 - pCO2_tref) / (pCO2_tref - params_local$preind_atm_pCO2)) * cant_pos,
    cstar_tref = cstar - cstar_tref_delta)

```


## Control plots

```{r cstar_reference_year_adjustment_histogram}

GLODAP %>% 
  ggplot(aes(cstar_tref_delta)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Histogramm with binwidth = 1")

```

```{r cstar_reference_year_adjustment_vs_time_cant}

GLODAP %>% 
  sample_n(1e4) %>% 
  ggplot(aes(year, cstar_tref_delta, col = cant_pos)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(title = "Time series of random subsample 1e4")


```

```{r cstar_reference_year_adjustment_vs_time_counts}

GLODAP %>% 
  ggplot(aes(year, cstar_tref_delta)) +
  geom_bin2d(binwidth = 1) +
  scale_fill_viridis_c(trans = "log10") +
  labs(title = "Heatmap with binwidth = 1")

```

# Selected section plots

A selected section is plotted to demonstrate the magnitude of various parameters and corrections relevant to C*.

```{r select_cruises}

GLODAP_cruise <- GLODAP %>% 
  filter(cruise %in% params_global$cruises_meridional)

```


```{r meridional_sections, fig.asp=0.5}

map +
  geom_path(data = GLODAP_cruise %>%
              arrange(date),
            aes(lon, lat)) +
  geom_point(data = GLODAP_cruise %>%
              arrange(date),
             aes(lon, lat, col = date)) +
  scale_color_viridis_c(trans = "date") +
  labs(title = paste("Cruise year:", mean(GLODAP_cruise$year)))

lat_section <- 
GLODAP_cruise %>%
  ggplot(aes(lat, depth)) +
  scale_y_reverse() +
  scale_fill_viridis_c() +
  theme(axis.title.x = element_blank())

for (i_var in c("tco2",
                "rCP_phosphate",
                "talk_05",
                "rNP_phosphate_05",
                "cstar",
                "cstar_tref")) {
  print(lat_section +
          stat_summary_2d(aes(z = !!sym(i_var))) +
          scale_fill_viridis_c(name = i_var)
        )
  
}

rm(lat_section, GLODAP_cruise)

```


# Isoneutral slabs

The following boundaries for isoneutral slabs were defined:

- Atlantic: `r params_local$slabs_Atl`
- Indo-Pacific: `r params_local$slabs_Ind_Pac`

Continuous neutral densities (gamma) values from GLODAP are grouped into isoneutral slabs.

```{r cut_isoneutral_slabs}

GLODAP <- m_cut_gamma(GLODAP, "gamma")

```

```{r meridional_sections_isoneutral_slabs, fig.asp=0.6}

GLODAP_cruise <- GLODAP %>% 
  filter(cruise %in% params_global$cruises_meridional)

lat_section <- 
GLODAP_cruise %>%
  ggplot(aes(lat, depth)) +
  scale_y_reverse() +
  theme(legend.position = "bottom")

lat_section +
  geom_point(aes(col = gamma_slab)) +
  scale_color_viridis_d()

rm(lat_section, GLODAP_cruise)

```


```{r test_isoneutral_density_calculation, eval=FALSE, hide=TRUE}

# this section was only used to calculate gamma locally, and compare it to the value provided in GLODAP data set

GLODAP_cruise <- GLODAP %>% 
  filter(cruise %in% params_global$cruises_meridional)

library(oce)
library(gsw)
# calculate pressure from depth

GLODAP_cruise <- GLODAP_cruise %>% 
  mutate(CTDPRS = gsw_p_from_z(-depth,
                               lat))

GLODAP_cruise <- GLODAP_cruise %>% 
  mutate(THETA = swTheta(salinity = sal,
                         temperature = temp,
                         pressure = CTDPRS,
                         referencePressure = 0,
                         longitude = lon-180,
                         latitude = lat))

GLODAP_cruise <- GLODAP_cruise %>% 
  rename(LATITUDE = lat,
         LONGITUDE = lon,
         SALNTY = sal,
         gamma_provided = gamma)

library(reticulate)
source_python(here::here("code/python_scripts",
                         "Gamma_GLODAP_python.py"))

GLODAP_cruise <- calculate_gamma(GLODAP_cruise)

GLODAP_cruise <- GLODAP_cruise %>% 
  mutate(gamma_delta = gamma_provided - GAMMA)

lat_section <- 
GLODAP_cruise %>%
  ggplot(aes(LATITUDE, CTDPRS)) +
  scale_y_reverse() +
  theme(legend.position = "bottom")

lat_section +
  stat_summary_2d(aes(z = gamma_delta)) +
  scale_color_viridis_c()

GLODAP_cruise %>% 
  ggplot(aes(gamma_delta))+
  geom_histogram()

rm(lat_section, GLODAP_cruise, cruises_meridional)

```


# Observations coverage

```{r obs_gamma_slab, fig.asp=0.8}

GLODAP <- GLODAP %>% 
  mutate(gamma_slab = factor(gamma_slab), 
         gamma_slab = factor(gamma_slab, levels = rev(levels(gamma_slab))))

for (i_basin in unique(GLODAP$basin)) {
  # i_basin <- unique(GLODAP$basin)[3]
  
  print(
    GLODAP %>%
      filter(basin == i_basin) %>%
      ggplot(aes(lat, gamma_slab)) +
      geom_bin2d(binwidth = 5) +
      scale_fill_viridis_c(
        option = "magma",
        direction = -1,
        trans = "log10"
      ) +
      scale_x_continuous(breaks = seq(-100, 100, 20),
                         limits = c(params_global$lat_min,
                                    params_global$lat_max)) +
      facet_grid(era ~ .) +
      labs(title = paste("MLR region: ", i_basin))
  )
  
}

```


## Histograms

```{r histogram_predictor, fig.asp=0.8}

GLODAP_vars <- GLODAP %>% 
  select(params_local$MLR_target,
         params_local$MLR_predictors)

GLODAP_vars_long <- GLODAP_vars %>%
  pivot_longer(
    cols = c(params_local$MLR_target,
             params_local$MLR_predictors),
    names_to = "variable",
    values_to = "value"
  )

GLODAP_vars_long %>% 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~ variable,
             ncol = 2,
             scales = "free")

rm(GLODAP_vars, GLODAP_vars_long)

```




# Cruise quality check

## MLR models

The the following full MLR model was fitted to all GLODAP, irrespective of the sampling era:

```{r fit_MLR_to_all_data}

#define full model
model <- paste("cstar",
               paste(params_local$MLR_predictors, collapse = " + "),
               sep = " ~ ")

model

# prepare nested data frame
GLODAP_nested <- GLODAP %>%
  group_by(gamma_slab, basin) %>%
  nest()

# expand with model definitions
GLODAP_nested_lm <- expand_grid(GLODAP_nested,
                                model)

# fit models and extract tidy model output
GLODAP_nested_lm_fit <- GLODAP_nested_lm %>%
  mutate(
    fit = map2(.x = data, .y = model,
               ~ lm(as.formula(.y), data = .x)),
    tidied = map(fit, tidy),
    glanced = map(fit, glance),
    augmented = map(fit, augment)
  )

# extract augmented model output (fitted values and residuals)
GLODAP_augmented <- GLODAP_nested_lm_fit %>%
  select(-c(data, fit, tidied, glanced)) %>%
  unnest(augmented)

# extract input data
GLODAP_data <- GLODAP_nested_lm_fit %>%
  select(-c(fit, tidied, glanced, augmented)) %>%
  unnest(data)

# append input data with augmented data
GLODAP_augmented <- bind_cols(GLODAP_data,
                              GLODAP_augmented %>% select(.fitted, .resid))

rm(GLODAP_data)
```

### Residuals

Below, the residuals of C* from the mean C* and from C* predicted with the global model are shown.

```{r Residuals_global_c_star, fig.asp=0.5}

GLODAP_augmented %>%
  ggplot(aes(year, cstar_tref - mean(cstar_tref))) +
  geom_bin2d(binwidth = c(1, 1)) +
  scale_fill_viridis_c()

GLODAP_augmented %>%
  ggplot(aes(year, .resid)) +
  geom_bin2d(binwidth = c(1, 1)) +
  scale_fill_viridis_c()


```

### Ranked RMSE

```{r ranked_cstar_rmse, fig.asp=0.5}


# calculate RMSE from augmented output per cruise
cruise_all <- GLODAP_augmented %>%
  group_by(cruise) %>%
  summarise(rmse = sqrt(c(crossprod(.resid)) / length(.resid))) %>%
  ungroup()

# rank RMSE
cruise_all <- cruise_all %>% 
  mutate(rank_rmse = rank(rmse),
         cruise = as.factor(cruise))

cruise_out <- cruise_all %>% 
  filter(rmse > params_local$c_star_rmse_max)

GLODAP_out <- GLODAP_augmented %>% 
  filter(cruise %in% cruise_out$cruise)

ggplot() +
  geom_hline(yintercept = params_local$c_star_rmse_max) +
  geom_point(data = cruise_all,
             aes(rank_rmse, rmse)) +
  geom_point(data = cruise_out,
             aes(rank_rmse, rmse, col = cruise))


```

### Removed cruises

Following fraction (%) of cruises was removed:

```{r cstar_rmse_removed_cruises, fig.asp=0.8}

nrow(GLODAP_out)/nrow(GLODAP_augmented)*100

if (nrow(GLODAP_out) > 0) {
  map +
  geom_raster(data = GLODAP_out %>% distinct(lat, lon, era),
              aes(lon, lat)) +
  facet_wrap( ~ era, ncol = 1) +
  labs(title = "Maps of removed cruises")

} else {
  print("no cruises removed")
  
}


```



## Individual cruise sections

Zonal and meridional section plots are produce for each cruise individually and are available under:  

`r paste(path_version_figures, "Cruise_sections_histograms/", sep = "")`

```{r plot_all_individual_cruises_clean}

if (params_local$plot_all_figures == "y") {

cruises <- GLODAP %>% 
  group_by(cruise) %>% 
  summarise(date_mean = mean(date, na.rm = TRUE),
            n = n()) %>% 
  ungroup() %>% 
  arrange(date_mean)

GLODAP <- full_join(GLODAP, cruises)

n <- 0
for (i_cruise in unique(cruises$cruise)) {

# i_cruise <- unique(cruises$cruise)[1]
# n <- n + 1
# print(n)  
  
GLODAP_cruise <- GLODAP %>%
  filter(cruise == i_cruise) %>% 
  arrange(date)

cruises_cruise <- cruises %>%
  filter(cruise == i_cruise)
  
map_plot <- 
  map +
  geom_point(data = GLODAP_cruise,
             aes(lon, lat, col = date)) +
  scale_color_viridis_c(trans = "date") +
  labs(title = paste("Mean date:", cruises_cruise$date_mean,
                     "| cruise:", cruises_cruise$cruise,
                     "| n(samples):", cruises_cruise$n))


lon_section <- GLODAP_cruise %>%
  ggplot(aes(lon, depth)) +
  scale_y_reverse() +
  scale_fill_viridis_c()

lon_tco2 <- lon_section+
  stat_summary_2d(aes(z=tco2))

lon_talk <- lon_section+
  stat_summary_2d(aes(z=talk))

lon_phosphate <- lon_section+
  stat_summary_2d(aes(z=phosphate))

lon_oxygen <- lon_section+
  stat_summary_2d(aes(z=oxygen))

lon_aou <- lon_section+
  stat_summary_2d(aes(z=aou))

lon_phosphate_star <- lon_section+
  stat_summary_2d(aes(z=phosphate_star))

lon_nitrate <- lon_section+
  stat_summary_2d(aes(z=nitrate))

lon_cstar <- lon_section+
  stat_summary_2d(aes(z=cstar_tref))


lat_section <- GLODAP_cruise %>%
  ggplot(aes(lat, depth)) +
  scale_y_reverse() +
  scale_fill_viridis_c()

lat_tco2 <- lat_section+
  stat_summary_2d(aes(z=tco2))

lat_talk <- lat_section+
  stat_summary_2d(aes(z=talk))

lat_phosphate <- lat_section+
  stat_summary_2d(aes(z=phosphate))

lat_oxygen <- lat_section+
  stat_summary_2d(aes(z=oxygen))

lat_aou <- lat_section+
  stat_summary_2d(aes(z=aou))

lat_phosphate_star <- lat_section+
  stat_summary_2d(aes(z=phosphate_star))

lat_nitrate <- lat_section+
  stat_summary_2d(aes(z=nitrate))

lat_cstar <- lat_section+
  stat_summary_2d(aes(z=cstar_tref))

hist_tco2 <- GLODAP_cruise %>%
  ggplot(aes(tco2)) +
  geom_histogram()

hist_talk <- GLODAP_cruise %>%
  ggplot(aes(talk)) +
  geom_histogram()

hist_phosphate <- GLODAP_cruise %>%
  ggplot(aes(phosphate)) +
  geom_histogram()

hist_oxygen <- GLODAP_cruise %>%
  ggplot(aes(oxygen)) +
  geom_histogram()

hist_aou <- GLODAP_cruise %>%
  ggplot(aes(aou)) +
  geom_histogram()

hist_phosphate_star <- GLODAP_cruise %>%
  ggplot(aes(phosphate_star)) +
  geom_histogram()

hist_nitrate <- GLODAP_cruise %>%
  ggplot(aes(nitrate)) +
  geom_histogram()

hist_cstar <- GLODAP_cruise %>%
  ggplot(aes(cstar_tref)) +
  geom_histogram()

(map_plot /
    ((hist_tco2 / hist_talk / hist_phosphate / hist_cstar) |
       (hist_oxygen / hist_phosphate_star / hist_nitrate / hist_aou)
    )) |
  ((lat_tco2 / lat_talk / lat_phosphate / lat_oxygen / lat_aou / lat_phosphate_star / lat_nitrate / lat_cstar) |
     (lon_tco2 / lon_talk / lon_phosphate / lon_oxygen /  lon_aou /lon_phosphate_star / lon_nitrate / lon_cstar))    

ggsave(
  path = paste(path_version_figures, "Cruise_sections_histograms/", sep = ""),
  filename = paste(
    "Cruise_date",
    cruises_cruise$date_mean,
    "count",
    cruises_cruise$n,
    "cruiseID",
    cruises_cruise$cruise,
    ".png",
    sep = "_"
  ),
width = 20, height = 12)

rm(map_plot,
   lon_section, lat_section,
   lat_tco2, lat_talk, lat_phosphate, lon_tco2, lon_talk, lon_phosphate,
   GLODAP_cruise, cruises_cruise)

}

}

```


# Write files

```{r write_csv_GLODAP_used_for_fitting}

# select relevant columns
GLODAP <- GLODAP %>%
  filter(!(cruise %in% cruise_out$cruise)) %>%
  select(
    year,
    date,
    era,
    basin,
    basin_AIP,
    lat,
    lon,
    depth,
    gamma,
    gamma_slab,
    params_local$MLR_predictors,
    params_local$MLR_target
  )

GLODAP %>% write_csv(paste(
  path_version_data,
  "GLODAPv2.2020_MLR_fitting_ready.csv",
  sep = ""
))

co2_atm_tref %>%  write_csv(paste(path_version_data,
                                  "co2_atm_tref.csv",
                                  sep = ""))

```

# Spatial autocorrelation

```{r spatial_autocorrelation, eval=FALSE}

GLODAP_sp <- GLODAP %>% 
  filter(depth == 150)

map +
  geom_raster(data = GLODAP_sp,
              aes(lon, lat, fill = temp)) +
  scale_fill_viridis_c()

class(GLODAP_sp)

GLODAP_sp <- GLODAP_sp %>% 
  mutate(lon = if_else(lon > 180, lon - 360, lon))

ggplot() +
  geom_raster(data = GLODAP_sp,
              aes(lon, lat, fill = temp)) +
  scale_fill_viridis_c() +
  coord_quickmap()

GLODAP_sp <- as.data.frame(GLODAP_sp)

library(sp)
coordinates(GLODAP_sp) = ~lon+lat
class(GLODAP_sp)

summary(GLODAP_sp)

is.projected(GLODAP_sp)
proj4string(GLODAP_sp) <-
  CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")


GLODAP_sp_grid <- GLODAP_sp

gridded(GLODAP_sp_grid) <- TRUE


spplot(GLODAP_sp,
       zcol = "temp")

spplot(GLODAP_sp_grid,
       zcol = "temp")

library(sf)
library(stars)
GLODAP_sf <- st_as_sf(GLODAP_sp_grid)
GLODAP_stars <- st_as_stars(GLODAP_sp_grid)
class(GLODAP_stars)

plot(GLODAP_stars)

ggplot() +
  geom_stars(data = GLODAP_stars,
             aes(x, y, fill = temp)) +
  scale_fill_viridis_c(na.value = "transparent") +
  coord_quickmap(expand = 0)

library(rnaturalearth)
coastlines <- ne_coastline(scale = "small", returnclass = "sf")

ggplot() +
  geom_sf(data = GLODAP_sf,
             aes(col = temp)) +
  scale_fill_viridis_c(na.value = "transparent") +
  geom_sf(data = st_wrap_dateline(coastlines),
          colour = "black") +
  coord_sf(crs = st_crs('ESRI:54030')) +
  theme_bw()

summary(GLODAP_sp)

library(gstat)
vg_temp <- variogram(temp~1,
                     data = GLODAP_sp_grid,
                     cutoff = 1e4)
fit_temp <- fit.variogram(vg_temp, vgm("Sph"))

plot(vg_temp, fit_temp)

vg_temp <- variogram(temp~1,
                     data = GLODAP_sp_grid,
                     alpha = c(0,90),
                     cutoff = 1e4)
plot(vg_temp)


vg <- gstat(id = params_local$MLR_target,
            formula = as.formula(paste(sym(
              params_local$MLR_target
            ), "~ 1")),
            data = GLODAP_sp_grid)

for (i_var in params_local$MLR_predictors) {
  #i_var <- params_local$MLR_predictors[1]
  vg <- gstat(vg,
            id = i_var,
            formula = as.formula(paste(sym(
              i_var
            ), "~ 1")),
            data = GLODAP_sp_grid)
}



plot(variogram(vg, cutoff = 1e4))

# ### kriging
# 
# lzn.kr1 = krige(formula = temp~1,
#                 GLODAP_sp,
#                 GLODAP_sp_grid,
#                 model = fit_temp)
# #> [using universal kriging]
# plot(lzn.kr1[1])

```


