---
title: "eMLR - model fitting"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}
params_local <-
  read_rds(here::here("data/auxillary",
                       "params_local.rds"))

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(lubridate)
library(GGally)
library(olsrr)
library(knitr)
library(kableExtra)
library(broom)
library(corrr)
library(GGally)
library(gt)
```

# Required data

Required are:

- cleaned and prepared GLODAPv2_2020 file

```{r read_GLODAP_used_for_fitting}

GLODAP <-
  read_csv(paste(path_version_data,
                 "GLODAPv2.2020_MLR_fitting_ready.csv",
                 sep = ""))

```



# Fitting all models

## RMSE calculation

Individual linear regression models were fitted for C* as a function of each predictor combination. Fitting was performed separately within each basin, era, and slab. The root mean squared error (RMSE) was calculated for each fitted model.

```{r eMLR_find_all_predictor_combinations, results='hide'}

i_basin <- unique(GLODAP$basin)[1]
i_era   <- unique(GLODAP$era)[1]

# subset one basin and era for fitting
GLODAP_basin_era <- GLODAP %>%
  filter(basin == i_basin, era == i_era)

i_gamma_slab <- unique(GLODAP_basin_era$gamma_slab)[1]
print(i_gamma_slab)

# subset one gamma slab
GLODAP_basin_era_slab <- GLODAP_basin_era %>%
  filter(gamma_slab == i_gamma_slab)

# fit the full linear model, i.e. all predictor combinations
lm_full <- lm(paste(
  params_local$MLR_target,
  paste(params_local$MLR_predictors, collapse = " + "),
  sep = " ~ "
),
data = GLODAP_basin_era_slab)

# fit linear models for all possible predictor combinations
# unfortunately, this functions does not provide model coefficients (yet)
lm_all <- ols_step_all_possible(lm_full)

# extract diagnostics of each linear model
lm_all <- as_tibble(lm_all$result)

rm(i_gamma_slab,
   i_era,
   i_basin,
   GLODAP_basin_era,
   GLODAP_basin_era_slab,
   lm_full)

```

RMSE was plotted against the number of predictors (limited to `r params_local$predictors_min` - `r params_local$predictors_max`).

```{r apply_MLR_predictor_threshold}

lm_all <- lm_all %>% 
  filter(n >= params_local$predictors_min,
         n <= params_local$predictors_max)

```

# Fitting best models

After selecting `r params_local$MLR_number` linear regression models with lowest summed RMSE across all eras, models are fitted again and model coefficients are saved to file.

```{r prepare_models}

# format model formula
lm_all <- lm_all %>% 
  select(predictors) %>% 
  mutate(lm_coeff = str_replace_all(predictors, " ", " + "),
         lm_coeff = paste(params_local$MLR_target, "~", lm_coeff))


```


```{r fit_all_models, results='hide'}


# run similar loop as for rmse calculation
# to get model coefficients for 10 best models
for (i_basin in unique(GLODAP$basin)) {
  for (i_era in unique(GLODAP$era)) {
    # i_basin <- unique(GLODAP$basin)[1]
    # i_era   <- unique(GLODAP$era)[1]
    print(i_basin)
    print(i_era)
    
    GLODAP_basin_era <- GLODAP %>%
      filter(basin == i_basin, era == i_era)
    
    for (i_gamma_slab in unique(GLODAP_basin_era$gamma_slab)) {
      # i_gamma_slab <- unique(GLODAP_basin_era$gamma_slab)[1]
      print(i_gamma_slab)
      
      GLODAP_basin_era_slab <- GLODAP_basin_era %>%
        filter(gamma_slab == i_gamma_slab)
      
      for (i_predictors in unique(lm_all$predictors)) {
        # i_predictors <- unique(lm_all$predictors)[2]
        print(i_predictors)
        
        lm_single <- lm_all %>%
          filter(predictors == i_predictors) %>%
          select(lm_coeff) %>%
          pull()
        
        lm_single_formula <- as.formula(lm_single)
        
        lm_fit_single <- lm(lm_single_formula,
                            data = GLODAP_basin_era_slab)
        
        # plotting model diagnostics
        if (params_local$plot_all_figures == "y") {
          p_model <- ggnostic(
            lm_fit_single,
            columnsY = c(params_local$MLR_target, ".fitted", ".resid"),
            title = paste(
              "eras:",
              i_eras,
              "| era:",
              i_era,
              "| basin:",
              i_basin,
              "| gamma slab:",
              i_gamma_slab,
              "| predictors:",
              i_predictors
            )
          )
          
          ggsave(
            plot = p_model,
            path = paste(path_version_figures, "eMLR_diagnostics/", sep = ""),
            filename = paste(
              "MLR_residuals",
              i_eras,
              i_era,
              i_basin,
              i_gamma_slab,
              i_predictors,
              "predictors.png",
              sep = "_"
            ),
            width = 14,
            height = 8
          )
          
          rm(p_model)
          
        }
        
        i_rmse <-
          c(crossprod(lm_fit_single$residuals)) / length(lm_fit_single$residuals)
        
        # collect model coefficients in data frame
        coefficients <- tidy(lm_fit_single)
        coefficients <- coefficients %>%
          mutate(
            basin = i_basin,
            era = i_era,
            gamma_slab = i_gamma_slab,
            model = lm_single,
            rmse = i_rmse
          )
        
        if (exists("lm_all_fitted")) {
          lm_all_fitted <- bind_rows(lm_all_fitted, coefficients)
        }
        
        if (!exists("lm_all_fitted")) {
          lm_all_fitted <- coefficients
        }
        
      }
      
      
    }
  }
  
}

rm(lm_fit_single, coefficients, i_rmse,
   GLODAP_basin_era, GLODAP_basin_era_slab,
   lm_single, lm_single_formula,
   i_basin, i_era, i_gamma_slab, i_predictors, lm_all)


```




## Prepare coeffcients

Coefficients are prepared for the mapping of Cant and the target variable.

```{r prepare_model_coefficients_for_mapping}


# select relevant columns
lm_all_fitted <- lm_all_fitted %>% 
  select(basin, gamma_slab, era, model, term, estimate, rmse)

# set coefficient to zero if not fitted (=NA)
lm_all_fitted <- lm_all_fitted %>% 
  mutate(estimate = if_else(is.na(estimate), 0, estimate))

# Prepare model coefficients for mapping of target variable
lm_all_fitted_wide <- lm_all_fitted %>% 
  pivot_wider(values_from = estimate,
              names_from = term,
              names_prefix = "coeff_",
              values_fill = 0)

```


## Predictor selection

Within each basin and slab, the `r params_local$MLR_number` linear regression models with lowest summed RMSE across two adjacent (ie compared) eras were selected.

```{r predictor_selection}

# calculate RMSE sum for adjacent eras
lm_all_fitted_wide_eras <- lm_all_fitted_wide  %>%
  arrange(era) %>% 
  group_by(basin, gamma_slab, model) %>% 
  mutate(rmse_sum = rmse + lag(rmse),
         eras = paste(lag(era), era, sep = " --> ")) %>% 
  ungroup() %>% 
  arrange(basin, gamma_slab, model, era) %>% 
  drop_na()

# subset models with lowest RMSE sum
lm_best <- lm_all_fitted_wide_eras %>%
  group_by(basin, gamma_slab, eras) %>%
  slice_min(order_by = rmse_sum,
            with_ties = FALSE,
            n = params_local$MLR_number) %>%
  ungroup() %>% 
  arrange(basin, gamma_slab, eras, model, era)

# plot table for website
lm_best %>% 
  select(basin, gamma_slab, era, eras, model, rmse, rmse_sum) %>% 
  kable() %>%
  add_header_above() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")

```


```{r prepare_model_table}

# create table with two era belonging to one eras
eras_forward <- lm_all_fitted_wide %>%
  arrange(era) %>% 
  group_by(basin, gamma_slab, model) %>% 
  mutate(eras = paste(era, lead(era), sep = " --> ")) %>% 
  ungroup() %>% 
  select(era, eras) %>% 
  unique()

eras_backward <- lm_all_fitted_wide %>%
  arrange(era) %>% 
  group_by(basin, gamma_slab, model) %>% 
  mutate(eras = paste(lag(era), era, sep = " --> ")) %>% 
  ungroup() %>% 
  select(era, eras) %>% 
  unique()

eras_era <- full_join(eras_backward, eras_forward) %>% 
  filter(str_detect(eras, "NA") == FALSE)

# extend best model selection from eras to era
lm_best <- lm_best %>% 
  select(basin, gamma_slab, model, eras)

lm_best <- full_join(lm_best, eras_era)
lm_all_target <- left_join(lm_best, lm_all_fitted_wide)

rm(eras_era, eras_forward, eras_backward,
   lm_all_fitted, lm_all_fitted_wide, lm_all_fitted_wide_eras)


```

```{r prepare_delta_coefficients_for_cant_mapping}


# Prepare model coefficients differences for mapping of cant
# this requires to subtract coefficients of adjacent era


lm_best_long <- lm_all_target %>%
  pivot_longer(`coeff_(Intercept)`:coeff_oxygen,
               names_to = "term",
               values_to = "estimate",
               names_prefix = "coeff_")

  
lm_best_long <- lm_best_long %>%
  arrange(era) %>%
  group_by(basin, gamma_slab, eras, model, term) %>%
  mutate(delta_coeff = estimate - lag(estimate)) %>%
  ungroup() %>%
  arrange(basin, gamma_slab, model, term, eras) %>%
  drop_na() %>%
  select(-c(era,estimate))


lm_all_cant <- lm_best_long %>%
  pivot_wider(values_from = delta_coeff,
              names_from = term,
              names_prefix = "delta_coeff_",
              values_fill = 0)

```


## Write files

```{r write_model_coefficients_files}

lm_all_target %>%
  select(-rmse) %>%
  write_csv(paste(path_version_data,
                  "lm_all_target.csv",
                  sep = ""))

lm_all_cant %>%
  select(-rmse) %>%
  write_csv(paste(path_version_data,
                  "lm_all_cant.csv",
                  sep = ""))


```







## Correlations RMSE between eras

### All models

```{r rmse_correlation_across_eras_all_models, eval=FALSE}

max_rmse <-
  max(c(lm_all_fitted_wide_eras$rmse,
        lm_all_fitted_wide_eras$rmse_sum - lm_all_fitted_wide_eras$rmse))

lm_all_fitted_wide_eras %>%
  ggplot(aes(rmse, rmse_sum - rmse, col = gamma_slab)) +
  geom_point() +
  scale_color_viridis_d() +
  coord_equal(xlim = c(0,max_rmse),
              ylim = c(0,max_rmse)) +
  facet_grid(eras ~ basin)

rm(max_rmse)

```


### Best models

```{r rmse_correlation_across_eras_10_models, eval=FALSE}

max_rmse <-
  max(c(lm_best$rmse,
        lm_best$rmse_sum - lm_best$rmse))

lm_best %>%
  ggplot(aes(rmse, rmse_sum - rmse, col = gamma_slab)) +
  geom_point() +
  scale_color_viridis_d() +
  coord_equal(xlim = c(0,max_rmse),
              ylim = c(0,max_rmse)) +
  facet_grid(eras ~ basin)

rm(max_rmse)
```





```{r eMLR_predictor_model_quality, fig.asp=2, eval=FALSE}

lm_all_stats <- lm_all %>% 
  group_by(basin, gamma_slab, era, model) %>% 
  summarise(n = n(),
            rmse = mean(rmse)) %>% 
  ungroup()

lm_all_stats %>%
  ggplot(aes(n, rmse, col = basin)) +
  geom_hline(yintercept = 10) +
  geom_point(shape = 21) +
  facet_grid(gamma_slab~era) +
  scale_color_brewer(palette = "Set1")

```




## Predictor counts

The number of models where a particular predictor was included were counted for each basin, density slab and compared eras

```{r predictor_counts, eval=FALSE}

lm_all_stats <- lm_all %>% 
  filter(term != "(Intercept)") %>% 
  group_by(basin, eras, gamma_slab) %>% 
  count(term) %>% 
  ungroup() %>% 
  pivot_wider(values_from = n, names_from = term)


lm_all_stats %>%
  gt(rowname_col = "gamma_slab",
     groupname_col = c("basin")) %>% 
  summary_rows(
    groups = TRUE,
    fns = list(total = "sum")
  )

```




## RMSE alternatives

AIC is an alternative criterion to RMSE to judge model quality, but not (yet) taken into account.

```{r AIC_RMSE, eval=FALSE ,fig.asp=1}
lm_all_stats %>% 
  ggplot(aes(rmse, aic, col = gamma_slab)) +
  geom_point() +
  scale_color_viridis_d() +
  facet_grid(era~basin)

```


```{r test_rmse_output, eval=FALSE}

# this code tests the rmse output of the ols_step_all_possible() function used
lm_1 <- GLODAP %>% 
  filter(basin == "Atlantic",
         gamma_slab == "(-Inf,26]",
         era == "new_era") %>% 
  lm(cstar ~ sal + tem + aou + oxygen + phosphate_star, data = .)

lm_2 <- GLODAP %>% 
  filter(basin == "Atlantic",
         gamma_slab == "(-Inf,26]",
         era == "new_era") %>% 
  lm(cstar ~ sal + tem + aou + oxygen + phosphate, data = .)

lm_1
lm_2

sqrt(mean(lm_1$residuals^2))
sqrt(mean(lm_2$residuals^2))

```


