---
title: "GLODAPv2_2020 subsetting"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
params:
  Version_ID: "v_XXX"
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup_obs.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}

if (params$Version_ID == "v_XXX") {
  params_local <-
    read_rds(here::here("data/auxillary", "params_local.rds"))
  
} else {
  params_local <-
    read_rds(
      file = paste(path_root, "observations",
                   params$Version_ID,
                   "data/params_local.rds",
                   sep = "/")
    )
  }

```

```{r define_paths, include = FALSE}

path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_preprocessing_model    <-
  paste(path_root, "/model/preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/GLODAPv2_2020_subsetting/",
        sep = "")
```

```{r load_libraries_specific, include = FALSE}
library(lubridate)
library(gt)
library(kableExtra)
```

```{r select_basin_mask, include=FALSE}

basinmask <- basinmask %>% 
  filter(MLR_basins == params_local$MLR_basins) %>% 
  select(-c(MLR_basins, basin_AIP))

```

# Version ID

The results displayed on this site correspond to the Version_ID:

```{r print_version_ID}

params$Version_ID

```

# Read files

Main data source for this project is the preprocessed version of GLODAPv2:


```{r print_GLODAPv2_version}

params_local$GLODAPv2_version

```


```{r read_GLODAPv2_2020_merged_master_file}

GLODAP <-
  read_csv(
    paste0(
      path_preprocessing_model,
      "GLODAPv2.",
      params_local$GLODAPv2_version,
      "_preprocessed_model_runA_both_NA_filled.csv")
  )

GLODAP_CB <-
  read_csv(
    paste0(
      path_preprocessing,
      "GLODAPv2.",
      params_local$GLODAPv2_version,
      "_Canyon-B.csv"
    )
  )

```


# Data preparation

## Filter eras

```{r assign_eras}

# create labels for era
era_labels <- bind_cols(
  start = params_local$era_start,
  end = params_local$era_end)

era_labels <- era_labels %>% 
  mutate(start = if_else(start == -Inf, max(GLODAP$year), start),
         end = if_else(end == Inf, max(GLODAP$year), end),
         era = as.factor(paste(start, end, sep = "-")))

era_labels %>% 
  kbl() %>% 
  kable_styling()
  

# filter GLODAP data within eras
GLODAP <- expand_grid(
  GLODAP,
  era_labels
)

# select data within each era
GLODAP <- GLODAP %>% 
  filter(year >= start & year <= end)

GLODAP <- GLODAP %>% 
  select(-c(start, end))

```

## Spatial boundaries

### Basin mask

The basin mask from the World Ocean Atlas was used. For details consult the data base subsection for [WOA18](https://jens-daniel-mueller.github.io/Cant_eMLR/read_World_Ocean_Atlas_2018.html) data.

Please note that some GLODAP observations were made outside the WOA18 basin mask (i.e. in marginal seas) and will be removed for further analysis.

```{r join_GLODAP_basin_mask}

# use only data inside basinmask
GLODAP <- inner_join(GLODAP, basinmask)

```

### Depth

Observations collected shallower than:

- minimum sampling depth:  `r params_local$depth_min`m

were excluded from the analysis to avoid seasonal bias.

```{r apply_depth_min}

GLODAP <- GLODAP %>% 
  filter(depth >= params_local$depth_min)

```

### Bottomdepth

Observations collected in an area with a:

- minimum bottom depth:  `r params_local$bottomdepth_min`m

were excluded from the analysis to avoid coastal impacts. Please note that minimum bottom depth criterion of 0m means that no filtering was applied here.

```{r apply_bottomdepth_min}

GLODAP <- GLODAP %>% 
  filter(bottomdepth >= params_local$bottomdepth_min)

```

### Coastal cruises

Coastal cruises can be removed by defining a minimum sampling depth threshold for the entire cruises. This criterion was not applied.

```{r apply_sampling_depth_min, eval=FALSE}

GLODAP <- GLODAP %>% 
  group_by(cruise) %>% 
  mutate(max_depth = max(depth)) %>% 
  ungroup()

GLODAP_grid <- GLODAP %>% 
  distinct(lat, lon, cruise, max_depth)

map +
  geom_raster(data = GLODAP_grid,
              aes(lon, lat, fill = "all depth")) +
  geom_raster(data = GLODAP_grid %>% filter(max_depth < 1000),
              aes(lon, lat, fill = "shallow filter")) +
  scale_fill_manual(values = c("grey60", "red"))

```

## CANYON-B cleaning

Data cleaning based on comparison to CANYOB-B predictions is done individually for each variable required in the eMLR approach and constrained to samples collected deeper than `r params_local$CANYON_B_depth` m.


```{r combine_GLODAP_CANYON_B}

# join data frames
GLODAP_combined_raw <- left_join(GLODAP,
                                 GLODAP_CB)

# calculate offset by parameter
GLODAP_combined <- GLODAP_combined_raw %>%
  mutate(
    offset_talk = talk - talk_CANYONB,
    offset_tco2 = tco2 - tco2_CANYONB,
    offset_nitrate = nitrate - nitrate_CANYONB,
    offset_phosphate = phosphate - phosphate_CANYONB,
    offset_silicate = silicate - silicate_CANYONB
  ) %>%
  select(row_number,
         year,
         cruise,
         basin_AIP,
         lat,
         lon,
         depth,
         starts_with("offset_"))


GLODAP_combined %>%
  filter(cruise == 695) %>%
  arrange(depth) %>%
  ggplot(aes(offset_talk, depth)) +
  geom_point() +
  scale_y_reverse() +
  labs(title = "GLODAPv2.2021 vs CANYON-B | cruise 695")

GLODAP %>%
  filter(cruise == 695) %>%
  arrange(depth) %>%
  ggplot(aes(talk, depth,
             fill = as.factor(talkf),
             col = as.factor(talkqc))) +
  geom_point(shape = 21) +
  scale_y_reverse() +
  # scale_color_manual(values = "black", name = "talkqc") +
  # scale_fill_manual(values = c("black", "white", "grey"), name = "talkf") +
  labs(title = "GLODAPv2.2021 | cruise 695")


# pivot to long format
GLODAP_combined <- GLODAP_combined %>%
  pivot_longer(
    starts_with("offset"),
    names_to = "parameter",
    names_prefix = "offset_",
    values_to = "offset"
  )

GLODAP_combined <- GLODAP_combined %>%
  filter(parameter %in% c("tco2", "talk", params_local$MLR_predictors))

GLODAP_combined <- left_join(GLODAP_combined,
                             GLODAP %>% select(row_number, era))


```

### Residual histograms

```{r CANYON_B_residual_histograms, fig.asp=0.6}

# flag data according to sampling depth
GLODAP_combined <- GLODAP_combined %>%
  mutate(sampling_depth =
           if_else(depth >= params_local$CANYON_B_depth,
                   "deep", "shallow"))

GLODAP_combined %>%
  # filter(cruise == 695) %>%
  ggplot(aes(offset, fill=sampling_depth)) +
  geom_histogram() +
  scale_y_continuous(trans = "log10") +
  facet_grid(basin_AIP ~ parameter, scales = "free_x")

# filter only deep water samples
GLODAP_combined <- GLODAP_combined %>% 
  filter(sampling_depth == "deep")

GLODAP_combined %>%
  # filter(cruise == 695) %>% 
  ggplot(aes(offset, fill=sampling_depth)) +
  geom_histogram() +
  scale_y_continuous(trans = "log10") +
  facet_grid(basin_AIP ~ parameter, scales = "free_x")

GLODAP_combined <- GLODAP_combined %>% 
  select(-sampling_depth)

```


### Property-property plots

```{r CANYON_B_property_property, fig.asp=0.5}

variables <- unique(GLODAP_combined$parameter)

for (i_variable in variables) {
  # i_variable <- variables[1]
  print(p_prop_prop(
    df = GLODAP_combined_raw %>%
      filter(depth >= params_local$CANYON_B_depth),
    var1 = i_variable,
    var2 = paste0(i_variable, "_CANYONB")
  ))
}

rm(variables, GLODAP_combined_raw, i_variable)

```

### Residual time series

```{r CANYON_B_residual_time_series, fig.asp=1}

GLODAP_combined %>%
  ggplot(aes(year, offset)) +
  geom_bin2d() +
  geom_hline(yintercept = 0, col = "red") +
  scale_fill_viridis_c(trans = "log10") +
  facet_grid(parameter ~ basin_AIP, scales = "free_y")

```

### Cruise offset metrics

```{r CANYON_B_residuals_cruise_metrics, fig.asp=0.4}

# calculate mean cruise offset by parameter
cruise_stats <- GLODAP_combined %>%
  group_by(cruise, parameter) %>%
  summarise(
    mean_offset = mean(offset, na.rm = TRUE),
    sd_offset = sd(offset, na.rm = TRUE)
  ) %>%
  ungroup()

cruise_stats <- cruise_stats %>%
  group_by(parameter) %>%
  mutate(
    mean_offset_rank = rank(mean_offset)
  ) %>%
  ungroup() %>% 
  arrange(mean_offset_rank)

# calculate offset parameters across all observations
cruise_stats_global <- cruise_stats %>%
  group_by(parameter) %>%
  summarise(
    mean_offset_max = sd(mean_offset, na.rm = TRUE) *
      params_local$CANYON_B_cruise_mean,
    sd_offset_max = median(sd_offset, na.rm = TRUE) *
      params_local$CANYON_B_sample_SD
  ) %>%
  ungroup()


# rank offsets and calculate offset thresholds
cruise_stats <- full_join(cruise_stats,
                          cruise_stats_global) %>%
  mutate(cruise = as.factor(cruise))


```

### Filter cruises

The following approach was tested but is currently not applied:

Cruises are removed, when the mean offset of the observation from the value predicted with CANYON-B is higher than `r params_local$CANYON_B_max` times the standard deviation of all cruise mean offsets.

```{r CANYON_B_filter_cruises, fig.asp=0.5}

# identify cruises to be removed
cruise_stats <- cruise_stats %>%
  mutate(cruise_flag = if_else(abs(mean_offset) > mean_offset_max,
                               "removed",
                               "included"))

cruise_stats %>% 
  ggplot() +
  geom_vline(data = cruise_stats_global,
             aes(xintercept = mean_offset_max * c(1),
             col = "threshold")) +
  geom_vline(data = cruise_stats_global,
             aes(xintercept = mean_offset_max * c(-1),
             col = "threshold")) +
  geom_histogram(aes(mean_offset, fill = cruise_flag)) +
  facet_grid(. ~ parameter, scales = "free_x") +
  scale_color_manual(values = "red") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  theme(legend.title = element_blank())

for (i_parameter in unique(cruise_stats$parameter)) {
  # i_parameter <- unique(cruise_stats$parameter)[4]
  
  i_cruise_stats <- cruise_stats %>%
    filter(parameter == i_parameter)
  
  i_cruise_out <- cruise_stats %>%
    filter(parameter == i_parameter,
           cruise_flag == "removed")
  
  print(
    ggplot() +
      geom_hline(
        data = i_cruise_stats,
        aes(yintercept = c(-1, 1) * mean_offset_max),
        lty = 2
      ) +
      geom_ribbon(
        data = i_cruise_stats,
        aes(
          x = mean_offset_rank,
          ymax = mean_offset + sd_offset,
          ymin = mean_offset - sd_offset
        ),
        alpha = 0.3
      ) +
      geom_path(data = i_cruise_stats,
                aes(mean_offset_rank, mean_offset)) +
      geom_point(data = i_cruise_out,
                 aes(mean_offset_rank, mean_offset, col = cruise)) +
      labs(title = i_parameter)
  )
  
  print(
    ggplot() +
      geom_point(data = i_cruise_stats,
                 aes(mean_offset, sd_offset)) +
      geom_point(data = i_cruise_out,
                 aes(mean_offset, sd_offset, col = cruise)) +
      labs(title = i_parameter)
  )
  
}

rm(i_parameter, i_cruise_stats, i_cruise_out)

```


```{r CANYON_B_removed_cruises_map, fig.asp=1}

cruise_removal_grid <-
  left_join(
    GLODAP_combined %>%
      distinct(era, parameter, cruise, lon, lat) %>%
      mutate(cruise = as.factor(cruise)),
    cruise_stats %>% select(cruise, parameter, cruise_flag)
  )


map +
  geom_raster(data = GLODAP %>% distinct(lat, lon, era),
              aes(lon, lat), fill = "grey70") +
  geom_raster(data = cruise_removal_grid %>% filter(cruise_flag == "removed"),
              aes(lon, lat, fill = cruise_flag)) +
  scale_fill_brewer(palette = "Set1") +
  facet_grid(parameter ~ era) +
  labs(title = "Cruise filtering") +
  theme(legend.title = element_blank())

map +
  geom_raster(data = GLODAP %>% distinct(lat, lon, era),
              aes(lon, lat), fill = "grey70") +
  geom_raster(data = cruise_removal_grid %>% filter(cruise_flag == "included"),
              aes(lon, lat, fill = cruise_flag)) +
  scale_fill_brewer(palette = "Set1") +
  facet_grid(parameter ~ era) +
  labs(title = "Cruise filtering") +
  theme(legend.title = element_blank())

map +
  geom_raster(data = GLODAP %>% distinct(lat, lon, era),
              aes(lon, lat), fill = "grey70") +
  geom_raster(data = cruise_removal_grid %>% filter(is.na(cruise_flag)),
              aes(lon, lat, fill = "NA")) +
  scale_fill_brewer(palette = "Set1") +
  facet_grid(parameter ~ era) +
  labs(title = "Cruise filtering") +
  theme(legend.title = element_blank())

```

The ratio (%) of removed observations is:

```{r CANYON_B_remove_cruises}

cruise_out <- cruise_stats %>% 
  filter(cruise_flag == "removed") %>% 
  distinct(cruise) %>% 
  pull()

GLODAP <- GLODAP %>%
  mutate(cruise_flag = if_else(
    cruise %in% cruise_out,
    "removed",
    "included"))


cruise_removal_ratio <-
  nrow(GLODAP %>% filter(cruise_flag == "removed")) /
  nrow(GLODAP) * 100

```

```{r stats_cruise_cleaning}

stats_CANYON_B_cleaning_cruises <- tibble(
  cruise_removal_ratio = cruise_removal_ratio,
  cruise_removal_n = length(cruise_out)
)

stats_CANYON_B_cleaning_cruises %>% 
  kbl() %>% 
  kable_styling()

```


### Filter samples

```{r filter_samples}

sample_stats <- GLODAP_combined %>%
  group_by(parameter) %>%
  summarise(
    mean_offset = mean(offset, na.rm = TRUE),
    sd_offset = sd(offset, na.rm = TRUE)
  ) %>%
  ungroup() %>% 
  mutate(sd_offset_max = sd_offset * params_local$CANYON_B_sample_SD)


GLODAP_combined %>%
  ggplot(aes(offset)) +
  geom_histogram() +
  geom_vline(data = sample_stats,
             aes(xintercept = sd_offset_max*c(-1))) +
  geom_vline(data = sample_stats,
             aes(xintercept = sd_offset_max*c(1))) +
  scale_y_continuous(trans = "log10") +
  facet_grid(. ~ parameter, scales = "free_x")

# cruise_stats %>% 
#   ggplot() +
#   geom_vline(aes(xintercept = sd_offset_max,
#              col = "threshold")) +
#   geom_histogram(aes(sd_offset)) +
#   facet_grid(. ~ parameter, scales = "free_x") +
#   scale_color_manual(values = "red") +
#   scale_fill_brewer(palette = "Set1", direction = -1) +
#   theme(legend.title = element_blank())


GLODAP_combined <-
  full_join(GLODAP_combined %>% mutate(cruise = as.factor(cruise)),
            sample_stats)

GLODAP_combined <- GLODAP_combined %>%
  mutate(sample_flag = if_else(abs(offset) <= sd_offset_max,
                               "included", 
                               "removed"))

GLODAP_combined %>%
  ggplot(aes(offset, fill=sample_flag)) +
  geom_histogram() +
  scale_y_continuous(trans = "log10") +
  facet_grid(basin_AIP ~ parameter, scales = "free_x")

GLODAP_combined %>%
  filter(sample_flag == "removed") %>% 
  ggplot(aes(offset, fill=sample_flag)) +
  geom_histogram() +
  scale_y_continuous() +
  facet_grid(basin_AIP ~ parameter, scales = "free_x")

samples_out <- GLODAP_combined %>% 
  filter(sample_flag == "removed") %>% 
  distinct(row_number, parameter)

stats_CANYON_B_cleaning_samples <- GLODAP_combined %>% 
  group_by(parameter) %>% 
  count(sample_flag)

stats_CANYON_B_cleaning_samples %>% 
  kbl() %>% 
  kable_styling()

```


```{r CANYON_B_removed_samples_map, fig.asp=1}

grid_CANYON_B_cleaning_samples <- GLODAP_combined %>% 
  count(lat, lon, era, parameter, sample_flag) %>% 
  pivot_wider(names_from = sample_flag,
              values_from = n,
              values_fill = 0) %>% 
  mutate(removal_ratio = removed / (removed + included))
  
map +
    geom_raster(data = grid_CANYON_B_cleaning_samples,
                aes(lon, lat, fill = removal_ratio)) +
    scale_fill_viridis_c(direction = -1,
                         na.value = "red") +
    facet_grid(parameter ~ era) +
    labs(title = "Maps of removed samples") +
    theme(legend.title = element_blank())

```


The ratio (%) of removed observations is:

```{r CANYON_B_remove_samples}


for (i_parameter in unique(samples_out$parameter)) {
  # i_parameter <- unique(samples_out$parameter)[1]
  
  i_row_number <- samples_out %>%
    filter(parameter == i_parameter) %>%
    pull(row_number)
  
  print(paste(i_parameter, ", NA data before sample filter:"))
  
  print(
  GLODAP %>%
    count(is.na(!!sym(i_parameter))) %>% 
    pull(n)
  )
  
  GLODAP <- GLODAP %>%
    mutate(!!sym(i_parameter) := if_else(row_number %in% i_row_number,
                                         NaN,
                                         !!sym(i_parameter)))
  
  
  print(paste(i_parameter, ", NA data after sample filter:"))
  
  print(
  GLODAP %>%
    count(is.na(!!sym(i_parameter))) %>% 
    pull(n)
  )
  
}


```

```{r final_data_removal, eval=FALSE}

GLODAP %>% 
  filter(sample_flag == "removed",
         cruise_flag == "removed") %>% 
  nrow()

GLODAP %>% 
  filter(sample_flag == "removed",
         cruise_flag == "included") %>% 
  nrow()

# the code below would completely remove rows
# making it impossible to replace with CANYON-B estimates
# this is not done any more

GLODAP %>% 
  filter(sample_flag == "included",
         cruise_flag == "removed") %>% 
  nrow()

GLODAP <- GLODAP %>% 
  filter(sample_flag != "removed",
         cruise_flag != "removed")

```



## Gap filling

```{r GLODAP_gap_filling_CANYON_B, fig.asp=1}

if (params_local$gap_filling == "CANYON-B") {
  
GLODAP <- left_join(GLODAP,
                    GLODAP_CB)

# label rows with missing values that should be filled
GLODAP <- GLODAP %>%
  mutate(
    fill_nitrate = if_else(is.na(nitrate), "filled", "measured"),
    fill_phosphate = if_else(is.na(phosphate), "filled", "measured"),
    fill_silicate = if_else(is.na(silicate), "filled", "measured"),
    fill_talk = if_else(is.na(talk), "filled", "measured")
  )

# fill missing values with CANYON-B estimate, if available
GLODAP <- GLODAP %>% 
  mutate(nitrate = if_else(is.na(nitrate), nitrate_CANYONB, nitrate),
         phosphate = if_else(is.na(phosphate), phosphate_CANYONB, phosphate),
         silicate = if_else(is.na(silicate), silicate_CANYONB, silicate),
         talk = if_else(is.na(talk), talk_CANYONB, talk))

# label rows with remaining missing values that could not be filled
GLODAP <- GLODAP %>%
  mutate(
    fill_nitrate = if_else(is.na(nitrate), "missing", fill_nitrate),
    fill_phosphate = if_else(is.na(phosphate), "missing", fill_phosphate),
    fill_silicate = if_else(is.na(silicate), "missing", fill_silicate),
    fill_talk = if_else(is.na(talk), "missing", fill_talk)
  )

GLODAP_filling_long <- GLODAP %>% 
  select(lon, lat, era, fill_nitrate, fill_phosphate, fill_silicate, fill_talk) %>% 
  pivot_longer(fill_nitrate:fill_talk,
               names_to = "parameter",
               values_to = "filling",
               names_prefix = "fill_") %>% 
  count(lon, lat, era, parameter, filling)

GLODAP_filling_wide <- GLODAP_filling_long %>% 
  pivot_wider(names_from = filling,
              values_from = n,
              values_fill = 0) %>% 
  mutate(total = measured + filled + missing,
         filled_ratio = 100*filled/total,
         missing_ratio = 100*missing/total)

print(
map + 
  geom_raster(data = GLODAP_filling_wide,
              aes(lon, lat, fill = filled_ratio)) +
  facet_grid(parameter ~ era) +
  scale_fill_viridis_c(direction = -1,
                       na.value = "red")
)

print(
map + 
  geom_raster(data = GLODAP_filling_wide,
              aes(lon, lat, fill = missing_ratio)) +
  facet_grid(parameter ~ era) +
  scale_fill_viridis_c(direction = -1,
                       na.value = "red")
)


print(
GLODAP_filling_long %>% 
  group_by(era, parameter, filling) %>% 
  summarise(n = sum(n)) %>% 
  ggplot(aes(parameter, n, fill = filling)) +
  coord_flip() +
  geom_col() +
  facet_grid(era~.) +
  scale_fill_viridis_d()
)

GLODAP <- GLODAP %>% 
  select(!ends_with("_CANYONB")) %>% 
  select(!starts_with("fill_"))

rm(GLODAP_filling_long, GLODAP_filling_wide)

}


```



## Flags and missing data

Only rows (samples) for which all relevant parameters are available were selected, ie NA's were removed.

According to Olsen et al (2020), flags within the merged master file identify:

- f:  

  - 2: Acceptable
  - 0: Interpolated (nutrients/oxygen) or calculated (CO[2] variables)
  - 9: Data not used (so, only NA data should have this flag)
  
  
- qc:  

  - 1: Adjusted or unadjusted data
  - 0: Data appear of good quality but have not been subjected to full secondary QC
  - data with poor or uncertain quality are excluded.

Following flagging criteria were taken into account:

- flag_f:     `r params_local$flag_f`
- flag_qc:    `r params_local$flag_qc`

The cleaning process was performed successively and the maps below represent the data coverage at various cleaning levels. 

Summary statistics were calculated during cleaning process.

### tco2

#### NA

Rows with missing tco2 observations were already removed in the preprocessing. The map below shows the coverage of preprocessed GLODAP data.

```{r tco2_na_subset}

cruise_stats_global <- GLODAP %>% 
  summarise(tco2_values = n())

GLODAP_obs_grid <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "tco2_values")

```


```{r WOA18_basin_mask_GLODAP_observations_map, fig.asp=0.6}

GLODAP_obs <- GLODAP %>% 
  group_by(lat, lon) %>% 
  summarise(n = n()) %>% 
  ungroup()

map +
  geom_raster(data = basinmask, aes(lon, lat, fill = basin)) +
  geom_raster(data = GLODAP_obs, aes(lon, lat)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "top",
        legend.title = element_blank())

rm(GLODAP_obs)

```

#### f flag

```{r tco2_f_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, tco2f)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ tco2f) +
  theme(legend.position = "top")

rm(GLODAP_obs_grid_temp)

GLODAP <- GLODAP %>%
  filter(tco2f %in% params_local$flag_f)

```

#### qc flag

```{r tco2_qc_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, tco2qc)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ tco2qc) +
  theme(legend.position = "top")

##

GLODAP <- GLODAP %>%
  filter(tco2qc %in% params_local$flag_qc)

cruise_stats_global_temp <- GLODAP %>%
  summarise(tco2_flag = n())

cruise_stats_global <- cbind(cruise_stats_global, cruise_stats_global_temp)
rm(cruise_stats_global_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "tco2_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### talk

#### NA

```{r talk_na_subset, fig.asp=0.6}

GLODAP <- GLODAP %>% 
  mutate(talkna = if_else(is.na(talk), "NA", "Value"))

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkna)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkna) +
  theme(legend.position = "top")

GLODAP <- GLODAP %>% 
  select(-talkna) %>% 
  filter(!is.na(talk))

##

cruise_stats_global_temp <- GLODAP %>% 
  summarise(talk_values = n())

cruise_stats_global <- cbind(cruise_stats_global, cruise_stats_global_temp)
rm(cruise_stats_global_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "talk_values")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

#### f flag

```{r talk_f_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkf)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkf) +
  theme(legend.position = "top",
        legend.title = element_blank())


# ###

GLODAP <- GLODAP %>%
  filter(talkf %in% params_local$flag_f)


```

#### qc flag

```{r talk_qc_subset, fig.asp=0.6}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, talkqc)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ talkqc) +
  theme(legend.position = "top",
        legend.title = element_blank())

###

GLODAP <- GLODAP %>%
  filter(talkqc %in% params_local$flag_qc)

##

cruise_stats_global_temp <- GLODAP %>%
  summarise(talk_flag = n())

cruise_stats_global <- cbind(cruise_stats_global, cruise_stats_global_temp)
rm(cruise_stats_global_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "talk_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### Phosphate

Phosphate was used as predictor:

```{r check_phosphate_predictor}

P_used <- "phosphate" %in% params_local$MLR_predictors
P_used

```


#### NA

```{r phosphate_na_subset, fig.asp=0.6, eval=P_used}

GLODAP <- GLODAP %>% 
  mutate(phosphatena = if_else(is.na(phosphate), "NA", "Value"))

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphatena)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era ~ phosphatena) +
  theme(legend.position = "top")

GLODAP <- GLODAP %>% 
  select(-phosphatena) %>% 
  filter(!is.na(phosphate))

##

cruise_stats_global_temp <- GLODAP %>% 
  summarise(phosphate_values = n())

cruise_stats_global <- cbind(cruise_stats_global, cruise_stats_global_temp)
rm(cruise_stats_global_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>% 
  count(lat, lon, era) %>% 
  mutate(cleaning_level = "phosphate_values")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

#### f flag

```{r phosphate_f_subset, fig.asp=0.6, eval=P_used}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphatef)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~phosphatef) +
  theme(legend.position = "top",
        legend.title = element_blank())

###

GLODAP <- GLODAP %>%
  filter(phosphatef %in% params_local$flag_f)


```

#### qc flag

```{r phosphate_qc_subset, fig.asp=0.6, eval=P_used}

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era, phosphateqc)

map +
  geom_raster(data = GLODAP_obs_grid_temp, aes(lon, lat, fill = n)) +
    scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  facet_grid(era~phosphateqc) +
  theme(legend.position = "top",
        legend.title = element_blank())


###

GLODAP <- GLODAP %>%
  filter(phosphateqc %in% params_local$flag_qc)

##

cruise_stats_global_temp <- GLODAP %>%
  summarise(phosphate_flag = n())

cruise_stats_global <- cbind(cruise_stats_global, cruise_stats_global_temp)
rm(cruise_stats_global_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "phosphate_flag")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)

```


### eMLR variables

Variables required as predictors for the MLR fits, are subsetted for NAs and flags.

```{r eMLR_subset}

if ("temp" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(temp))
}

##

if ("sal" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(sal))
  
  GLODAP <- GLODAP %>%
    filter(salinityf %in% params_local$flag_f)
  
  GLODAP <- GLODAP %>%
    filter(salinityqc %in% params_local$flag_qc)
}

##

if ("silicate" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(silicate))
  
  GLODAP <- GLODAP %>%
    filter(silicatef %in% params_local$flag_f)
  
  GLODAP <- GLODAP %>%
    filter(silicateqc %in% params_local$flag_qc)
}

##

if ("oxygen" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(oxygen))
  
  GLODAP <- GLODAP %>%
    filter(oxygenf %in% params_local$flag_f)
  
  GLODAP <- GLODAP %>%
    filter(oxygenqc %in% params_local$flag_qc)
}

##

if ("aou" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(aou))
  
  GLODAP <- GLODAP %>%
    filter(aouf %in% params_local$flag_f)
}

##

if ("nitrate" %in% params_local$MLR_predictors) {
  GLODAP <- GLODAP %>%
    filter(!is.na(nitrate))
  
  GLODAP <- GLODAP %>%
    filter(nitratef %in% params_local$flag_f)
  
  GLODAP <- GLODAP %>%
    filter(nitrateqc %in% params_local$flag_qc)
}

##

GLODAP <- GLODAP %>%
  filter(!is.na(depth))

GLODAP <- GLODAP %>%
  filter(!is.na(gamma))

##

cruise_stats_global_temp <- GLODAP %>%
  summarise(eMLR_variables = n())

cruise_stats_global <- cbind(cruise_stats_global, cruise_stats_global_temp)

rm(cruise_stats_global_temp)

##

GLODAP_obs_grid_temp <- GLODAP %>%
  count(lat, lon, era) %>%
  mutate(cleaning_level = "eMLR_variables")

GLODAP_obs_grid <-
  bind_rows(GLODAP_obs_grid, GLODAP_obs_grid_temp)

rm(GLODAP_obs_grid_temp)


```

```{r remove_flag_columns}

GLODAP <- GLODAP %>% 
  select(-ends_with(c("f", "qc")))

```



## Reference years (tref)

### Median years

Median years of each era (tref) were determined as:

```{r calc_tref}

# calculate reference year
tref <- GLODAP %>%
  group_by(era) %>%
  summarise(median_year = median(year)) %>%
  ungroup()

tref

```

### Fixed tref

tref were manually set to:

```{r set_predefined_tref}

tref <- tref %>%
  arrange(median_year)

tref1 <- tref %>%
  head(1)

tref2 <- tref %>%
  tail(1)

if (!is.null(params_local$tref1)) {
  tref1 <- tref1 %>%
    mutate(median_year = params_local$tref1)
}

if (!is.null(params_local$tref2)) {
  tref2 <- tref2 %>%
    mutate(median_year = params_local$tref2)
}

tref <- bind_rows(tref1, tref2)

tref <- full_join(tref, era_labels)
tref

```


## Write files
 
```{r write_clean_data_files}

# select relevant columns for further analysis
GLODAP <- GLODAP %>% 
  select(year, date, era, basin, basin_AIP, lat, lon, cruise,
         bottomdepth, depth,
         temp, sal, gamma,
         tco2, talk, phosphate,
         oxygen, aou, nitrate, silicate,
         temp_model, sal_model = so_model, gamma_model,
         tco2_model = dissic_model, talk_model, phosphate_model = po4_model,
         oxygen_model = o2_model, aou_model,
         nitrate_model = no3_model, silicate_model = si_model
         )


GLODAP  %>%  write_csv(paste(path_version_data,
                             "GLODAPv2.2020_clean.csv",
                             sep = ""))

tref  %>%  write_csv(paste(path_version_data,
                           "tref.csv",
                           sep = ""))

cruise_stats  %>%  write_csv(paste(
  path_version_data,
  "CANYON_B_cleaning_cruise_stats.csv",
  sep = ""
))

cruise_stats_global  %>%  write_csv(paste(
  path_version_data,
  "CANYON_B_cleaning_cruise_stats_global.csv",
  sep = ""
))

stats_CANYON_B_cleaning_cruises  %>%  write_csv(paste(
  path_version_data,
  "CANYON_B_cleaning_GLODAP_cruise_removal_stats.csv",
  sep = ""
))

stats_CANYON_B_cleaning_samples  %>%  write_csv(paste(
  path_version_data,
  "CANYON_B_cleaning_GLODAP_sample_removal_stats.csv",
  sep = ""
))

grid_CANYON_B_cleaning_samples  %>%  write_csv(paste(
  path_version_data,
  "CANYON_B_cleaning_GLODAP_sample_removal_grid.csv",
  sep = ""
))



```


# Overview plots

## Number of overservations

```{r counts}

GLODAP %>% 
  group_by(era, basin_AIP) %>% 
  count() %>% 
  ggplot(aes(basin_AIP, n, fill = era)) +
  geom_col() +
  scale_fill_brewer(palette = "Dark2")

```


## Cleaning stats

Number of observations at various steps of data cleaning.

```{r GLODAP_cleaning_stats}

cruise_stats_global_long <- cruise_stats_global %>%
  pivot_longer(1:length(cruise_stats_global),
               names_to = "parameter",
               values_to = "n")

cruise_stats_global_long <- cruise_stats_global_long %>%
  mutate(parameter = fct_reorder(parameter, n))

cruise_stats_global_long %>% 
  ggplot(aes(parameter, n/1000)) +
  geom_col() +
  coord_flip() +
  theme(axis.title.y = element_blank())

rm(cruise_stats_global_long)

```


## Assign coarse spatial grid

For the following plots, the cleaned data set was re-opened and observations were gridded spatially to intervals of:  

- 5° x 5°

```{r grid_spatially_5x5}

GLODAP <- m_grid_horizontal_coarse(GLODAP)

```


## Histogram Zonal coverage

```{r coverage_histogram_zonal}

GLODAP_histogram_lat <- GLODAP %>%
  group_by(era, lat_grid, basin_AIP) %>%
  tally() %>%
  ungroup()

GLODAP_histogram_lat %>%
  ggplot(aes(lat_grid, n, fill = era)) +
  geom_col() +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap( ~ basin_AIP) +
  coord_flip(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank())

rm(GLODAP_histogram_lat)

```


## Histogram temporal coverage

```{r coverage_histogram_temporal}

GLODAP_histogram_year <- GLODAP %>%
  group_by(year, basin_AIP) %>%
  tally() %>%
  ungroup()

GLODAP_histogram_year %>%
  ggplot() +
  geom_vline(xintercept = sort(params_local$era_end)[1] + 0.5) +
  geom_col(aes(year, n,
               fill = basin_AIP)) +
  geom_point(
    data = tref,
    aes(median_year, 100, shape = "tref"),
    size = 2,
    fill = "white"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  scale_shape_manual(values = 24, name = "") +
  scale_y_continuous() +
  coord_cartesian(expand = 0) +
  theme(
    legend.position = "top",
    legend.direction = "vertical",
    legend.title = element_blank(),
    axis.title.x = element_blank()
  )

rm(GLODAP_histogram_year,
   era_median_year)

```

## Zonal temporal coverage (Hovmoeller)

```{r coverage_hovmoeller, fig.asp=1.3}

GLODAP_hovmoeller_year <- GLODAP %>%
  group_by(year, lat_grid, basin_AIP) %>%
  tally() %>%
  ungroup()

GLODAP_hovmoeller_year %>%
  ggplot(aes(year, lat_grid, fill = n)) +
  geom_tile() +
  geom_vline(xintercept = sort(params_local$era_end)[1] + 0.5) +
  scale_fill_viridis_c(option = "magma",
                       direction = -1,
                       trans = "log10") +
  coord_cartesian(expand = 0) +
  facet_wrap( ~ basin_AIP, ncol = 1) +
  theme(legend.position = "top",
        axis.title.x = element_blank())

rm(GLODAP_hovmoeller_year)

```

## Coverage maps by era

### Subsetting process

The following plots show the remaining data after **individual cleaning steps**, separately for each era.

```{r coverage_maps_era_subsetting_era_cleaning_level, fig.asp=1.1}

GLODAP_obs_grid <- GLODAP_obs_grid %>%
  mutate(cleaning_level = factor(
           cleaning_level,
           unique(GLODAP_obs_grid$cleaning_level)
         ))

map +
  geom_raster(data = GLODAP_obs_grid %>%
                filter(cleaning_level == "tco2_values") %>%
                select(-cleaning_level),
              aes(lon, lat, fill = "tco2_values")) +
  geom_raster(data = GLODAP_obs_grid %>%
                filter(cleaning_level != "tco2_values"),
              aes(lon, lat, fill = "subset")) +
  scale_fill_brewer(palette = "Set1", name = "") +
  facet_grid(cleaning_level ~ era) +
  theme(legend.position = "top",
        axis.title = element_blank())

```

### Final input data

The following plots show the remaining data density in each grid cell **after all cleaning steps**, separately for each era.

```{r coverage_maps_era_clean, fig.asp=1}

GLODAP_tco2_grid <- GLODAP %>%
  count(lat, lon)

map +
  geom_bin2d(data = GLODAP,
             aes(lon, lat),
             binwidth = c(1,1)) +
  scale_fill_viridis_c(option = "magma", direction = -1, trans = "log10") +
  facet_wrap(~era, ncol = 1) +
  labs(title = "Cleaned GLODAP observations",
       subtitle = paste("Version:", params_local$Version_ID)) +
  theme(axis.title = element_blank())

ggsave(path = path_version_figures,
       filename = "data_distribution_era.png",
       height = 8,
       width = 5)

```


## Model vs observation

```{r model_obs_comparison}

variables <-
  c(
    "temp",
    "sal",
    "gamma",
    "tco2",
    "talk",
    "phosphate",
    "oxygen",
    "aou",
    "nitrate",
    "silicate"
  )

for (i_var in variables) {
  #  i_var <- variables[1]
  
  p_prop_prop(df = GLODAP,
              var1 = i_var,
              var2 = paste0(i_var, "_model"))
  
}

```

