---
title: "Mapping cant"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
params:
  Version_ID: "v_XXX"
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup_obs.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}

if (params$Version_ID == "v_XXX") {
  params_local <-
    read_rds(here::here("data/auxillary", "params_local.rds"))
  
} else {
  params_local <-
    read_rds(
      file = paste(path_root, "observations",
                   params$Version_ID,
                   "data/params_local.rds",
                   sep = "/")
    )
}


```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_model_preprocessing    <-
  paste(path_root, "/model/preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r load_libraries_specific, include=FALSE}
library(marelac)
```


# Version ID

The results displayed on this site correspond to the Version_ID: `r params$Version_ID`

# Required data

- tcant 3D fields at tref1 and tref2 for variable and constant climate model runs


```{r read_model_tcant_files_vc}

tref  <-
  read_csv(paste(path_version_data,
                 "tref.csv",
                 sep = ""))

tcant_tref_1 <-
  read_csv(
    paste(
      path_model_preprocessing,
      "cant_annual_field_AD",
      "/cant_",
      unique(tref$median_year[1]),
      ".csv",
      sep = ""
    )
  )

tcant_tref_1 <- tcant_tref_1 %>%
  rename(tcant_tref_1 = cant_total) %>%
  select(-year)

tcant_tref_2 <-
  read_csv(
    paste(
      path_model_preprocessing,
      "cant_annual_field_AD",
      "/cant_",
      unique(tref$median_year[2]),
      ".csv",
      sep = ""
    )
  )

tcant_tref_2 <- tcant_tref_2 %>%
  rename(tcant_tref_2 = cant_total) %>%
  select(-year)

```

```{r read_model_tcant_files_cc}

tcant_cc_tref_1 <-
  read_csv(
    paste(
      path_model_preprocessing,
      "cant_annual_field_CB",
      "/cant_",
      unique(tref$median_year[1]),
      ".csv",
      sep = ""
    )
  )

tcant_cc_tref_1 <- tcant_cc_tref_1 %>%
  rename(tcant_tref_1 = cant_total) %>%
  select(-year)

tcant_cc_tref_2 <-
  read_csv(
    paste(
      path_model_preprocessing,
      "cant_annual_field_CB",
      "/cant_",
      unique(tref$median_year[2]),
      ".csv",
      sep = ""
    )
  )

tcant_cc_tref_2 <- tcant_cc_tref_2 %>%
  rename(tcant_tref_2 = cant_total) %>%
  select(-year)

```


```{r load_model_gamma_climatology}

climatology <-
  read_csv(paste(path_model_preprocessing, "climatology_runA_2007.csv", sep = ""))

climatology <- climatology %>%
  select(lon,lat,depth, gamma)

```

# dcant calculation

```{r calc_model_cant_between_tref_vc}

dcant_3d <- left_join(tcant_tref_1, tcant_tref_2) %>%
  mutate(dcant = tcant_tref_2 - tcant_tref_1)

rm(tcant_tref_1, tcant_tref_2)

dcant_3d <- dcant_3d %>%
  mutate(dcant_pos = if_else(dcant <= 0, 0, dcant))

dcant_3d <- full_join(dcant_3d, climatology)

dcant_3d <- m_cut_gamma(dcant_3d, "gamma")

dcant_3d_vc <- dcant_3d %>% 
  mutate(data_source = "mod_truth") %>% 
  select(lon, lat, depth, basin_AIP, data_source,
            tcant_tref_1, dcant, dcant_pos,
            gamma, gamma_slab)

```

```{r calc_model_cant_between_tref_cc}

dcant_3d <- left_join(tcant_cc_tref_1, tcant_cc_tref_2) %>%
  mutate(dcant = tcant_tref_2 - tcant_tref_1)

rm(tcant_cc_tref_1, tcant_cc_tref_2)

dcant_3d <- dcant_3d %>%
  mutate(dcant_pos = if_else(dcant <= 0, 0, dcant))

dcant_3d <- full_join(dcant_3d, climatology)

dcant_3d <- m_cut_gamma(dcant_3d, "gamma")

dcant_3d_cc <- dcant_3d %>% 
  mutate(data_source = "mod_truth_cc") %>% 
  select(lon, lat, depth, basin_AIP, data_source,
            tcant_tref_1, dcant, dcant_pos,
            gamma, gamma_slab)

rm(climatology, dcant_3d)

```

```{r join_climate_runs}

dcant_3d <- bind_rows(
  dcant_3d_cc,
  dcant_3d_vc
)

rm(
  dcant_3d_cc,
  dcant_3d_vc
)


```

# Maps

```{r cant_deep_climatology_map, fig.asp=0.6}

dcant_3d %>%
  group_split(data_source) %>%
  # head(1) %>%
  map(~ p_map_climatology(
    df = .x,
    var = "dcant_pos",
    subtitle_text = paste("Climate: ", .x$data_source)
  ))

```


```{r cant_deep_climatology_sections, fig.asp=1}

dcant_3d %>%
  group_split(data_source) %>%
  # head(1) %>%
  map(~ p_section_climatology_regular(
    df = .x,
    var = "dcant_pos",
    subtitle_text = paste("Climate: ", .x$data_source)
  ))

```

# Averaging and integration

## Zonal sections

```{r calc_cant_zonal_mean_sections}

dcant_zonal <- dcant_3d %>%
  group_by(data_source) %>%
  nest() %>%
  mutate(zonal = map(.x = data, ~m_zonal_mean_sd(.x))) %>%
  select(-data) %>%
  unnest(zonal)


dcant_zonal <- m_cut_gamma(dcant_zonal, "gamma_mean")

```

## Inventories

To calculate dcant column inventories, we:  

1. Convert Cant concentrations to volumetric units
2. Multiply layer thickness with volumetric Cant concentration to get a layer inventory
3. For each horizontal grid cell and era, sum cant layer inventories for different inventory depths (`r params_global$inventory_depths` m)

Step 2 is performed separately for all Cant and positive Cant values only.

```{r calc_dcant_inventories, fig.asp=1}

dcant_inv <- dcant_3d %>%
  group_by(data_source) %>%
  nest() %>%
  mutate(inv = map(.x = data, ~m_dcant_inv(.x))) %>%
  select(-data) %>%
  unnest(inv)

p_map_cant_inv(df = dcant_inv,
               var = "dcant_pos",
               subtitle_text = "for predefined integration depths") +
  facet_grid(inv_depth ~ data_source)

```

```{r cant_total_inventories, fig.asp=1}

tcant_inv <- dcant_3d %>%
  rename(tcant = tcant_tref_1) %>%
  mutate(tcant_pos = if_else(tcant <= 0, 0, tcant)) %>% 
  group_by(data_source) %>%
  nest() %>%
  mutate(inv = map(.x = data, ~m_tcant_inv(.x))) %>%
  select(-data) %>%
  unnest(inv)

p_map_cant_inv(df = tcant_inv,
               var = "tcant_pos",
               breaks = seq(0,70,10),
               subtitle_text = "for predefined integration depths") +
  facet_grid(inv_depth ~ data_source)

```

## Budgets

Global dcant budgets were estimated in units of Pg C. Please note that here we added dcant (all vs postitive only) values and do not apply additional corrections for areas not covered.

```{r calculate_budget_per_grid_cell}

molC_to_PgC <- 12*1e-15

dcant_budget <- dcant_inv %>% 
  mutate(surface_area = earth_surf(lat, lon),
         dcant_grid = dcant*surface_area*molC_to_PgC,
         dcant_pos_grid = dcant_pos*surface_area*molC_to_PgC)
```

```{r budget_global}

dcant_budget_global <- dcant_budget %>%
  group_by(data_source, inv_depth) %>% 
  summarise(dcant = sum(dcant_grid),
            dcant = round(dcant,1),
            dcant_pos = sum(dcant_pos_grid),
            dcant_pos = round(dcant_pos,1)) %>% 
  ungroup() %>% 
  pivot_longer(cols = dcant:dcant_pos,
               names_to = "estimate",
               values_to = "value")

dcant_budget_global %>%
  filter(inv_depth == params_global$inventory_depth_standard) %>%
  ggplot(aes(estimate, value)) +
  scale_fill_brewer(palette = "Dark2") +
  geom_col() +
  facet_grid(~data_source)

```


```{r budget_basin_AIP}

dcant_budget_basin_AIP <- dcant_budget %>% 
  group_by(basin_AIP, data_source, inv_depth) %>% 
  summarise(dcant = sum(dcant_grid),
            dcant = round(dcant,1),
            dcant_pos = sum(dcant_pos_grid),
            dcant_pos = round(dcant_pos,1)) %>% 
  ungroup() %>% 
  pivot_longer(cols = dcant:dcant_pos,
               names_to = "estimate",
               values_to = "value")

dcant_budget_basin_AIP %>%
  filter(inv_depth == params_global$inventory_depth_standard) %>%
  ggplot(aes(estimate, value, fill=basin_AIP)) +
  scale_fill_brewer(palette = "Dark2") +
  geom_col() +
  facet_grid(~data_source)

```


```{r budget_MLR_regions, fig.asp=1}

dcant_budget_basin_MLR <-
  full_join(dcant_budget, basinmask) %>% 
  group_by(basin, MLR_basins, data_source, inv_depth) %>% 
  summarise(dcant = sum(dcant_grid),
            dcant = round(dcant,1),
            dcant_pos = sum(dcant_pos_grid),
            dcant_pos = round(dcant_pos,1)) %>% 
  ungroup() %>% 
  pivot_longer(cols = dcant:dcant_pos,
               names_to = "estimate",
               values_to = "value")

dcant_budget_basin_MLR %>%
  filter(inv_depth == params_global$inventory_depth_standard) %>%
  ggplot(aes(MLR_basins, value, fill=basin)) +
  geom_col() +
  facet_grid(estimate~data_source)

```


```{r budget_lat_grid, fig.asp=1}

dcant_budget_lat_grid <- 
  dcant_budget %>% 
  m_grid_horizontal_coarse() %>%
  group_by(lat_grid, basin_AIP, data_source, inv_depth) %>% 
  summarise(dcant = sum(dcant_grid),
            dcant = round(dcant,1),
            dcant_pos = sum(dcant_pos_grid),
            dcant_pos = round(dcant_pos,1)) %>% 
  ungroup() %>% 
  pivot_longer(cols = dcant:dcant_pos,
               names_to = "estimate",
               values_to = "value")

dcant_budget_lat_grid %>%
  filter(inv_depth == params_global$inventory_depth_standard) %>%
  ggplot(aes(lat_grid, value, fill = basin_AIP)) +
  geom_col() +
  scale_fill_brewer(palette = "Dark2") +
  coord_flip() +
  facet_grid(estimate ~ data_source)

```




# Alpha

## 3d fields

```{r alpha_3d}

alpha_3d <- dcant_3d %>%
  mutate(
    alpha = dcant / tcant_tref_1,
    alpha = if_else(dcant <= 0 | tcant_tref_1 <= 0,
                    NaN, alpha)
  )

alpha_3d %>%
  filter(alpha < 1) %>%
  ggplot(aes(alpha)) +
  geom_histogram() +
  facet_grid(basin_AIP ~ data_source) +
  scale_x_continuous(breaks = seq(0, 1, 0.1))

median_alpha <- 
  alpha_3d %>% 
  group_by(depth, basin_AIP, data_source) %>% 
  summarise(alpha_median = median(alpha, na.rm = TRUE),
            alpha_mean = weighted.mean(alpha, tcant_tref_1, na.rm = TRUE)) %>% 
  ungroup()

alpha_3d %>%
  filter(alpha < 1) %>%
  ggplot(aes(alpha, depth)) +
  geom_bin2d(binwidth = c(0.01, 200)) +
  geom_path(data = median_alpha,
            aes(alpha_median, depth, col="Median")) +
  geom_path(data = median_alpha,
            aes(alpha_mean, depth, col="Weighted mean")) +
  scale_color_manual(name = "Alpha",
                     values = c("red", "orange")) +
  scale_y_reverse() +
  scale_x_continuous(breaks = seq(0, 1, 0.1),
                     limits = c(0,1)) +
  scale_fill_viridis_c(trans = "log10") +
  coord_cartesian(expand = 0) +
  facet_grid(basin_AIP~data_source)


print(paste("mean:",
            round(mean(
              alpha_3d$alpha, na.rm = TRUE
            ), 3)))

print(paste("mean, weighted with Cant:",
            round(
              weighted.mean(
                x = alpha_3d$alpha,
                w = alpha_3d$tcant_tref_1,
                na.rm = TRUE
              ),
              3
            )))

print(paste("median:",
            round(median(
              alpha_3d$alpha, na.rm = TRUE
            ), 3)))


```

## Inventory maps

```{r alpha_surface_distribution, fig.asp=1.5}

alpha_inv <- full_join(
  dcant_inv,
  tcant_inv)

alpha_inv <- alpha_inv %>% 
  mutate(alpha = dcant / tcant)


map +
  geom_raster(data = alpha_inv,
              aes(lon, lat, fill = alpha)) +
  facet_grid(inv_depth ~ data_source) +
  scale_fill_divergent(
    midpoint = median(alpha_inv$alpha)
  )


```



# Write csv

```{r write_cant_files}

dcant_3d <- dcant_3d %>%
  select(-c(tcant_tref_1))

dcant_3d %>%
  filter(data_source == "mod_truth") %>%
  write_csv(paste(path_version_data,
                  "dcant_3d_mod_truth.csv", sep = ""))

dcant_3d %>%
  filter(data_source == "mod_truth_cc") %>%
  write_csv(paste(path_version_data,
                  "dcant_3d_mod_truth_cc.csv", sep = ""))



dcant_zonal %>%
  filter(data_source == "mod_truth") %>%
  write_csv(paste(path_version_data,
                  "dcant_zonal_mod_truth.csv", sep = ""))

dcant_zonal %>%
  filter(data_source == "mod_truth_cc") %>%
  write_csv(paste(path_version_data,
                  "dcant_zonal_mod_truth_cc.csv", sep = ""))



dcant_inv %>%
  filter(data_source == "mod_truth") %>%
  write_csv(paste(path_version_data,
                  "dcant_inv_mod_truth.csv", sep = ""))

dcant_inv %>%
  filter(data_source == "mod_truth_cc") %>%
  write_csv(paste(path_version_data,
                  "dcant_inv_mod_truth_cc.csv", sep = ""))



dcant_budget_global %>%
  filter(data_source == "mod_truth") %>%
  write_csv(paste(path_version_data,
                  "dcant_budget_global_mod_truth.csv", sep = ""))

dcant_budget_global %>%
  filter(data_source == "mod_truth_cc") %>%
  write_csv(paste(path_version_data,
                  "dcant_budget_global_mod_truth_cc.csv", sep = ""))

dcant_budget_basin_AIP %>%
  filter(data_source == "mod_truth") %>%
  write_csv(paste(path_version_data,
                  "dcant_budget_basin_AIP_mod_truth.csv", sep = ""))

dcant_budget_basin_AIP %>%
  filter(data_source == "mod_truth_cc") %>%
  write_csv(paste(path_version_data,
                  "dcant_budget_basin_AIP_mod_truth_cc.csv", sep = ""))

dcant_budget_basin_MLR %>%
  filter(data_source == "mod_truth") %>%
  write_csv(paste(path_version_data,
                  "dcant_budget_basin_MLR_mod_truth.csv", sep = ""))

dcant_budget_basin_MLR %>%
  filter(data_source == "mod_truth_cc") %>%
  write_csv(paste(path_version_data,
                  "dcant_budget_basin_MLR_mod_truth_cc.csv", sep = ""))

dcant_budget_lat_grid %>%
  filter(data_source == "mod_truth") %>%
  write_csv(paste(path_version_data,
                  "dcant_budget_lat_grid_mod_truth.csv", sep = ""))

dcant_budget_lat_grid %>%
  filter(data_source == "mod_truth_cc") %>%
  write_csv(paste(path_version_data,
                  "dcant_budget_lat_grid_mod_truth_cc.csv", sep = ""))

```


