---
title: "Mapping"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
params:
  Version_ID: "v_XXX"
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}

if (params$Version_ID == "v_XXX") {
  params_local <-
    read_rds(here::here("data/auxillary", "params_local.rds"))
  
} else {
  params_local <-
    read_rds(
      file = paste(path_root, "observations",
                   params$Version_ID,
                   "data/params_local.rds",
                   sep = "/")
    )
  }

if (params_local$gobm == "CESM-ETHZ_v20211122_monthly") {
  params_local$gobm <- "CESM-ETHZ_v20211122"
}


```

```{r define_paths, include = FALSE}

path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_preprocessing_model    <-
  paste(path_root, "/model/preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")


```

```{r select_basin_mask, include=FALSE}

basinmask <- basinmask %>% 
  filter(MLR_basins == params_local$MLR_basins) %>% 
  select(-MLR_basins)

```

```{r load_libraries_specific, include=FALSE}
library(metR)
library(marelac)
library(gsw)
```

# Version ID

The results displayed on this site correspond to the Version_ID: `r params$Version_ID`

# Required data

## Observational

Currently, following data sets are used for mapping:

- GLODAPv2_2016b_MappedClimatologies, with variables  
  - Phosphate (-> phosphate_star)
  - Silicate
  - Oxygen (-> AOU)
  - TAlk (surface only)
  - TCO2 (surface only)


```{r load_GLODAPv2_2016b_MappedClimatologies_interior_mapping}

variables <-
  c("oxygen", "PO4", "silicate", "NO3", "TAlk")

# i_variable <- variables[1]

for (i_variable in variables) {
  temp <- read_csv(paste(
    path_preprocessing,
    paste("GLODAPv2_2016_MappedClimatology_", i_variable, ".csv", sep = ""),
    sep = ""
  ))
  
  if (exists("GLODAP_predictors")) {
    GLODAP_predictors <- full_join(GLODAP_predictors, temp)
  }
  
  if (!exists("GLODAP_predictors")) {
    GLODAP_predictors <- temp
  }
}

rm(temp, i_variable, variables)

GLODAP_predictors <- GLODAP_predictors %>%
  rename(phosphate = PO4,
         nitrate = NO3,
         talk = TAlk)

# removed na's attributable to slightly different coverage of predictor fields
GLODAP_predictors <- GLODAP_predictors %>%
  select(lon, lat, depth, basin_AIP,
         phosphate, oxygen, nitrate,
         any_of(params_local$MLR_predictors)) %>% 
  drop_na()

```


```{r load_GLODAPv2_2016b_MappedClimatologies_surface_mapping}

variables <-
  c("PO4", "silicate", "TAlk", "TCO2")

for (i_variable in variables) {
  temp <- read_csv(paste(
    path_preprocessing,
    paste("GLODAPv2_2016_MappedClimatology_", i_variable, ".csv", sep = ""),
    sep = ""
  ))
  
  if (exists("GLODAP_predictors_CO2")) {
    GLODAP_predictors_CO2 <- full_join(GLODAP_predictors_CO2, temp)
  }
  
  if (!exists("GLODAP_predictors_CO2")) {
    GLODAP_predictors_CO2 <- temp
  }
}

rm(temp, i_variable, variables)


GLODAP_predictors_CO2 <- GLODAP_predictors_CO2 %>%
  rename(phosphate = PO4)

# removed na's attributable to slightly different coverage of predictor fields
GLODAP_predictors_CO2 <- GLODAP_predictors_CO2 %>%
  drop_na()

```

- World Ocean Atlas 2018  
  - Salinity
  - Temperature
  - Neutral density

```{r load_WOA18}

WOA18_predictors <-
  read_csv(paste(path_preprocessing,
                 "WOA18_sal_temp.csv",
                 sep = ""))

```


## Model

- Cmorized model runA climatologies in 2007, with variables  
  - Phosphate (-> phosphate_star)
  - Silicate
  - Oxygen 
  - AOU (already calculated)
  - TAlk (surface only)
  - TCO2 (surface only)


```{r load_model_climatologies_interior_mapping}

climatology <-
  read_csv(paste0(path_preprocessing_model,
                  "Climatology_A_2007_all_models/",
                  params_local$gobm,
                  "_climatology_runA_2007.csv"
                  ))

climatology <- climatology %>%
  rename(aou = AOU, TCO2 = tco2, TAlk = talk) %>% 
  select(-aou)

climatology <- inner_join(climatology, basinmask)

model_predictors <- climatology %>%
  select(lon, lat, depth, basin_AIP, gamma, 
         sal, phosphate, oxygen, temp, nitrate,
         any_of(params_local$MLR_predictors))

```


```{r load_model_climatologies_surface_mapping}

model_predictors_surface <- climatology %>%
  select(lon,
         lat,
         depth,
         basin_AIP,
         gamma,
         sal,
         temp,
         TCO2,
         TAlk,
         phosphate,
         silicate)

```


```{r load_gamma_offset_values}

mean_gamma_offset_Atl  <-
  read_rds(paste(path_version_data,
                  "mean_gamma_offset_Atl.rds",
                  sep = ""))

mean_gamma_offset_Indo_Pac  <-
  read_rds(paste(path_version_data,
                  "mean_gamma_offset_Indo_Pac.rds",
                  sep = ""))
```


# Join data

## WOA18 + GLODAP

WOA18 and GLODAP predictor climatologies are merged. Only horizontal grid cells with observations from both predictor fields are kept.

```{r join_WOA18_GLODAP_predictors}

# join deep water predictors
predictors <- full_join(
  GLODAP_predictors,
  WOA18_predictors)

rm(GLODAP_predictors)

predictors <- predictors %>% 
  drop_na()

# join surface water predictors
predictors_surface <- full_join(
  GLODAP_predictors_CO2,
  WOA18_predictors)

predictors_surface <- predictors_surface %>% 
  drop_na()

```

## Model + observational


```{r join_model_observational_predictors}

# join deep water predictors
predictors <- bind_rows(
  predictors %>% mutate(data_source = "obs"),
  model_predictors %>% mutate(data_source = "mod"))

rm(model_predictors)

# join surface water predictors
predictors_surface <- bind_rows(
  predictors_surface %>% mutate(data_source = "obs"),
  model_predictors_surface %>% mutate(data_source = "mod"))

rm(model_predictors_surface)


```


# Apply density threshold

The predictor field was split into two parts:

1. Deep water: neutral densities >= `r params_local$gamma_min` and depth >= `r params_local$depth_min`m
2. Shallow water: rest

```{r apply_gamma_offset_mod}

predictors <- predictors %>%
  mutate(
    gamma = if_else(
      basin_AIP %in% c("Indian", "Pacific") &
        data_source == "mod",
      gamma - mean_gamma_offset_Indo_Pac,
      gamma
    ),
    gamma = if_else(
      basin_AIP %in% c("Atlantic") &
        data_source == "mod",
      gamma - mean_gamma_offset_Atl,
      gamma
    )
  )

predictors_surface <- predictors_surface %>%
  mutate(
    gamma = if_else(
      basin_AIP %in% c("Indian", "Pacific") &
        data_source == "mod",
      gamma - mean_gamma_offset_Indo_Pac,
      gamma
    ),
    gamma = if_else(
      basin_AIP %in% c("Atlantic") &
        data_source == "mod",
      gamma - mean_gamma_offset_Atl,
      gamma
    )
  )

```



```{r subset_surface_WOA18_GLODAP_predictors}

predictors_surface <- predictors_surface %>% 
  group_by(data_source) %>% 
  mutate(surface_depth = min(depth, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(depth == surface_depth) %>% 
  select(-surface_depth)

```


## Apply basin mask

Data outside the WOA18 basin mask were removed for further analysis.

```{r join_basin_mask}

predictors <- inner_join(predictors, basinmask)
predictors_surface <- inner_join(predictors_surface, basinmask)

```

## Control plots

Plots below are generated to control successful merging of data sets.

### Maps interior

```{r joined_climatology_control_PO4, fig.asp=0.6}

predictors %>%
  group_split(data_source) %>%
  # head(1) %>%
  map(~ p_map_climatology(
    df = .x,
    var = "phosphate",
    subtitle_text = paste("Data source: ", unique(.x$data_source))
  ))


```

```{r joined_climatology_control_temp, fig.asp=0.6}

predictors %>%
  group_split(data_source) %>%
  map(~ p_map_climatology(
    df = .x,
    var = "temp",
    subtitle_text = paste("Data source: ", unique(.x$data_source))
  ))

```

### Maps surface

```{r joined_climatology_control_TAlk_surface, fig.asp=0.6}

predictors_surface %>%
  group_split(data_source) %>%
  map(~ p_map_climatology(
    df = .x,
    var = "TAlk",
    subtitle_text = paste("Data source: ", unique(.x$data_source))
  ))

```

```{r joined_climatology_control_TCO2_surface, fig.asp=0.6}
predictors_surface %>%
  group_split(data_source) %>%
  map(~ p_map_climatology(
    df = .x,
    var = "TCO2",
    subtitle_text = paste("Data source: ", unique(.x$data_source))
  ))
```

```{r joined_climatology_control_sal_surface, fig.asp=0.6}
predictors_surface %>%
  group_split(data_source) %>%
  map(~ p_map_climatology(
    df = .x,
    var = "sal",
    subtitle_text = paste("Data source: ", unique(.x$data_source))
  ))
```

```{r joined_climatology_control_temp_surface, fig.asp=0.6}
predictors_surface %>%
  group_split(data_source) %>%
  map(~ p_map_climatology(
    df = .x,
    var = "temp",
    subtitle_text = paste("Data source: ", unique(.x$data_source))
  ))
```


### Predictor profiles

Likewise, predictor profiles for the North Atlantic (`r params_global$lat_Atl_profile` / `r params_global$lon_Atl_section`) are plotted to control successful merging of the data sets.

```{r predictor_profiles_N_Atl, fig.asp=1.5}

# subset data
N_Atl <- predictors %>% 
  filter(lat == params_global$lat_Atl_profile,
         lon == params_global$lon_Atl_section)

# pivot table to long format
N_Atl <- N_Atl %>% 
  select(-c(basin, basin_AIP)) %>% 
  pivot_longer(c(any_of(params_local$MLR_predictors), gamma),
               names_to = "parameter", values_to = "value")

# plot profiles
N_Atl %>% 
  ggplot(aes(value, depth, col=data_source)) +
  geom_path() +
  geom_point() +
  scale_y_reverse() +
  facet_wrap(~parameter,
             scales = "free_x",
             ncol = 2)

rm(N_Atl)

```

# Prepare predictor fields

Additional predictor fields need to calculated from available climatologies

## AOU

### Calculation

AOU was calculated as the difference between saturation concentration and observed concentration. 
**CAVEAT**: Algorithms used to calculate oxygen saturation concentration are not yet identical in GLODAP data set (fitting) and predictor climatologies (mapping).

```{r calculate_aou_climatology}

predictors <- predictors %>% 
  mutate(aou = b_aou(
    sal = sal,
    tem = temp,
    depth = depth,
    oxygen = oxygen))


```

### Maps

```{r aou_climatology_maps, fig.asp=0.6}

predictors %>%
  group_split(data_source) %>%
  map(
    ~ p_map_climatology(
      df = .x,
      var = "aou",
      col = "divergent",
      subtitle_text = paste("Data source: ", unique(.x$data_source))
    )
  )

```

### Global section

```{r aou_climatology_section}

predictors %>%
  group_split(data_source) %>%
  map(
    ~ p_section_global(
      df = .x,
      var = "aou",
      col = "divergent",
      subtitle_text = paste("Data source: ", unique(.x$data_source))
    )
  )

```

## Isoneutral slabs

The following boundaries for isoneutral slabs were defined:

- Atlantic: `r params_local$slabs_Atl`
- Indo-Pacific: `r params_local$slabs_Ind_Pac`

Continuous neutral density (gamma) values based on WOA18 are grouped into isoneutral slabs.

```{r cut_isoneutral_slabs}

predictors <- m_cut_gamma(predictors, "gamma")

```

# Plot al predictor sections

## Deep waters

Predictor sections along with lines are shown below for each (potential) predictor variable.

```{r predictors_observations_sections_map, fig.asp=0.8}

map +
  geom_bin2d(data = predictors,
             aes(lon, lat),
             binwidth = c(1,1)) +
  geom_vline(xintercept = params_global$longitude_sections_regular,
             col = "white") +
  scale_fill_viridis_c(direction = -1) +
  facet_wrap(~ data_source, ncol = 1)

```

```{r all_predictor_sections, fig.asp=1.2}

for (i_var in params_local$MLR_predictors) {
  # i_var <- params_local$MLR_predictors[1]
  print(predictors %>%
          group_split(data_source) %>%
          map(
            ~ p_section_climatology_regular(
              df = .x,
              var = i_var,
              subtitle_text = paste("Data source: ", unique(.x$data_source))
            )
          ))
}

```

## Surface waters

Predictor sections along with lines are shown below for each (potential) predictor variable.

```{r predictors_observations_sections_map_surface, fig.asp=0.6}

if (nrow(predictors_surface) > 0) {
  map +
    geom_bin2d(data = predictors_surface,
               aes(lon, lat),
               binwidth = c(1, 1)) +
    geom_vline(xintercept = params_global$longitude_sections_regular,
               col = "white") +
    scale_fill_viridis_c(direction = -1) +
    theme(legend.position = "bottom") +
    facet_wrap( ~ data_source)
}

```

```{r all_predictor_sections_surface, fig.asp=1.2}

vars <-
  c(
    "gamma",
    "sal",
    "temp",
    "TCO2",
    "TAlk"
  )

# i_var <- vars[1]

for (i_var in vars) {
  print(
    predictors_surface %>%
      group_split(data_source) %>%
      map(
        ~ p_section_climatology_regular(
          df = .x,
          var = i_var,
          surface = "y",
          subtitle_text = paste("Data source: ", unique(.x$data_source))
        )
      )
  )
}


```

# Write csv

```{r write_joined_predictor_fields}

predictors %>%
  write_csv(paste(path_version_data,
                  "W18_st_G16_opsn.csv",
                  sep = ""))

predictors_surface %>%
  write_csv(paste(path_version_data,
                  "W18_st_G16_opsn_surface.csv",
                  sep = ""))

```

