---
title: "Mapping target variable"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}
params_local <-
  read_rds(here::here("data/auxillary",
                       "params_local.rds"))

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

# Scope

Map 3d fields of the predicted target variable (either tco2, cstar or cstar_tref) based on fitted MLRs and climatological fields of predictor variables, for each era.

# Predictor fields

Currently, we use combined predictor fields:

- WOA18: S, T, and derived variables
- GLODAP16: Oxygen, PO4, NO3, Silicate, and derived variables

```{r read_predictor_file}

predictors <-
  read_csv(paste(path_version_data,
                 "W18_st_G16_opsn.csv",
                 sep = ""))

```


# Load MLR models

```{r load_eMLR_models}

lm_best_target <-
  read_csv(paste(path_version_data,
                 "lm_best_target.csv",
                 sep = ""))

```

# Merge MLRs + climatologies

```{r merge_model_coeff_predictor_climatology}

lm_best_target <- lm_best_target %>%
  mutate(model = str_remove(model, paste(params_local$MLR_target, "~ ")))

target <- full_join(predictors, lm_best_target)

rm(predictors, lm_best_target)

```


# Map target variable

## Apply MLRs to predictor

```{r calculate_target}

target <- b_target_model(target)

target <- target %>% 
  select(lon, lat, depth, era, basin, basin_AIP, gamma,
         params_local$MLR_target)

```

```{r target_average}

target_average <- m_target_model_average(target)

rm(target)

target_average <- m_cut_gamma(target_average, "gamma")

```

## Mean target variable sections

For each basin and era combination, the zonal mean of the target variable is calculated. Likewise, sd is calculated for the averaging of the mean basin fields.

```{r Calculate_target_mean_sections}

target_average_zonal <- m_target_zonal_mean(target_average)

target_average_zonal <- m_cut_gamma(target_average_zonal, "gamma_mean")

```

# Write csv

```{r write_target_files}

target_average %>%
  write_csv(paste(path_version_data,
                  "target_3d.csv",
                  sep = ""))

target_average_zonal %>%
  write_csv(paste(path_version_data,
                  "target_zonal.csv",
                  sep = ""))

```
