---
title: "Mapping cstar"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = here::here("analysis/child/setup.Rmd")}
```

# Predictor fields

Currently, we use combined predictor fields:

- WOA18: S, T, and derived variables
- GLODAP16: Oxygen, PO4, NO3, Silicate, and derived variables

```{r read_predictor_file}

predictors <- 
    read_csv(here::here("data/mapping",
                         "W18_st_G16_opsn.csv"))

```


# Load MLR models

```{r load_eMLR_models}

lm_all_wide_cstar <-
  read_csv(here::here("data/eMLR",
                       "lm_all_wide_cstar.csv"))

```

# Merge MLRs + climatologies

```{r merge_model_coeff_predictor_climatology}

lm_all_wide_cstar <- lm_all_wide_cstar %>% 
  mutate(model = str_remove(model, "cstar ~ "))
         
cstar <- full_join(predictors, lm_all_wide_cstar)

rm(predictors, lm_all_wide_cstar)

```


# Map cstar

## Apply MLRs to predictor

```{r calculate_cstar}

cstar <- b_cstar_model(cstar)

cstar <- cstar %>% 
  select(lon, lat, depth, era, basin, cstar, gamma)

```

```{r cstar_average}

cstar_average <- m_cstar_model_average(cstar)

rm(cstar)

cstar_average <- m_cut_gamma(cstar_average, "gamma")

```

## Mean cstar sections

For each basin and era combination, the zonal mean cstar is calculated. Likewise, sd is calculated for the averaging of the mean basin fields.

```{r Calculate_cstar_mean_sections}

cstar_average <- left_join(cstar_average,
                           basinmask %>% select(-basin))


cstar_average_zonal <- m_cstar_zonal_mean(cstar_average)

cstar_average_zonal <- m_cut_gamma(cstar_average_zonal, "gamma_mean")

```

# Write csv

```{r write_cstar_files}

cstar_average %>%
    write_csv(here::here("data/output",
                         "cstar_3d.csv"))

cstar_average_zonal %>%
    write_csv(here::here("data/output",
                         "cstar_zonal.csv"))

```
