---
title: "Mapping"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r read_params_local, include = FALSE}
params_local <-
  read_rds(here::here("data/auxillary",
                       "params_local.rds"))

```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")

path_version_data     <-
  paste(path_observations,
        params_local$Version_ID,
        "/data/",
        sep = "")

path_version_figures  <-
  paste(path_observations,
        params_local$Version_ID,
        "/figures/",
        sep = "")
```

```{r select_basin_mask, include=FALSE}

basinmask <- basinmask %>% 
  filter(MLR_basins == params_local$MLR_basins) %>% 
  select(-MLR_basins)

```

```{r load_libraries_specific, include=FALSE}
library(metR)
library(marelac)
library(gsw)
```

# Required data

All required data sets were subsetted spatially in the read-in section *Data base*.
Currently, following data sets are used for mapping:

## GLODAPv2_2016b_MappedClimatologies

Following variables are currently used:  

- Phosphate (+Phosphate*)
- Silicate
- Oxygen (+AOU)
- TAlk (surface only)
- TCO2 (surface only)


```{r load_GLODAPv2_2016b_MappedClimatologies}

variables <-
  c("oxygen", "PO4", "silicate")

# i_variable <- variables[1]

for (i_variable in variables) {
  temp <- read_csv(paste(
    path_preprocessing,
    paste("GLODAPv2_2016_MappedClimatology_", i_variable, ".csv", sep = ""),
    sep = ""
  ))
  
  if (exists("GLODAP_predictors")) {
    GLODAP_predictors <- full_join(GLODAP_predictors, temp)
  }
  
  if (!exists("GLODAP_predictors")) {
    GLODAP_predictors <- temp
  }
}

rm(temp, i_variable, variables)

GLODAP_predictors <- GLODAP_predictors %>%
  rename(phosphate = PO4)

# removed na's attributable to slightly different coverage of predictor fields
GLODAP_predictors <- GLODAP_predictors %>%
  drop_na()

```


```{r load_GLODAPv2_2016b_MappedClimatologies_surface}

variables <-
  c("PO4", "silicate", "TAlk", "TCO2")

for (i_variable in variables) {
  temp <- read_csv(paste(
    path_preprocessing,
    paste("GLODAPv2_2016_MappedClimatology_", i_variable, ".csv", sep = ""),
    sep = ""
  ))
  
  if (exists("GLODAP_predictors_CO2")) {
    GLODAP_predictors_CO2 <- full_join(GLODAP_predictors_CO2, temp)
  }
  
  if (!exists("GLODAP_predictors_CO2")) {
    GLODAP_predictors_CO2 <- temp
  }
}

rm(temp, i_variable, variables)


GLODAP_predictors_CO2 <- GLODAP_predictors_CO2 %>%
  rename(phosphate = PO4)

# removed na's attributable to slightly different coverage of predictor fields
GLODAP_predictors_CO2 <- GLODAP_predictors_CO2 %>%
  drop_na()

```


## World Ocean Atlas 2018

- Salinity
- Temperature
- Neutral density

```{r load_WOA18}

WOA18_predictors <-
  read_csv(paste(path_preprocessing,
                 "WOA18_sal_temp.csv",
                 sep = ""))

```

# Join WOA18 + GLODAP

WOA18 and GLODAP predictor climatologies are merged. Only horizontal grid cells with observations from both predictor fields are kept.

**CAVEAT**: Coverage of GLODAP climatologies differs slightly for parameters (some are NA in some regions)

```{r join_WOA18_GLODAP_predictors}

# join deep water predictors
predictors <- full_join(
  GLODAP_predictors,
  WOA18_predictors)

rm(GLODAP_predictors)

predictors <- predictors %>% 
  drop_na()


# join surface water predictors
predictors_surface <- full_join(
  GLODAP_predictors_CO2,
  WOA18_predictors)

predictors_surface <- predictors_surface %>% 
  drop_na()

```


## Apply density threshold

The predictor field was split into two parts:

1. Deep water: neutral densities >= `r params_local$gamma_min` and depth >= `r params_local$depth_min`m
2. Shallow water: rest

```{r split_WOA18_GLODAP_predictors_deep_surface}

# predictors for deep waters
predictors <- predictors %>%
  filter(depth >= params_local$depth_min | gamma >= params_local$gamma_min)

# predictors for surface waters
predictors_surface <- predictors_surface %>%
  filter(depth < params_local$depth_min,
         gamma < params_local$gamma_min)


```


## Apply basin mask

Data outside the WOA18 basin mask were removed for further analysis.

```{r join_basin_mask}

predictors <- inner_join(predictors, basinmask)
predictors_surface <- inner_join(predictors_surface, basinmask)

```

## Control plots

### Maps

Three maps are generated to control successful merging of data sets.

```{r joined_climatology_control_PO4, fig.asp=0.6}
p_map_climatology(
  df = predictors, 
  var = "phosphate")
```

```{r joined_climatology_control_temp, fig.asp=0.6}
p_map_climatology(
  df = predictors, 
  var = "temp")
```

### Maps surface

Three maps are generated to control successful merging of data sets.

```{r joined_climatology_control_TAlk_surface, fig.asp=0.6}
p_map_climatology(
  df = predictors_surface, 
  var = "TAlk")
```

```{r joined_climatology_control_TCO2_surface, fig.asp=0.6}
p_map_climatology(
  df = predictors_surface, 
  var = "TCO2")
```

```{r joined_climatology_control_sal_surface, fig.asp=0.6}
p_map_climatology(
  df = predictors_surface, 
  var = "sal")
```

```{r joined_climatology_control_temp_surface, fig.asp=0.6}
p_map_climatology(
  df = predictors_surface, 
  var = "temp")
```


### Predictor profiles

Likewise, predictor profiles for the North Atlantic (`r params_global$lat_Atl_profile` / `r params_global$lon_Atl_section`) are plotted to control successful merging of the data sets.

```{r predictor_profiles_N_Atl, fig.asp=1.5}

N_Atl <- predictors %>% 
  filter(lat == params_global$lat_Atl_profile,
         lon == params_global$lon_Atl_section)

N_Atl <- N_Atl %>% 
  select(-c(basin, basin_AIP)) %>% 
  pivot_longer(oxygen:gamma, names_to = "parameter", values_to = "value")

N_Atl %>% 
  ggplot(aes(value, depth)) +
  geom_path() +
  geom_point() +
  scale_y_reverse() +
  facet_wrap(~parameter,
             scales = "free_x",
             ncol = 2)

rm(N_Atl)

```

# Prepare predictor fields

## PO~4~* calculation 

The predictor PO~4~* was be calculated according to Clement and Gruber (2018), ie based on oxygen. Please note that an errornous equations for PO4* calculation is given in the supplement of Gruber et al (2019), based on nitrate.

```{r calculate_phosphate_star_clement}

predictors <- predictors %>% 
  mutate(phosphate_star = b_phosphate_star(phosphate, oxygen))

```

### Maps

```{r PO4_star_climatology_maps, fig.asp=0.6}

p_map_climatology(
  df = predictors,
  var = "phosphate_star",
  col = "divergent")

```

### Sections

```{r PO4_star_climatology_section}

p_section_global(
  df = predictors,
  var = "phosphate_star",
  col = "divergent")

```


## AOU

### Calculation

AOU was calculated as the difference between saturation concentration and observed concentration. 
**CAVEAT**: Algorithms used to calculate oxygen saturation concentration are not yet identical in GLODAP data set (fitting) and predictor climatologies (mapping).

```{r calculate_aou_climatology}

predictors <- predictors %>% 
  mutate(aou = b_aou(
    sal = sal,
    tem = temp,
    depth = depth,
    oxygen = oxygen))

```

### Maps

```{r aou_climatology_maps, fig.asp=0.6}

p_map_climatology(
  df = predictors,
  var = "aou",
  col = "divergent")

```

### Sections

```{r aou_climatology_section}

p_section_global(
  df = predictors,
  var = "aou",
  col = "divergent")

```

## Isoneutral slabs

The following boundaries for isoneutral slabs were defined:

- Atlantic: `r params_local$slabs_Atl`
- Indo-Pacific: `r params_local$slabs_Ind_Pac`

Continuous neutral density (gamma) values based on WOA18 are grouped into isoneutral slabs.

```{r cut_isoneutral_slabs}

predictors <- m_cut_gamma(predictors, "gamma")

```

# Plot al predictor sections

## Deep waters

Predictor sections along with lines are shown below for each (potential) predictor variable.

```{r predictors_observations_sections_map, fig.asp=0.6}

map +
  geom_bin2d(data = predictors,
             aes(lon, lat),
             binwidth = c(1,1)) +
  geom_vline(xintercept = params_global$longitude_sections_regular,
             col = "white") +
  scale_fill_viridis_c(direction = -1) +
  theme(legend.position = "bottom")

```

```{r all_predictor_sections, fig.asp=1.2}

vars <-
  c(
    "gamma",
    "sal",
    "temp",
    "phosphate",
    "phosphate_star",
    "oxygen",
    "aou",
    "silicate"
  )

# i_var <- vars[1]

for (i_var in vars) {
  print(
    p_section_climatology_regular(
      df = predictors,
      var = i_var)
    )
}

```

## Surface waters

Predictor sections along with lines are shown below for each (potential) predictor variable.

```{r predictors_observations_sections_map_surface, fig.asp=0.6}

map +
  geom_bin2d(data = predictors_surface,
             aes(lon, lat),
             binwidth = c(1,1)) +
  geom_vline(xintercept = params_global$longitude_sections_regular,
             col = "white") +
  scale_fill_viridis_c(direction = -1) +
  theme(legend.position = "bottom")

```

```{r all_predictor_sections_surface, fig.asp=1.2}

vars <-
  c(
    "gamma",
    "sal",
    "temp",
    "TCO2",
    "TAlk"
  )

for (i_var in vars) {
  print(
    p_section_climatology_regular(
      df = predictors_surface,
      var = i_var)
    )
}

```

# Write csv

```{r write_joined_predictor_fields}

predictors %>%
  write_csv(paste(path_version_data,
                  "W18_st_G16_opsn.csv",
                  sep = ""))

predictors_surface %>%
  write_csv(paste(path_version_data,
                  "W18_st_G16_opsn_surface.csv",
                  sep = ""))

```

