---
title: "Anthropogenic CO2 from 1994 to 2007"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = here::here("analysis/child/setup.Rmd")}
```

# Data source

- Anthropogenic CO2 estimates (1800-1994) by Sabine et al. (2004) downloaded in August 2020 from [NOAA/NCEI Ocean Carbon Data System (OCADS)](ftp://ftp.nodc.noaa.gov/pub/data.nodc/ncei/ocads/data/0001644/){target="_blank"}

# Read ncdfs

```{r read_Sabine_2004}

AnthCO2_data <-
  read_csv(
    "data/input/GLODAPv1_1/GLODAP_gridded.data/AnthCO2.data/AnthCO2.data.txt",
    col_names = FALSE,
    na = "-999",
    col_types = list(.default = "d")
  )

Depth_centers <-
  read_file("data/input/GLODAPv1_1/GLODAP_gridded.data/Depth.centers.txt")

Depth_centers <- Depth_centers %>%
  str_split(",") %>%
  as_vector()

Lat_centers <-
  read_file("data/input/GLODAPv1_1/GLODAP_gridded.data/Lat.centers.txt")

Lat_centers <- Lat_centers %>%
  str_split(",") %>%
  as_vector()

Long_centers <-
  read_file("data/input/GLODAPv1_1/GLODAP_gridded.data/Long.centers.txt")

Long_centers <- Long_centers %>%
  str_split(",") %>%
  as_vector()

names(AnthCO2_data) <- Lat_centers

Long_Depth <-
  expand_grid(depth = Depth_centers, lon = Long_centers) %>%
  mutate(lon = as.numeric(lon),
         depth = as.numeric(depth))

cant_3d <- bind_cols(AnthCO2_data, Long_Depth)

cant_3d <- cant_3d %>%
  pivot_longer(1:180, names_to = "lat", values_to = "cant") %>%
  mutate(lat = as.numeric(lat))

cant_3d <- cant_3d %>%
  drop_na()

cant_3d <- cant_3d %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))

rm(AnthCO2_data,
   Long_Depth,
   Depth_centers,
   Lat_centers,
   Long_centers)

```

# Apply basin mask

```{r apply_basin_mask}

cant_3d <- inner_join(cant_3d, basinmask)

```


# Inventory calculation

```{r calculate_inventory}

cant_3d <- cant_3d %>% 
  mutate(cant_pos = if_else(cant <= 0, 0, cant),
         eras = "1800-1994")

cant_inv <- m_cant_inv(cant_3d) 

# cant_inv <- cant_3d %>% 
#   filter(depth <= parameters$inventory_depth) %>% 
#   group_by(lon, lat, basin, basin_AIP) %>% 
#   summarise(cant_inv = sum(layer_inv_pos, na.rm = TRUE) / 1000,
#             cant_inv_incl_neg = sum(layer_inv, na.rm = TRUE) / 1000) %>% 
#   ungroup()

```

# Cant plots

Below, following subsets of the climatologies are plotted for all relevant parameters:  

- Horizontal planes at `r parameters$depth_levels`m
- Meridional sections at longitudes: `r parameters$longitude_sections_basin`

Section locations are indicated as white lines in maps.


## Horizontal plane maps

```{r cant_maps, fig.asp=0.6}

p_map_climatology(
  df = cant_3d,
  var = "cant",
  col = "divergent")

```

## Sections

```{r cant_sections}

p_section_global(
  df = cant_3d,
  var = "cant",
  col = "divergent")

```

## Sections at regular longitudes

```{r cant_sections_regular, fig.asp=2}

p_section_climatology_regular(
  df = cant_3d,
  var = "cant",
  col = "divergent")

```

## Zonal mean section

```{r calculate_zonal_mean_section}

cant_zonal <- m_zonal_mean_section(cant_3d %>% select(-basin))

```

## Inventory maps

```{r cant_inv_pos_maps, fig.asp=0.6}

p_map_cant_inv(
  df = cant_inv,
  var = "cant_pos_inv",
  breaks = seq(0,max(cant_inv$cant_pos_inv),5))

```


## Write files

```{r write_Sabine_2004_cant_file}

cant_3d %>% 
  write_csv(here::here("data/interim",
                       "S04_cant_3d.csv"))

cant_inv %>% 
  write_csv(here::here("data/interim",
                       "S04_cant_inv.csv"))

cant_zonal %>% 
  write_csv(here::here("data/interim",
                       "S04_cant_zonal.csv"))


```
