---
title: "Anthropogenic CO2 from 1994 to 2007"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r parent, child = here::here("analysis/child/setup.Rmd")}
```

# Libraries

Loading libraries specific to the the analysis performed in this section.

```{r load_libraries_specific, message=FALSE, warning=FALSE}
library(tidync)
```


# Data source

- Anthropogenic CO2 estimates (1994-2007) by Gruber et al. (2019) downloaded in August 2020 from [NOAA/NCEI Ocean Carbon Data System (OCADS)](https://www.nodc.noaa.gov/ocads/oceans/ndp_100/ndp100.html){target="_blank"}

# Read ncdfs

## Public data sets

The publicly available data sets contain only positive Cant estimates.

```{r read_Gruber_dcant}

dcant <- tidync(here::here("data/input/Gruber_2019",
                               "dcant_emlr_cstar_gruber_94-07_vs1.nc"))

dcant <- dcant %>%  activate(DCANT_01)
dcant <- dcant %>% hyper_tibble()

# harmonize column names and coordinates
dcant <- dcant %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         depth = DEPTH,
         cant_pos = DCANT_01) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```

```{r read_Gruber_dcant_inv}

dcant_inv <- tidync(here::here("data/input/Gruber_2019",
                               "inv_dcant_emlr_cstar_gruber_94-07_vs1.nc"))

dcant_inv <- dcant_inv %>%  activate(DCANT_INV01)
dcant_inv <- dcant_inv %>% hyper_tibble()

# harmonize column names and coordinates
dcant_inv <- dcant_inv %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         cant_pos_inv = DCANT_INV01) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon)) %>% 
  mutate(eras = "JGOFS_GO")

```

## Raw data

Internally available data sets also contain negative Cant estimates, as they are contained in the "raw" output of the eMLR mapping step.

```{r read_Gruber_V101}

V101 <- tidync(here::here("data/input/Gruber_2019",
                               "Cant_V101new.nc"))

V101 <- V101 %>%  activate(Cant)
V101 <- V101 %>% hyper_tibble()

# harmonize column names and coordinates
V101 <- V101 %>% 
  rename(lon = longitude,
         lat = latitude,
         cant = Cant) %>% 
  filter(cant != -999) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```


# Apply basin mask

```{r apply_basin_mask}

# basinmask <- basinmask %>% 
#   select(-basin)

dcant <- inner_join(dcant, basinmask)
dcant_inv <- inner_join(dcant_inv, basinmask)
V101 <- inner_join(V101, basinmask)

```

# Join pos and all Cant

```{r join_V101_and_all}

cant_3d <- inner_join(dcant, V101)
cant_3d <- cant_3d %>% 
  mutate(eras = "JGOFS_GO")

rm(dcant, V101)

```

# Zonal mean section

```{r calculate_zonal_mean_section}

cant_zonal <- m_zonal_mean_section(cant_3d %>% select(-basin))

```

# Column inventory 

## From 3d fields

```{r calculate_column_inventory}

cant_inv <- m_cant_inv(cant_3d)

```

### Cant - all

```{r G19_cant_inv_maps}

p_map_cant_inv(
  df = cant_inv,
  var = "cant_inv",
  col = "divergent")

```

### Cant - pos

```{r G19_cant_pos_inventory_maps}

p_map_cant_inv(
  df = cant_inv,
  var = "cant_pos_inv")

```

## From pubished inventory data

```{r G19_cant_pos_inventory_maps_published}

p_map_cant_inv(
  df = dcant_inv,
  var = "cant_pos_inv")

```

## Published - 3d

```{r G19_cant_offset_inventory}

cant_offset <- inner_join(
  cant_inv %>% rename(cant_re = cant_pos_inv),
  dcant_inv %>% rename(cant_pub = cant_pos_inv)
)

cant_offset <- cant_offset %>% 
  mutate(delta_cant = cant_re - cant_pub)

p_map_cant_inv_offset(df = cant_offset,
                      var = "delta_cant",
                      breaks = seq(-3,3,0.25))

rm(cant_offset, dcant_inv)
```


# Cant plots

Below, following subsets of the climatologies are plotted for all relevant parameters:  

- Horizontal planes at `r parameters$depth_levels`m
- Meridional sections at longitudes: `r parameters$longitude_sections_basin`

Section locations are indicated as white lines in maps.


## Horizontal plane maps

### All values

```{r Cant_maps, fig.asp=0.6}

p_map_climatology(
  df = cant_3d,
  var = "cant",
  col = "divergent")

```

### Positive values

```{r Cant_maps_V101, fig.asp=0.6}

p_map_climatology(
  df = cant_3d,
  var = "cant_pos")

```


## Sections basin

### All values

```{r Cant_sections_V101, fig.asp=0.8}

p_section_global(
  df = cant_3d,
  var = "cant",
  col = "divergent")

```

### Positive values

```{r Cant_sections, fig.asp=0.8}

p_section_global(
  df = cant_3d,
  var = "cant_pos")
```


## Sections at regular longitudes

### All values

```{r Cant_sections_regular, fig.asp=2}

p_section_climatology_regular(
  df = cant_3d,
  var = "cant",
  col = "divergent")

```

### Positive values

```{r Cant_sections_regular_V101, fig.asp=2}

p_section_climatology_regular(
  df = cant_3d,
  var = "cant_pos")

```

# Write files

```{r write_Gruber_Cant_file}

cant_3d %>% 
  write_csv(here::here("data/interim",
                       "G19_cant_3d.csv"))

cant_inv %>% 
  write_csv(here::here("data/interim",
                       "G19_cant_inv.csv"))

cant_zonal %>% 
  write_csv(here::here("data/interim",
                       "G19_cant_zonal.csv"))

```

